### 2021-10-13 paper/paper.tex
added rough draft of paper

\documentclass[11pt]{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry} % page settings
\usepackage{float}
\usepackage[sort,numbers]{natbib}

\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}

\usepackage{amsmath}
\usepackage{amsthm, thmtools}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}   
\usepackage{enumitem}
\usepackage{listings}
\usepackage[hidelinks]{hyperref}
%\numberwithin{equation}

\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows}
\usepackage[makeroom,samesize]{cancel}

\usepackage{
nameref,
%\nameref
hyperref,
%\autoref
% n.b. \Autoref is defined by thmtools
cleveref,
% \cref
% n.b. cleveref after! hyperref
}

\usepackage{setspace}
\usepackage{lineno}
\title{No robust coexistence in a canonical model of plant-soil feedbacks}

\begin{document}
\maketitle
\setstretch{1.5}
\linenumbers

\begin{abstract}
  [Abstract]
\end{abstract}

\section{Introduction}

It has become well understood that reciprocal interactions between plants and the soil biota, known as plant-soil feedbacks (PSFs), play an important role in structuring the composition and dynamics of plant communities. PSFs operate alongside other factors, including abiotic drivers \cite{bennett2019mechanisms} and above-ground trophic interactions \cite{van2009empirical}, but are thought to be a key mechanism generating negative frequency-dependent feedbacks that promote coexistence and maintain plant diversity \cite{kulmatiski2008plant,van2013plant,bever2015maintenance}. The existence of PSFs has long been known \cite{van1993plant,bever1994feedback}, but our understanding of their importance -- particularly in relation to patterns of coexistence -- has developed rapidly in recent years \cite{klironomos2002feedback,petermann2008janzen,mangan2010negative}. Broad interest in PSFs was ignited by the development of simple mathematical models, which illustrated the potential of PSFs to mediate plant coexistence \cite{bever1997incorporating,bever2003soil,ke2015incorporating}. These models have played a guiding role for a wide range of empirical studies, as well \cite{kulmatiski2008plant,kulmatiski2011testing,pernilla2010plant}.

The first, and still most widely known and used, model for PSFs was introduced by Bever and colleagues in the 1990s \cite{bever1992ecological,bever1997incorporating,bever1999dynamics,bever2003soil}. In this framework, often referred to simply as the Bever model, each plant species is assumed to promote the growth of a specific soil component (i.e. associated bacteria, fungi, invertebrates, considered collectively) in the vicinity of individual plants. In turn, the fitness of each plant species is impacted by the relative frequency of different soil components. Starting from minimalist assumptions, Bever et al. \cite{bever1997incorporating} derived a set of differential equations to capture these dynamics. PSFs can be either positive (fitness of a plant species is increased by its corresponding soil component) or negative (a plant species experiences lower relative fitness in its own soil). Bever et al. introduced a single quantity to summarize whether PSFs are positive or negative, and showed that this value characterizes the dynamical behavior of the model. In the original Bever model of two plant species, positive PSFs lead to exclusion of one species, while negative PSFs result in neutral oscillations. It is thus widely suggested that negative PSFs help sustain coexistence in real-world plant communities \cite{kulmatiski2008plant,van2013plant}, perhaps with spatial asynchrony playing a role in stabilizing the cyclic dynamics \cite{revilla2013plant,bever2003soil}.

Subsequent studies have generalized this model to include, for example, more realistic functional forms \cite{umbanhowar2005simple, eppinga2006accumulation}, more explicit representations of the soil community \cite{bever2010rooting}, spatial structure \cite{eppinga2006accumulation,molofsky2002negative,suding2013consequences}, or additional processes such as direct competitive interactions between plants \cite{bever2003soil}. However, the original Bever model remains an important touchstone for the theory of PSFs \cite{ke2015incorporating,ke2020effects}, and informs empirical research through the interaction coefficient, $I_s$, derived by Bever et al., which is commonly measured and used to draw conclusions about coexistence in experimental studies. Despite the ubiquity of this model, and the fruitful interplay of theory and experiment in the PSF literature, extensions to communities with more than two or three species have appeared only rarely and recently (but see \cite{eppinga2018frequency,mack2019plant}). While PSF models motivate hypotheses and conclusions about species-rich natural communities, there is much still unknown about the behavior of these models with natural levels of diversity \cite{van2013plant}.

In this Letter, we extend the Bever model to include any number of plant species, and show that the model is equivalent to a special form of the replicator equation studied in evolutionary game theory \cite{hofbauer1998evolutionary}. In particular, this model corresponds to the class of bimatrix games, where there are two players (here, plants and soil components) which interact with asymmetric strategies and payoffs. The replicator dynamics of bimatrix games are well-studied, allowing us to characterize many properties of the Bever model with $n$ species. Surprisingly, using this equivalence, we show that coexistence of more than two species in this model is never robust.

\section{Results}

\subsection{Generalizing a classic PSF model}

Inspired by emerging empirical evidence for the important role of PSFs in plant community dynamics and coexistence \cite{van1993plant,bever1994feedback}, Bever et al. \cite{bever1997incorporating} introduced a simple mathematical model to investigate their behavior. In this model, two plant species, $1$ and $2$, grow exponentially with growth rates determined by the state of the soil biota in the system (note we adopt a different notation than Bever et al.). These effects of soil on plants are specified by parameters $\alpha_{ij}$, the growth rate of plant species $i$ in soil type $j$. There is a soil component corresponding to each plant species, which grows exponentially in the presence of its associated plant at a rate $\beta_i$. Bever et al. set an important precedent by considering dynamics of \emph{relative} abundances in such a system; starting from dynamics of the form

\begin{align} \label{2spp_absolute}
	\begin{cases}
	\frac{dx_i}{dt} &= x_i \left( \frac{\alpha_{ii} \, y_i + \alpha_{ij} \, y_j}{y_i + y_j} \right) \, , \quad  i, j = 1,2 \\
	\frac{dy_i}{dt} &= y_i \left( \frac{\beta_i \, x_i}{x_i + x_j} \right)
	\end{cases}
\end{align}
for the \emph{absolute} abundances of plants ($x_i$) and soil components ($y_i$), one considers the relative abundances (frequencies), $p_i = x_i \, / \sum_j x_j$ and $q_i = y_i \, / \sum_{j} y_j$. The dynamics for these frequencies are easily derived from Eq.~\ref{2spp_absolute}, and using the facts $p_i = 1 - p_j$ and $q_i = 1 - q_j$, can be written as:

\begin{align} \label{2spp_relative}
\begin{cases}
\frac{dp_i}{dt} &= p_i \, p_j \left( (\alpha_{ii} - \alpha_{ji}) \, q_i + (\alpha_{ij} - \alpha_{jj}) \, q_j \right) \, , \quad  i, j = 1,2 \\
\frac{dq_i}{dt} &= q_i \, q_j (\beta_i \, p_i - \beta_j \, p_j) \, .
\end{cases}
\end{align}
This model may admit a coexistence equilibrium where

\begin{align} \label{2spp_eq}
\begin{cases}
p_i^\star &= \frac{1}{1 + \frac{\beta_i}{\beta_j}} \, , \quad  i, j = 1,2 \\
q_i^\star &= \frac{\alpha_{jj} - \alpha_{ji}}{\alpha_{ii} - \alpha_{ji} - \alpha_{ij} + \alpha_{jj}} \, .
\end{cases}
\end{align}

A central finding of the analysis by Bever et al. was that the denominator of $q_i^*$, which they termed the ``interaction coefficient'', $I_s$ ($= \alpha_{ii} - \alpha_{ji} - \alpha_{ij} + \alpha_{jj}$), controls the model dynamics: When $I_s > 0$, which represents a community with positive feedbacks, the equilibrium in Eq.~\ref{2spp_eq} is unstable, and the two species cannot coexist. On the other hand, when $I_s < 0$, the equilibrium is neutrally stable, and the dynamics cycle around it, providing a form of non-equilibrium coexistence. In fact, these conclusions depend on the existence of a feasible equilibrium (i.e. positive equilibrium values), which further requires that $\alpha_{ii} < \alpha_{ji}$, for both $i, j = 1, 2$, in order for the model to exhibit coexistence \cite{bever1997incorporating,ke2015incorporating}.

This coexistence is fragile. Species frequencies oscillate neutrally, similar to the textbook example of Lotka-Volterra predator-prey dynamics. Any stochasticity, external forcing, or time variation in the model parameters can destroy these finely balanced oscillations and cause one species to go extinct \cite{revilla2013plant}. However, coupled with mechanisms that buffer the system from extinctions, such as migration between desynchronized patches or the presence of a seed bank, the negative feedbacks in this model might produce sustained coexistence \cite{revilla2013plant, bever2003soil}.

Of course, most natural plant communities feature more than two coexisting species, and it is precisely in the most diverse communities where mechanisms of coexistence hold the greatest interest \cite{van2013plant}. While it is not immediately clear how to generalize Eq.~\ref{2spp_relative} to more than two species, Eq.~\ref{2spp_absolute} is naturally extended by maintaining the assumption that the overall growth rate for any plant is a weighted average of its growth rate in each soil type:

\begin{align} \label{nspp_absolute}
\begin{cases}
\frac{dx_i}{dt} &= x_i \left(\sum_{j} \alpha_{ij} \, q_j \right) \, , \quad  i = 1, \dots n \\
\frac{dy_i}{dt} &= y_i \left( \beta_i \, p_i \right) \, .
\end{cases}
\end{align}
From Eq.~\ref{nspp_absolute}, one can derive the $n$-species analogue of Eq.~\ref{2spp_relative},

\begin{align} \label{nspp_relative}
\begin{cases}
\frac{dp_i}{dt} &= p_i \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right) \, , \quad  i = 1, \dots n \\
\frac{dq_i}{dt} &= q_i \left(\beta_{i} \, p_i - \sum_{j} \beta_{j} \, p_j \, q_j  \right) \,
\end{cases}
\end{align}
giving the dynamics for species and soil component frequencies. Eq.~\ref{nspp_relative} is conveniently expressed in matrix form as

\begin{align} \label{matrix_form}
\begin{cases}
\frac{d\bm{p}}{dt} &= D(\bm{p}) \left(A \bm{q} - (\bm{p}^T A \bm{q}) \bm{1} \right) \\
\frac{d\bm{q}}{dt} &= D(\bm{q}) \left(B \bm{p} - (\bm{q}^T B \bm{p}) \bm{1}  \right) \,
\end{cases}
\end{align}
where vectors are indicated in boldface (e.g. $\bm{p}$ is the vector of species frequencies $(p_1, p_2, \dots p_n)$ and $\bm{1}$ is a vector of $n$ ones) and $D(\bm{z})$ is the diagonal matrix with vector $\bm{z}$ on the diagonal. We have introduced the matrices $A = (\alpha_{ij})$ and $B = D(\beta_1, \beta_2, \dots \beta_n)$, specifying soil effects on plants and plant effects on soil, respectively. Because $\bm{p}$ and $\bm{q}$ are vectors of frequencies, they must sum to one: $\bm{1}^T \bm{p} = \bm{1}^T \bm{p} = 1$. Using these constraints, one can easily show that the Bever model (Eq.~\ref{2spp_relative}) is a special case of Eqs.~\ref{nspp_relative} and \ref{matrix_form} when $n = 2$.

\subsection{Equivalence to bimatrix game dynamics}

Systems taking the form of Eq.~\ref{matrix_form} are well-known and well-studied in evolutionary game theory. Our generalization of the Bever model is a special case of the \emph{replicator equation}, corresponding to the class of \emph{bimatrix games} \cite{taylor1979evolutionarily,hofbauer1996evolutionary,hofbauer1998evolutionary,cressman2014replicator}. Bimatrix games arise in diverse contexts, such as animal behavior \cite{taylor1979evolutionarily,selten1988note}, evolutionary theory \cite{hofbauer1998evolutionary,cressman2014replicator}, and economics \cite{friedman1991evolutionary}, where they model games with asymmetric players. In a bimatrix game, each player (here, the plant community and the soil) has a distinct set of strategies (plants species and soil components, respectively) and payoffs (realized growth rates).

Much is known about bimatrix game dynamics, and we can draw on this body of knowledge to characterize the behavior of the Bever model with $n$ species. Essential mathematical background and details are presented in the Supplemental Methods; for a detailed introduction to bimatrix games, see Hofbauer and Sigmund (\cite{hofbauer1998evolutionary}).

Under the mild condition that matrix $A$ is invertible, Eq.~\ref{matrix_form} admits a unique coexistence equilibrium given by $\bm{p}^\star = k_p \, B^{-1} \bm{1}$ and $\bm{q}^\star = k_q \, A^{-1} \bm{1}$, where $k_p = 1 / (\bm{1}^T B^{-1} \bm{1})$ and $k_q = 1 / (\bm{1}^T A^{-1} \bm{1})$ are constants of proportionality that ensure the equilibrium frequencies sum to one for both plants and soil. Because $B$ is a diagonal matrix, and all $\beta_i$ are assumed positive, the equilibrium plant frequencies, $\bm{p}^\star$, are always positive, as well. Thus, feasibility of the equilibrium hinges on the soil frequencies, $\bm{q}^\star$, which are all positive if the elements of $A^{-1} \bm{1}$ all share the same sign.

As we have seen, when the community consists of two species, the coexistence equilibrium, if feasible, can be either unstable or neutrally stable. In fact, the same is true for the $n$-species extension (and, more generally, for any bimatrix game dynamics \cite{eshel1983coevolutionary,selten1988note,hofbauer1998evolutionary}). This can be established using straightforward local stability analysis, after accounting for the relative abundance constraints, which imply $p_n = 1 - \sum_{i = 1}^{n-1} p_i$ and $q_n = 1 - \sum_{i = 1}^{n-1} q_i$. Using these substitutions, Eq.~\ref{nspp_relative} can be written as a system of $2 n - 2$ (rather than $2 n$) equations, and the community matrix for this reduced model has a very simple form (see Supplemental Methods). In particular, due to the bipartite structure of the model, the community matrix has all zero diagonal elements, which implies that the eigenvalues of this matrix sum to zero. These eigenvalues govern the stability of the coexistence equilibrium, and this property leaves two qualitatively distinct possibilities: either the eigenvalues have a mix of positive and negative real parts (in which case the equilibrium is unstable), or the eigenvalues all have zero real part (in which case the equilibrium is neutrally stable). Already, we can see that the model never exhibits equilibrium coexistence, regardless of the number of species.

Another notable property of bimatrix game dynamics is that the vector field defined by the model equations is divergence-free or incompressible (see \cite{hofbauer1998evolutionary} for a proof). The divergence theorem from vector calculus \cite{arfken1985mathematical} then tells us that Eq.~\ref{matrix_form} cannot have any attractors -- that is, regions of the phase space that ``pull in'' trajectories -- with multiple species. This rules out coexistence in a stable limit cycle or other non-equilibrium attractors (e.g. chaotic attractors). Thus, only the relatively fragile coexistence afforded by neutral oscillations is possible, as in the two-species model.

Based on the local stability properties of the coexistence equilibrium, Bever et al. concluded that such neutral cycles arise for two species when $\alpha_{11} < \alpha_{21}$ and $\alpha_{22} < \alpha_{12}$. The equivalence between their model and a bimatrix game with two strategies allows us to give a fuller picture of these cycles. Namely, we can identify a constant of motion for the two-species dynamics:

\begin{equation}
	H = (\alpha_{12} - \alpha_{22}) \log q_1 + (\alpha_{21} - \alpha_{11}) \log q_2 + \beta_2 \log p_1 + \beta_1 \log p_2 .
\end{equation} 
Using the chain rule and time derivatives in Eq.~\ref{2spp_relative}, it is easy to show that $\frac{dH}{dt} = 0$ for any plant and soil frequencies (see Supplemental Methods). The level curves of $H$ form closed orbits around the equilibrium when the equilibrium is neutrally stable. Thus, $H$ implicitly defines the trajectories of the model, and can be used to determine characteristics such as the amplitude of oscillations arising from particular initial frequencies.

Because neutral cycles provide the only possible form of coexistence in this model, a key question becomes whether and when neutral cycles with $n$ species can arise. Do the ``negative feedback'' conditions identified by Bever et al. generalize in richer communities? Indeed, they do; however, for more than two species, these conditions are very severe. The model in Eq.~\ref{matrix_form} supports oscillations with $n$ species -- for any $n$ -- if matrices $A$ and $B$ satisfy a precise relationship (Fig ???; see Supplemental Methods for details). In particular, the model parameters must satisfy the conditions $\alpha_{ij} = \gamma_i + \delta_j$ for some constants $\gamma_i, \delta_i$ in $i = 1, \dots, n$ (when $i \neq j$), and $\alpha_{ii} = \gamma_i + \delta_i- c \beta_i$ (where $c$ is a positive constant independent of $i$). In the language of bimatrix games, such systems are called \emph{rescaled zero-sum games} \cite{hofbauer1996evolutionary,hofbauer1998evolutionary}. It is a long-standing conjecture in evolutionary game theory that these parameterizations are the \emph{only} cases where $n$-species coexistence can occur \cite{hofbauer1996evolutionary,hofbauer2011deterministic}.

Ecologically, these conditions mean there is a fixed effect of each soil type and plant species identity, and the growth rate of plant $i$ in soil type $j$ is the additive combination of these two, with no interaction effects. The only exception is for plants growing in their own soil type, which must experience a fitness cost exactly proportional to the rate at which they promote growth of the soil type ($\beta_i$).  
These conditions clearly extend the intuitive notion that each plant must have a disadvantage in its corresponding soil type to allow coexistence. But the parameters of the model are constrained so strongly that we never expect to observe cycles with more than two species in practice. When $n > 2$, a great deal of fine-tuning is necessary to satisfy the zero-sum game condition; the probability that random parameters will be suitable is infinitesimally small. We confirm this numerically with simulations shown in Fig ???. Although $n$-species cycles are clearly possible (as in Fig ???), for parameters drawn independently at random, communities always collapse to zero, one, or two species, regardless of the initial richness.

Not only are parameter combinations permitting many-species oscillations rare, they are also extremely sensitive to small changes to the parameter values. The rescaled zero-sum condition imposes many exact equality constraints on the matrix $A$ (e.g. $\alpha_{ij} - \alpha_{ik} = \alpha_{lj} - \alpha_{lk}$ for all $i, j, k,$ and $l$). Even if mechanisms exist to generate the requisite qualitative patterns, inevitable quantitative variation in real-world communities will disrupt coexistence (Fig ???). Coexistence of $n > 2$ species -- even in the weak sense of neutral cycles -- is not robust to small changes in the model parameters.

Interestingly, the two-species model is not subject to the same fragility. It can be shown (see Supplemental Methods) that all $2 \times 2$ bimatrix games take the same general form as a rescaled zero-sum game, although the constant $c$ may be positive or negative, depending on the parameters. When $I_s$, the interaction coefficient identified by Bever et al., is negative, $c$ is positive, ensuring (neutral) stability. This condition amounts to an inequality constraint, rather than an equality constraint, and so it \emph{is} generally robust to small variations in model parameters (Fig ???). As we can now see, the case $n = 2$ is unique in this regard.

\section{Discussion}

The Bever model has played a central role in motivating PSF research, and continues to guide both theory and experiment in the fast-growing field \cite{bever2015maintenance,kandlikar2019winning,ke2020effects}. In this Letter, we extend the Bever model to any number of species, and highlight its equivalence to bimatrix game dynamics. Taking advantage of the well-developed theory for these dynamics, we are able to characterize the behavior of the generalized Bever model in detail.

Our central finding is that there can be no robust coexistence of plant species in this model. Regardless of the number of species, $n$, the model never exhibits equilibrium coexistence or other attractors. Coexistence can be attained through neutral oscillations, but these dynamics lack any restoring force and are easily destabilized by stochasticity or exogenous forcing. In this respect, the generalized model behaves similarly to the now-classic two-species system. However, unlike the two-species model, oscillations with $n > 2$ species can only result under very restricted parameter combinations. These parameterizations are vanishingly unlikely to arise by chance, and highly sensitive to small deviations. Thus, coexistence of more than two species is neither dynamically nor structurally stable.

This result may be surprising, because a significant body of experimental evidence indicates that PSFs play a key role in mediating the coexistence of more than two species in natural communities \cite{kulmatiski2008plant,petermann2008janzen,mangan2010negative,bever2015maintenance}. Apparently, the picture suggested by the two-species Bever model generalizes in nature, but not in the model framework itself. We note that this framework was introduced as an intentional simplification to illustrate the potential role of PSFs in mediating coexistence, not to accurately model the biological details of PSFs. In this capacity, the model has been wildly successful. And indeed, alongside empirical study of PSFs, other modeling approaches have emerged, accounting for more biological realism (e.g. \cite{umbanhowar2005simple,eppstein2007invasiveness,bever2010rooting}), or with the demonstrated capacity to produce multispecies coexistence (e.g. \cite{bonanomi2005negative,miller2021metapopulations}). Some of these are minor modifications of the Bever model framework; others build on distinct foundations \cite{ke2015incorporating,ke2020effects}. Our results suggest that these various avenues are worth pursuing further.

Alternative modeling approaches are particularly important for better integration of theory and data. The predictions of the Bever model are commonly used to guide the design and analysis of PSF experiments, especially in drawing conclusions about coexistence. Our analysis cautions that applications of this model in multispecies communities might lead to incorrect inference. For example, attempts to parameterize the Bever model for three species using empirical data have produced predictions of non-coexistence in plant communities that seem to coexist experimentally \cite{kulmatiski2011testing}. In many other studies, the pairwise interaction coefficient, $I_s$, is calculated for species pairs and used to assess whole-community coexistence \cite{kulmatiski2008plant,fitzsimons2010importance,pendergast2013belowground,suding2013consequences,kuebbing2015plant,smith2015plant,bauer2017effects,kandlikar2019winning}. However, we have seen that whole-community coexistence is virtually impossible within the generalized model, and there is no guarantee that the pairwise coexistence conditions for this model will agree with $n$-species coexistence conditions in other frameworks. For example, $I_s < 0$ for all species pairs is neither necessary nor sufficient to produce coexistence in a metapopulation-based model for PSFs \cite{miller2021metapopulations}.

In contrast to other extensions of the Bever model to many-species communities \cite{eppinga2018frequency,mack2019plant}, our approach keeps the dynamics of both plants and soil fully explicit. As such, we make no assumptions about the relative timescales of plant and soil dynamics, or whether either of these reach equilibrium. This difference likely explains the discrepancy between our conclusions and previously published predictions for $n$-species PSFs \cite{kulmatiski2011testing,eppinga2018frequency,mack2019plant}.

PSFs as envisioned in the classic Bever model might facilitate coexistence in conjunction with other mechanisms, such as direct species interactions, or through long transient dynamics, but our analysis shows that they cannot produce robust $n$-species coexistence in isolation. This finding calls for renewed theoretical investigation of PSFs. One important consideration is grounding PSF modeling frameworks in more realistic models for absolute abundances or densities. As various researchers have noted \cite{revilla2013plant,eppinga2018frequency,ke2020effects}, the Bever model and its extensions are in fact projections (onto the space of relative abundances or frequencies) of dynamics for plant and soil \emph{abundances}. Consequently, the projected dynamics can mask unbiological outcomes in the original model (e.g. relative abundances oscillate around equilibrium while absolute abundances shrink to zero or explode to infinity). Indeed, the absolute abundance model (Eq.~\ref{nspp_absolute}) used to derive our $n$-species frequency dynamics (Eqs.~\ref{nspp_relative}-\ref{matrix_form}) does not generally possess any fixed points, which is a basic requirement for species coexistence \cite{hutson1990existence,hutson1992permanence}. It is usually seen as desirable to study PSFs in the space of species frequencies, both because this facilitates connections to data, and because frequencies are considered a more appropriate metric for analyzing processes that stabilize coexistence (\cite{adler2007niche,eppinga2018frequency} but see \cite{kandlikar2019winning,ke2020effects}). But models that introduce frequencies through a natural constraint, such as competition for finite space, will likely produce more realistic dynamics.

From a broader theoretical perspective, the qualitative change in model behavior that we observe as the number of species increases from two to three or more is a striking phenomenon, but not an unprecedented one. Ecologists have repeatedly found that intuitions from two-species models can generalize (or fail to generalize) to more diverse communities in surprising ways \cite{strobeck1973n,smale1976differential,barabas2016effect}. Our analysis provides another illustration of the fact that ``more is different'' \cite{anderson1972more} in ecology, and highlights the importance of developing theory for species-rich communities.

\section*{Acknowledgments}
We thank ...

\bibliographystyle{unsrt}
\bibliography{references}

\section*{Supplemental Methods}

\subsection{Model derivation}

As described in the Main Text, we begin with the system

\begin{align}
\begin{cases}
\frac{dx_i}{dt} &= x_i \left(\sum_{j} \alpha_{ij} \, q_j \right) \, , \quad  i = 1, \dots n \\
\frac{dy_i}{dt} &= y_i \left( \beta_i \, p_i \right)
\end{cases}
\end{align}
governing the time-evolution of plant abundances $x_i$ and soil components $y_i$, where $p_i = x_i / \sum_j x_j$, $q_i = y_i / \sum_j y_j$, and Greek letters denote nonnegative parameters. These equations capture the assumptions outlined by Bever et al. \cite{bever1997incorporating} for two species and extend them straightforwardly to any $n$ species. Following the approach of Bever et al. for two species (and consistent with other generalizations of this model \cite{kulmatiski2008plant,eppinga2018frequency}), we derive dynamics for the $p_i$ by applying the chain rule:

\begin{align} \label{derive_relative}
\begin{split}
	\frac{dp_i}{dt} &= \frac{d}{dt} \frac{x_i}{\sum x_j}\\  
	&= \frac{1}{\sum_j x_j} \frac{dx_i}{dt} - \frac{x_i}{(\sum_j x_j)^2} \sum_j \frac{dx_j}{dt} \\
	&= \frac{x_i}{\sum_j x_j} \left( \sum_{j} \alpha_{ij} \, q_j \right) - \frac{x_i}{\sum_j x_j} \left(\sum_j \frac{x_j}{\sum_k x_k} \sum_{l} \alpha_{jl} \, q_l \right)\\
	&= p_i \left(\sum_{j} \alpha_{ij} \, q_j -  \sum_{j, k} \alpha_{jk} p_j  q_k \right) \, .
\end{split}
\end{align}
This last expression is identical to the first line of Eq.~\ref{nspp_relative} in the Main Text. The dynamics for $q_i$ can be derived in exactly the same way (using the definitions $\beta_{ii} = \beta_i$ and $\beta_{ij} = 0$ to make the parallel clear). The two terms of each per capita growth rate in Eq.~\ref{nspp_relative} have natural interpretations in the language and notation of linear algebra: $\sum_{j} \alpha_{ij} \, q_j$ is the $i$th component of the matrix-vector product $A \bm{q}$ and $\sum_{j, k} \alpha_{jk} p_j  q_k$ is the bilinear form $\bm{p}^T A \bm{q}$. Here, $A$ (and $B$) is an $n \times n$ matrix and $\bm{p}$ ($\bm{q}$) is a vector of length $n$, as described in the Main Text. We can re-write Eq.~\ref{nspp_relative} as
\begin{align}
\begin{split}
	\frac{dp_i}{dt} &= p_i \left((A \bm{q})_i - \bm{p}^T A \bm{q} \right) \\
	\frac{dq_i}{dt} &= q_i \left((B \bm{p})_i - \bm{q}^T B \bm{p} \right) \\
\end{split}
\end{align}
or even more compactly as
\begin{align}
\begin{cases}
\frac{d\bm{p}}{dt} &= D(\bm{p}) \left(A \bm{q} - (\bm{p}^T A \bm{q}) \bm{1} \right) \\
\frac{d\bm{q}}{dt} &= D(\bm{q}) \left(B \bm{p} - (\bm{q}^T B \bm{p}) \bm{1}  \right) \,
\end{cases}
\end{align}
which is Eq.~\ref{matrix_form} in the Main Text.

These expressions are identical to standard bimatrix replicator dynamics \cite{hofbauer1996evolutionary,hofbauer1998evolutionary}. Bimatrix games have two strategy sets (here, the $p_i$ and $q_i$), and interactions take place only between strategies from opposite sets. The growth rate terms we considered above now have interpretations as payoffs or fitnesses: $\sum_{j} \alpha_{ij} \, q_j = (A \bm{q})_i$ is the payoff for strategy $i$ (an average of payoffs playing against each strategy of the other ``player'', weighted by the frequency of each strategy, $q_j$) and $\sum_{j, k} \alpha_{jk} p_j  q_k = \bm{p}^T A \bm{q}$ is the average payoff across the population of strategies. A general bimatrix game may have any nonnegative $B$; our model assumptions lead to the special case where $B$ is diagonal. We note that one could easily and plausibly consider an extension of the Bever model where each plant species has some affect on (up to) all $n$ of the soil components. Then, our PSF model would be map exactly onto the full space of bimatrix game dynamics (rather than just a subset). However, all of the results we consider hold for arbitrary bimatrix games, meaning the same conclusions about the dynamics of Eqs.~\ref{nspp_relative}-\ref{matrix_form} would apply to this extended model, as well.

We note two useful properties of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, as they will be important for the analysis that follows. First, we have the constraint $\sum_i p_i = \sum_i q_i = 1$ at every point in time. Second, the dynamics are completely unchanged by adding a constant to any \emph{column} of the parameter matrices $A$ or $B$. The first fact is a direct consequence of our definition for $p_i$ and $q_i$; the second can easily be shown. Suppose we have added a constant $c$ to each element in the $l$th column of $A$. Then

\begin{align}
\begin{split}
\frac{dp_i}{dt} &= p_i \left(\sum_{j} \alpha_{ij} \, q_j + c \, q_l - \sum_{j, k} \alpha_{jk} \, p_j \, q_k - \sum_{j} c \, p_j \, q_l \right)\\
&= p_i \left(\sum_{j} \alpha_{ij} \, q_j + c \, q_l - \sum_{j, k} \alpha_{jk} \, p_j \, q_k - c \, q_l \right) \\
&= p_i \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right)
\end{split}
\end{align}
which is precisely the differential equation we obtained prior to adding $c$. Clearly the trajectories of both systems (with and without the column shift) must be identical. The same considerations apply for the matrix $B$. Intuitively, this property reflects the fact that we are always subtracting the average payoff, and so any change to the payoffs that benefits (or harms) each species equally is ``invisible'' to the dynamics.

In the remaining sections, we outline the main behaviors of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, especially with regard to coexistence. We closely follow the treatment by Hofbauer and Sigmund \cite{hofbauer1998evolutionary}, and urge interested readers to consult this excellent introduction (see especially chapters 10 and 11). Here, we reproduce or sketch the essential details needed to justify the results in the Main Text.

\subsection{Coexistence equilibrium}

Written in matrix form, it is easy to see that the model admits a unique fixed point where all species are present at non-zero frequency. This fixed point, $(\bm{p}^\star, \bm{q}^\star)$, must take the form $(c_1 B^{-1} \bm{1}, c_2 A^{-1} \bm{1})$ for some undetermined constants $c_1$ and $c_2$. Substituting this ansatz into the growth rates in Eq.~\ref{matrix_form} and equating them to zero, we have

\begin{align}
	\begin{split}
	A \bm{q}^\star - ((\bm{p}^\star)^T A \bm{q}^\star) \bm{1} = c_2 A A^{-1} \bm{1} - (c_1 c_2 \bm{1}^T  (B^{-1})^T A A^{-1} \bm{1}) \bm{1} &= c_2 (1 - c_1 \bm{1}^T  (B^{-1})^T \bm{1}) \bm{1} = 0 \\
	B \bm{p}^\star - ((\bm{q}^\star)^T B \bm{p}^\star) \bm{1} = c_1 B B^{-1} \bm{1} - (c_1 c_2 \bm{1}^T  (A^{-1})^T B B^{-1} \bm{1}) \bm{1} &= c_1 (1 - c_2 \bm{1}^T  (A^{-1})^T \bm{1}) \bm{1} = 0
	\end{split}
\end{align}
From the final two equations, it is clear that $c_1 = \frac{1}{\bm{1}^T  (B^{-1})^T \bm{1}} = \frac{1}{\bm{1}^T  B^{-1} \bm{1}}$ and $c_2 = \frac{1}{\bm{1}^T  (A^{-1})^T \bm{1}} = \frac{1}{\bm{1}^T  A^{-1} \bm{1}}$.

These rescaling factors make intuitive sense, as they ensure that $\sum_i p^\star_i = \sum_i q^\star_i = 1$, consistent with their definition as frequencies.

Describing these equilibrium frequencies in terms of the parameters is a difficult problem that has received significant attention elsewhere \cite{eppinga2018frequency,mack2019plant,saavedra2017structural,servan2018coexistence,pettersson2020stability,saavedra2021feasibility}. In particular, one is usually interested in identifying whether all of the frequencies are nonnegative (such a fixed point is said to be feasible). The existence of a feasible fixed point is a requirement for the model to exhibit permanence, meaning that no species go extinct or grow to infinity. Throughout our analysis, we assume the existence of a feasible fixed point; considering the question of feasibility simultaneously would only make coexistence less likely in each case.

\subsection{Local stability analysis}

Perturbations around the coexistence equilibrium are constrained to respect the conditions $\sum_i p_i = \sum_i q_i = 1$. For this reason, it is convenient to remove these constraints before performing a local stability analysis. As in the two species case \cite{bever1997incorporating}, this can be done by eliminating the $n$th species and soil component, which leaves us with a $2n - 2$ dimensional system with no special constraints.

We use $p_n = 1 - \sum_{i = 1}^{n-1} p_i \equiv f(\bm{p})$ and $q_n = 1 - \sum_{i=1}^{n-1} q_i \equiv g(\bm{q})$ and write these frequencies as functions of the others. The reduced dynamics are given by

\begin{align}
\begin{cases} \label{reduced}
\frac{dp_i}{dt} &= p_i \left(\sum_{j}^{n-1} \alpha_{ij} \, q_j + \alpha_{in} g(\bm{q}) - \sum_{j, k}^{n-1} \alpha_{jk} \, p_j \, q_k - f(\bm{p}) \sum_{j}^{n-1} \alpha_{nj} q_j - g(\bm{q}) \sum_{j}^{n-1} \alpha_{jn} p_j - \alpha_{nn} f(\bm{p}) g(\bm{q})  \right) \\
\frac{dq_i}{dt} &= q_i \left(\beta_{i} \, p_i - \sum_{j}^{n-1} \beta_{j} \, p_j \, q_j  - \beta_n f(\bm{p}) g(\bm{q})  \right) \, , \quad  i = 1, \dots n-1
\end{cases}
\end{align}
Although these equations appear more complex, it is now straightforward to analyze the local stability of the coexistence equilibrium.

The elements of the community matrix (the Jacobian evaluated at the coexistence equilibrium) are easily computed from Eq.~\ref{reduced}. First we consider the plant dynamics differentiated with respect to the plant frequencies. In these calculations, all frequencies are evaluated at their equilibrium values.

\begin{align} \label{firstblock}
\begin{split} 
	\frac{\partial}{\partial p_j} \frac{dp_i}{dt} &= p_i \left( - \sum_{k}^{n-1} \alpha_{jk} \, q_k + \sum_{k}^{n-1} \alpha_{nk} q_k - \alpha_{jn} g(\bm{q}) + \alpha_{nn} g(\bm{q}) \right) \\
	&= p_i \left( - \sum_{k}^{n} \alpha_{jk} \, q_k + \sum_{k}^{n} \alpha_{nk} q_k \right)\\
	&= 0
\end{split}
\end{align}
Here, we have used the fact that $A \bm{q^*} \propto \bm{1}$. Notice that, because the factors in parentheses in Eq.~\ref{reduced} are zero at equilibrium, these community matrix calculations are valid even for $i = j$.

The other elements are computed similarly:

\begin{align}
\begin{split} 
\frac{\partial}{\partial q_j} \frac{dq_i}{dt} &= q_i \left( - \beta_i q_i + \beta_n f(\bm{p}) \right) \\
&= 0
\end{split}
\end{align}

\begin{align}
\begin{split} 
\frac{\partial}{\partial q_j} \frac{dp_i}{dt} &= p_i \left(\alpha_{ij} - \alpha_{in} - \sum_{k}^{n-1} \alpha_{kj} \, p_k - \alpha_{nj} \, f(\bm{p}) + \sum_{k}^{n-1} \alpha_{kn} p_k + \alpha_{nn} f(\bm{p}) \right) \\
&= p_i(\alpha_{ij} - \alpha_{in})
\end{split}
\end{align}

\begin{align} \label{lastblock}
\begin{split} 
\frac{\partial}{\partial p_j} \frac{dq_i}{dt} &= 
\begin{cases}
q_i \left(\beta_{i} - \beta_{n} \right) \, , \quad i = j\\
0 \, , \quad i \neq j
\end{cases} \\
\end{split}
\end{align}

From these calculations, it is apparent that the trace of the community matrix, given by $\sum_i^{n-1} \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_j^{n-1} \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$ , is zero. The trace of a square matrix is equal to the sum of its eigenvalues \cite{horn2012matrix}, so the eigenvalues of the community matrix must include either (i) a mix of positive and negative real parts or (ii) only purely imaginary values. In the first case, the coexistence equilibrium is locally unstable, because at least one eigenvalue has positive real part. In the second case, the coexistence equilibrium is a neutrally or marginally stable. These two possibilities exclude locally stable equilibria. In this respect, the behavior of the two-species model is the generic behavior of the generalized $n$-species model.

\subsection{Divergence and attractors}

We can extend this picture beyond a local neighborhood of the coexistence equilibrium by considering the divergence of the vector field associated with Eqs.~\ref{nspp_relative}-\ref{matrix_form}. The divergence, defined by $\sum_i \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_i \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$, measures the outgoing flux around a given point. It can be shown (see \cite{eshel1983coevolutionary} and \cite{hofbauer1998evolutionary}) that up to a change in velocity (i.e. rescaling time by a positive factor), the vector field corresponding to any bimatrix game dynamics has zero divergence everywhere in the interior of the positive orthant (i.e. where $p_i, q_i > 0$ for all $i$).

The divergence theorem \cite{arfken1985mathematical} equates the integral of the divergence of a vector field over some $n$-dimensional region to the net flux over the boundary of the region. For a vector field with zero divergence, this implies that every closed surface has zero net flux. As a consequence, such \emph{divergence-free} vector fields cannot have attractors, or subsets of phase space toward which trajectories of the corresponding dynamical system tend to evolve. If an attractor existed, one could define a surface enclosing it sufficiently tightly, and the net flux over this surface would be negative (as trajectories enter, but do not exit, this region). But this would present a contradiction, and so we conclude that there can be no attractors, such as limit cycles, for the dynamics.

For our model, these facts mean that attractors can only exist on the boundary of the phase space. Because each boundary face for the $n$-dimensional system is another bimatrix replicator system on $n-2$ dimensions, the same logic applies, and the only possible attractors are points where a single species (and corresponding soil component) is present \cite{hofbauer1998evolutionary}. States with multiple species present are never attractive. This leaves neutrally-stable oscillations as the only potential form of species coexistence.

\subsection{Rescaled zero-sum games are neutrally stable}

In the context of bimatrix games, a zero-sum game is one where $A = -B^T$. A rescaled zero-sum game is one where there exist constants $\gamma_i, \delta_j$ and $c > 0$ such that $a_{ij} + \delta_j = -c b_{ji} + \gamma_i$ for all $i$ and $j$ (here, we understand $A = (a_{ij}), B = (b_{ij})$) \cite{hofbauer1998evolutionary}. Any rescaled zero-sum game can be turned into a zero-sum game by adding constants (in particular, $-\delta_j$ and $-\gamma_j$) to each column of $A$ and $B$, and then multiplying $B$ by a positive constant $1 / c$. As such, the dynamics of a rescaled zero-sum game and its corresponding zero-sum game are the same up to a rescaling of time.

If a rescaled zero-sum game has a feasible coexistence equilibrium, this equilibrium is neutrally stable. We can see this by considering the associated community matrix. First, we assume without loss of generality that $A = -c B^T$ (otherwise, we shift columns to obtain this form, without altering the dynamics in the process) Now we add the column-constant matrix $\frac{1}{c} \bm{b}_n \bm{1}^T$ to $A$ and $c \bm{a}_n \bm{1}^T$ to $B$, where $\bm{a}_n$ ($\bm{b}_n$) denotes the $n$th column of $A$ ($B$). Again, the dynamics, including both equilibrium values and stability properties, are unchanged by this operation. From Eqs.~\ref{firstblock}-\ref{lastblock}, we see that the community matrix, $J$, of the resulting system is given by

\begin{equation}
\begin{pmatrix}
	0 && D(\bm{p}^\star) (\bar{A} + \frac{1}{c} \bm{b}_n \bm{1}^T - \bm{1} \bm{a}_n^T) \\
	D(\bm{q}^\star) (\bar{B} + c \bm{a}_n \bm{1}^T - \bm{1} \bm{b}_n^T) && 0
\end{pmatrix}
\end{equation}
where $\bar{A}$ ($\bar{B}$) denotes the $(n-1) \times (n-1)$ submatrix of $A$ ($B$) obtained by dropping the $n$th row and column. Finally, we consider the similarity transform $P^{-1} J P$, defined by the change of basis matrix

\begin{equation}
P = \begin{pmatrix}
	\sqrt{c} D(\bm{p}^\star)^{1/2} && 0 \\
	0 && D(\bm{q}^\star)^{1/2}
\end{pmatrix} \, .
\end{equation}
The resulting matrix, $J'$, which shares the same eigenvalues as $J$ \cite{horn2012matrix}, is given by

\begin{equation}
\begin{pmatrix}
0 && \sqrt{c} D(\bm{p}^\star)^{1/2} (\bar{A} + \frac{1}{c} \bm{b}_n \bm{1}^T - \bm{1} \bm{a}_n^T) D(\bm{q}^\star)^{1/2} \\
\sqrt{c} D(\bm{q}^\star)^{1/2} (-\bar{A^T} + \bm{a}_n \bm{1}^T - \frac{1}{c} \bm{1} \bm{b}_n^T) D(\bm{p}^\star)^{1/2} 
\end{pmatrix}
\end{equation}
which is a skew-symmetric matrix. Every eigenvalue of a skew-symmetric matrix must have zero real part \cite{horn2012matrix}. Thus, the eigenvalues of $J$, the community matrix, have zero real part, and the coexistence equilibrium of our original system is neutrally stable.

Here, we have outlined a proof that applies to all rescaled zero-sum games. When $B$ is a diagonal matrix, as in our model of PSFs, the condition for $A$ and $B$ to constitute a rescaled zero-sum game reduces to the condition given in the Main Text.

Rescaled zero-sum games are the only bimatrix games known to produce neutrally stable oscillations. It is a long-standing conjecture that no other bimatrix games have this property \cite{hofbauer1996evolutionary,hofbauer1998evolutionary,hofbauer2011deterministic}.

\subsection{Two-species bimatrix games}

For $n > 2$, the rescaled zero-sum game condition is very stringent -- it places exacting equality constraints on the elements of $A$ and $B$. However, for $n = 2$, every bimatrix game satisfies $a_{ij} + \delta_j = -c b_{ji} + \gamma_i$ for some $c$ potentially positive (in which case we have a rescaled zero-sum game) or negative (in which case the game is called a \emph{partnership game}, and the coexistence equilibrium is unstable) \cite{hofbauer1998evolutionary}. Thus, neutral oscillations arise whenever $c > 0$.

To see that this is true, we first suppose that $A$ and $B$ have the form

\begin{equation}
	A = \begin{pmatrix}
	0 && a_1 \\
	a_2 && 0
	\end{pmatrix}
	\quad \quad 
	B = \begin{pmatrix}
	0 && b_1 \\
	b_2 && 0 
	\end{pmatrix} \, . 
\end{equation}
If this is not the case, we can use constant column shifts to arrive at this form (e.g., in general, $a_1 = a_{12} - a_{22}$). Now consider the constants $c = -\frac{a_1 + a_2}{b_1 + b_2}$ and $\gamma_1 = \delta_1 = a_1 + c b_2$ and $\gamma_2 = \delta_2 = 0$. Examining the equation $a_{ij} + \delta_j - \gamma_i = -c b_{ji}$ for each $i$ and $j$, one verifies

\begin{align}
	\begin{split}
	0 + \gamma_1 - \delta_1  &= 0 \\
	a_1 + \gamma_2 - \delta_1 &= -c b_2 \\
	a_2 + \gamma_1 - \delta_2 = -c (b_1 + b_2 - b2) &= -c b_1 \\
	0  &= 0
	\end{split}
\end{align}
and so the parameters $A$ and $B$ always constitute a rescaled zero-sum or partnership game. In the particular case of our model, $a_1 + a_2 = -\alpha_{11} + \alpha_{21} + \alpha_{12} - \alpha_{22} = -I_s$ and $b_1 + b_2 = -\beta_1 - \beta_2$. $c$ is positive (as needed for cycles) when these signs disagree; since $b_1 + b_2 = -\beta_1 - \beta_2$ is always negative, $a_1 + a_2$ must be positive, meaning $I_s < 0$, as found by Bever et al.

\subsection{Constant of motion}

When $A$ and $B$ satisfy the rescaled zero-sum game condition, the function

\begin{equation}
	H(\bm{p}, \bm{q}) = \sum_i p_i^\star \log p_i + c \sum_j q_j^\star \log q_j 
\end{equation}
is a constant of motion for the dynamics \cite{hofbauer1998evolutionary}. As above, we suppose that $A = -c B^T$, and shift the columns of each matrix as needed if this is not the case. Then consider the time derivative

\begin{align}
	\begin{split}
	\frac{dH}{dt} &= \sum_i p_i^\star \frac{1}{p_i} \frac{dp_i}{dt} + c \sum_j q_j^\star \frac{1}{q_i} \frac{dq_i}{dt} \\
	&= \sum_i p_i^\star \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right) + c \sum_j q_j^\star \left( \beta_{i} \, p_i - \sum_{j} \beta_{j} \, p_j \, q_j \right) \\
	&= \sum_{i, j} \alpha_{ij} \, p_i^\star \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k + c \sum_{i} \beta_{i} \, q_i^\star \, p_i - c \sum_{j} \beta_{j} \, p_j \, q_j \\
	&= \sum_{i, j} \alpha_{ij} \, (p_i^\star - p_i) \, q_j + c \sum_{i} \beta_{i} \, (q_i^\star - q_i) \, p_i 
	\intertext{Now, because $A = -c B^T$, we have}
	&= c \sum_i \beta_i \left( -(p_i^\star - p_i) \, q_i + (q_i^\star - q_i) \, p_i \right)\\
	&= c \sum_i \beta_i \left( -p_i^\star \, q_i + q_i^\star \, p_i \right)
	\intertext{and because $q_i^* = p_i^* = \frac{Z}{\beta_{i}}$, with $Z$ the normalizing constant,}
	&= c \, Z \, \sum_i (-q_i + p_i) \\
	&= 0
	\end{split}
\end{align}
In the last line, we use the fact that both sets of frequencies always sum to one.

Each orbit remains in the level set defined by the initial conditions,$(\bm{p}_0, \bm{q}_0)$:

\begin{equation}
H(\bm{p}_0, \bm{q}_0) = \sum_i p_i^\star \log p_i + c \sum_j q_j^\star \log q_j 
\end{equation}
For the two-species model studied by Bever et al., these level sets precisely define the trajectories in the $(p, q)$ phase plane.

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-10-23 code/functions.py
started storing solutions of the search

@@ -24,7 +24,10 @@ def integrate_PSF(fun, t_span, z0, args):
     Wrapper for integrator
     '''
     #Solve diferential equations
-    sol = solve_ivp(model, t_span, z0, method = 'BDF', args = args)
+    try:
+        sol = solve_ivp(model, t_span, z0, method = 'BDF', args = args)
+    except:
+        import ipdb; ipdb.set_trace(context = 20)
     #Get number of species 
     n = args[-1]
     #Get plant and soil abundances

### 2021-10-23 code/perturbation.py

@@ -21,10 +21,11 @@ import pandas as pd
 def main(argv):
     '''Main function'''
     #Number of species
-    n = 6
+    n = 4
+    #Set extinct threshold
+    tol = 1e-9
     like = 0
     n_it_max = 20
-    n_it = 0
     #Perturbation strenght
     p = 0.01
     #Set to False convergence flags
@@ -33,6 +34,16 @@ def main(argv):
     #Set number of simulations
     n_sim = 2
     sim_it = 0
+    #Initialize emtpy lists containin variables of interest.
+    sim_vec = []
+    t_vec = []
+    t_per_vec = []
+    p_l = []
+    p_ab = []
+    p_ab_per = []
+    s_l = []
+    s_ab = []
+    s_ab_per = []
     while sim_it < n_sim:
         #Keep sampling and integrating until convergence
         while not conv or not conv_per:
@@ -40,8 +51,7 @@ def main(argv):
             A, B = sampling_matrices(n)
             A_per = A + np.random.uniform(1-p, 1+p, size = np.shape(A))
             #Integrate unperurbed system
-            plant_ab, soil_ab, t = integrate_PSF(model, [1, 2000], 
-                                                 2*n*[1/n], 
+            plant_ab, soil_ab, t = integrate_PSF(model, [1, 2000], 2*n*[1/n], 
                                                  (A, B, n))
             #Check convergence of unperurbed system
             conv = check_convergence(plant_ab, soil_ab)
@@ -51,24 +61,85 @@ def main(argv):
                                                              (A_per, B, n))
             #Check convergence of perurbed system
             conv_per = check_convergence(plant_ab_per, soil_ab_per)
+        #Remove extinct species from perturbed system
+        rem_plant, rem_soil = remove_extinctions(plant_ab_per[:, -1], 
+                                                 soil_ab_per[:, -1], 
+                                                 tol) 
         #Check if the convergent perturbed system has arrived to equilibrium
-        equilibrium = check_equilibrium(plan_ab_per[:, -1], 
-                                        soil_ab_per[:, -1])
+        equilibrium = check_equilibrium(rem_plant, 
+                                        rem_soil)
+        #Set number of integration cycles to zero
+        n_int = 0
         #Keep integrating both systems until equilibrium is reached
         while not equilibrium:
-            #Organize new initial conditions and re-integrate, evaluating
-            #equilibrium at the end of each simulation
+            #Remove extinct species from payoff matrices
[…]

### 2021-10-23 code/plant_feedback_model.py

@@ -116,9 +116,13 @@ def main(argv):
             if not convergence:
                 #If it did, flag as non-convergent integration
                 df.loc[n_act + n_sim*n_vec_it, 'n_p_f'] = -1
+                #Don't count this simulation
+                n_act -= 1
             elif n_int > n_int_max:
                 #couln't trim the system down to two species
                 df.loc[n_act + n_sim*n_vec_it, 'n_p_f'] = -2
+                #Don't count this simulation
+                n_act -= 1
             else:
                 df.loc[n_act + n_sim*n_vec_it, 'n_p_f'] = len(rem_plant)
             n_act += 1

### 2021-10-25 code/perturbation.py
first working version of systematic search

@@ -25,7 +25,7 @@ def main(argv):
     #Set extinct threshold
     tol = 1e-9
     like = 0
-    n_it_max = 20
+    n_int_max = 20
     #Perturbation strenght
     p = 0.01
     #Set to False convergence flags
@@ -44,7 +44,9 @@ def main(argv):
     s_l = []
     s_ab = []
     s_ab_per = []
+    cyc_vec = []
     while sim_it < n_sim:
+        print('\nSimulation number: ', sim_it)
         #Keep sampling and integrating until convergence
         while not conv or not conv_per:
             #Sample matrices A and B with H-S constraints
@@ -54,13 +56,14 @@ def main(argv):
             plant_ab, soil_ab, t = integrate_PSF(model, [1, 2000], 2*n*[1/n], 
                                                  (A, B, n))
             #Check convergence of unperurbed system
-            conv = check_convergence(plant_ab, soil_ab)
+            conv = check_convergence(plant_ab[:, -1], soil_ab[:, -1], tol)
             #Integrate perurbed system
             plant_ab_per, soil_ab_per, t_per = integrate_PSF(model, [1, 2000], 
                                                              2*n*[1/n], 
                                                              (A_per, B, n))
             #Check convergence of perurbed system
-            conv_per = check_convergence(plant_ab_per, soil_ab_per)
+            conv_per = check_convergence(plant_ab_per[:, -1], 
+                                         soil_ab_per[:, -1], tol)
         #Remove extinct species from perturbed system
         rem_plant, rem_soil = remove_extinctions(plant_ab_per[:, -1], 
                                                  soil_ab_per[:, -1], 
@@ -70,8 +73,13 @@ def main(argv):
                                         rem_soil)
         #Set number of integration cycles to zero
         n_int = 0
-        #Keep integrating both systems until equilibrium is reached
+        #Initialize list to keep track of extinct plants and soils
+        ext_ind = []
+        #Keep integrating both systems until equilibrium in perturbed one 
+        #is reached
         while not equilibrium:
+            #Update number of plants
+            n_plant = len(rem_plant)
             #Remove extinct species from payoff matrices
             ext_plant, ext_soil = find_extinct_indices(plant_ab_per[:, -1], 
                                                        rem_plant, 
@@ -79,6 +87,8 @@ def main(argv):
                                                        rem_soil)
             A_per = remove_extinctions_matrix(A_per, ext_plant)
[…]

### 2021-10-26 code/dynamics_collapse.R
first draft of figure 2

@@ -1,0 +1,116 @@
+library(tidyverse)
+library(tidyquant)
+library(reshape2)
+library(cowplot)
+library(gridExtra)
+
+
+data = read_csv('../data/zach_results.csv')
+dev.off()
+p1 = ggplot(data = data,
+       aes(x = t, y = p_ab,
+           color = as.factor(p_l)))+
+  geom_line(alpha = 0.2)+
+  geom_ma(ma_fun = EMA, n = 150, linetype = 'solid')+
+theme(aspect.ratio = 3/4,
+            panel.background = element_blank(),
+            panel.border = element_rect(fill = NA),
+            legend.position = 'none',
+            axis.title.x = element_text(size = 15),
+            axis.title.y = element_text(size = 15),
+            axis.text = element_blank(),
+            axis.ticks = element_blank())+
+  labs(y = "Abundance",
+       x = "Time")
+print(p1)
+
+p2 = ggplot(data = data,
+            aes(x = t, y = p_ab_per,
+                color = as.factor(p_l)))+
+  geom_line(alpha = 0.2)+
+  geom_ma(ma_fun = EMA, n = 150, linetype = 'solid')+
+  theme(aspect.ratio = 3/4,
+        panel.background = element_blank(),
+        panel.border = element_rect(fill = NA),
+        legend.position = 'none',
+        axis.title.x = element_text(size = 15),
+        axis.title.y = element_text(size = 15),
+        axis.text = element_blank(),
+        axis.ticks = element_blank())+
+  labs(y = "",
+       x = "Time")
+print(p2)
+
+#Load matrices
+load('../data/example_perturbation_parameters.RData')
+
+A_pre = my_list$A_pre
+A_pre_melt = melt(A_pre)
+
+A_pre_plot =  ggplot(data = A_pre_melt, 
+                     aes(x = Var2, y = rev(Var1), fill = value)) +
+  geom_tile()+
+  theme(panel.background = element_rect(fill = NA),
+        plot.background = element_blank(),
+        legend.position = 'none',
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.title = element_blank(),
+        aspect.ratio = 1)+
+    labs(x = '', y = '')
+
+print(A_pre_plot)
+
+A_post = my_list$A_post
+A_post_melt = melt(A_post)
+
[…]

### 2021-10-26 code/functions.py

@@ -13,21 +13,22 @@ def model(t, z, A, B, n):
     Ip = np.ones(shape = n).reshape(n, 1)
     Iq = np.ones(shape = n).reshape(n, 1)
     #Model equations
-    dpdt = np.diag(p.transpose()[0]) @ (A @ q - \
-            (p.transpose()[0] @ A @ q) * Iq) 
-    dqdt = np.diag(q.transpose()[0]) @ (B @ p - \
-            (q.transpose()[0] @ B @ p) * Ip)
+    try:
+        dpdt = np.diag(p.transpose()[0]) @ (A @ q - \
+                (p.transpose()[0] @ A @ q) * Iq) 
+        dqdt = np.diag(q.transpose()[0]) @ (B @ p - \
+                (q.transpose()[0] @ B @ p) * Ip)
+    except:
+        import ipdb; ipdb.set_trace(context = 20)
     return(list(dpdt.reshape(n)) + list(dqdt.reshape(n)))
 
-def integrate_PSF(fun, t_span, z0, args):
+def integrate_PSF(fun, t_span, z0, args, method = 'BDF'):
     '''
     Wrapper for integrator
     '''
     #Solve diferential equations
-    try:
-        sol = solve_ivp(model, t_span, z0, method = 'BDF', args = args)
-    except:
-        import ipdb; ipdb.set_trace(context = 20)
+    sol = solve_ivp(model, t_span, z0, method = method, args = args)
+
     #Get number of species 
     n = args[-1]
     #Get plant and soil abundances
@@ -136,6 +137,22 @@ def check_equilibrium(plant_vector, soil_vector):
         equilibrium = True
     return equilibrium
 
+def check_equilibrium_bis(plant_vector, soil_vector):
+    '''
+    Check if only 2 species are greater than the tolerance
+    '''
+    n_plants = sum(plant_vector > 0)
+    n_soils = sum(soil_vector > 0)
+    if n_plants > 2 or n_soils > 2:
+        #If either plant/soil have more than 2, we have not reached
+        #equilibnrium
+        equilibrium = False
+    else:
+        #Otherwise (neither have more than 2), declare equilibrium
+        equilibrium = True
+    return equilibrium
+
+
 def check_convergence(plants, soils, tol):
     '''
     Check that integration has converged by verifying that all abundances at
@@ -163,4 +180,18 @@ def find_extinct_indices(plants, plants_rem, soils, soils_rem):
     o, ext_soil_ind, o = np.intersect1d(soils, ext_soil, 
[…]

### 2021-10-26 code/perturbation.py

@@ -27,7 +27,7 @@ def main(argv):
     like = 0
     n_int_max = 20
     #Perturbation strenght
-    p = 0.01
+    p = 0.1
     #Set to False convergence flags
     conv = False
     conv_per = False
@@ -48,19 +48,22 @@ def main(argv):
     while sim_it < n_sim:
         print('\nSimulation number: ', sim_it)
         #Keep sampling and integrating until convergence
+        #Initialize list to keep track of extinct plants and soils
+        ext_ind = []
         while not conv or not conv_per:
             #Sample matrices A and B with H-S constraints
             A, B = sampling_matrices(n)
             A_per = A + np.random.uniform(1-p, 1+p, size = np.shape(A))
+            B_per = B
             #Integrate unperurbed system
-            plant_ab, soil_ab, t = integrate_PSF(model, [1, 2000], 2*n*[1/n], 
+            plant_ab, soil_ab, t = integrate_PSF(model, [0, 2000], 2*n*[1/n], 
                                                  (A, B, n))
             #Check convergence of unperurbed system
             conv = check_convergence(plant_ab[:, -1], soil_ab[:, -1], tol)
             #Integrate perurbed system
-            plant_ab_per, soil_ab_per, t_per = integrate_PSF(model, [1, 2000], 
+            plant_ab_per, soil_ab_per, t_per = integrate_PSF(model, [0, 2000], 
                                                              2*n*[1/n], 
-                                                             (A_per, B, n))
+                                                             (A_per, B_per, n))
             #Check convergence of perurbed system
             conv_per = check_convergence(plant_ab_per[:, -1], 
                                          soil_ab_per[:, -1], tol)
@@ -71,41 +74,64 @@ def main(argv):
         #Check if the convergent perturbed system has arrived to equilibrium
         equilibrium = check_equilibrium(rem_plant, 
                                         rem_soil)
+        #Update number of plants
+        n_plant = len(rem_plant)
+        #Remove extinct species from payoff matrices
+        ext_plant, ext_soil = find_extinct_indices(plant_ab_per[:, -1], 
+                                                   rem_plant, 
+                                                   soil_ab_per[:, -1],
+                                                   rem_soil)
+        #Record extinctions
+        ext_ind += list(ext_plant)
         #Set number of integration cycles to zero
         n_int = 0
-        #Initialize list to keep track of extinct plants and soils
-        ext_ind = []
+        #Store
[…]

### 2021-10-26 code/plant_feedback_model.py

@@ -109,6 +109,7 @@ def main(argv):
                     rem_plant, rem_soil = remove_extinctions(plant_ab[:, -1], 
                                                              soil_ab[:, -1], 
                                                              tol) 
+                    n_plant = len(rem_plant)
                     #Check equilibrium
                     equilibrium = check_equilibrium(rem_plant, rem_soil)
             #Check if while loop exited due to exceeding integration cycles or

### 2021-10-26 code/quick_simulation.py

@@ -1,0 +1,106 @@
+#!/usr/bin/env python3
+
+__appname__ = '[App_name_here]'
+__author__ = 'Pablo Lechon (plechon@ucm.es)'
+__version__ = '0.0.1'
+
+## IMPORTS ##
+
+import sys
+import numpy as np
+import pandas as pd
+from functions import *
+import matplotlib.pylab as plt
+
+## CONSTANTS ##
+
+
+## FUNCTIONS ##
+
+def main(argv):
+    '''Main function'''
+
+    A = np.array(pd.read_csv('../data/A_pre.csv', delimiter = ' ', 
+                 header = None))
+    A_per = np.array(pd.read_csv('../data/A_post.csv', delimiter = ' ', 
+                     header = None))
+    mu = pd.read_csv('../data/mu.csv', delimiter = ' ', header = None)
+    B = np.diag(np.array(mu).flatten())
+    B_per = B
+    n = len(A)
+    n_per = n
+    z0 = 2*n*[1/n] 
+    z0_per = z0
+    tol = 1e-9
+    equilibrium = 0
+    n_int = 0
+    ext_ind = []
+    p_ab_per = []
+    t_vec = []
+    p_l = []
+    present = np.ones(n)
+    t_max = 2000
+    while not equilibrium:
+        #Integrate unperurbed system
+        plant_ab_per, soil_ab_per, t_per = integrate_PSF(model, [0, t_max], 
+                                                         z0_per, 
+                                                         (A_per, B, n), 
+                                                         method = 'RK45')
+        #Delete extinctions 
+        rem_plant, rem_soil = remove_extinctions(plant_ab_per[:, -1], 
+                                                 soil_ab_per[:, -1], 
+                                                 tol) 
+        #Check if we have reached equilibrium
+        equilibrium = check_equilibrium_bis(plant_ab_per[:, -1], 
+                                            soil_ab_per[:, -1])
+        #Update number of plants
+        n_per = len(rem_plant)
+        #Find extinct indices of plants and soils 
+        ext_plant, ext_soil = find_extinct_indices(plant_ab_per[:, -1], 
+                                                   rem_plant, 
+                                                   soil_ab_per[:, -1],
+                                                   rem_soil)
+        #Set extinct species to 0
+        plant_ab_per[ext_plant, :] = 0
+        soil_ab_per[ext_soil, :] = 0
[…]

### 2021-10-26 code/dynamics_collapse.R
some  edits to figure 2

@@ -91,14 +91,14 @@ correlation_plot = ggplot(data = data,
         legend.position = 'none',
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15),
-        axis.text = element_text(size = 15))+
+        axis.text = element_text(size = 10))+
   labs(x = 'A', y = expression(A[p]))
   
 print(correlation_plot)
 
-yheight = 0.6
-xheight = -0.335
-height = 0.15
+yheight = 0.63
+xheight = -0.30
+height = 0.34
 ptot1 = ggdraw() +
   draw_plot(p1) +
   draw_plot(A_pre_plot, x = xheight, y = yheight, height = height)
@@ -110,7 +110,8 @@ ptot2 = ggdraw() +
 lay <- rbind(c(1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 3),
              c(1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 3),
              c(1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 3))
+pdf('../sandbox/dynamics_collapse.pdf', height = 3.5, width  = 12)
 grid.arrange(ptot1, ptot2, correlation_plot,
              layout_matrix = lay)
-
+dev.off()
 

### 2021-10-26 code/dynamics_collapse.R
some  edits to figure 2

@@ -3,6 +3,7 @@ library(tidyquant)
 library(reshape2)
 library(cowplot)
 library(gridExtra)
+library(RColorBrewer)
 
 
 data = read_csv('../data/zach_results.csv')
@@ -12,7 +13,7 @@ p1 = ggplot(data = data,
            color = as.factor(p_l)))+
   geom_line(alpha = 0.2)+
   geom_ma(ma_fun = EMA, n = 150, linetype = 'solid')+
-theme(aspect.ratio = 3/4,
+  theme(aspect.ratio = 3/4,
             panel.background = element_blank(),
             panel.border = element_rect(fill = NA),
             legend.position = 'none',
@@ -20,6 +21,7 @@ theme(aspect.ratio = 3/4,
             axis.title.y = element_text(size = 15),
             axis.text = element_blank(),
             axis.ticks = element_blank())+
+  scale_color_brewer(palette = 'Set1')+
   labs(y = "Abundance",
        x = "Time")
 print(p1)
@@ -37,6 +39,7 @@ p2 = ggplot(data = data,
         axis.title.y = element_text(size = 15),
         axis.text = element_blank(),
         axis.ticks = element_blank())+
+  scale_color_brewer(palette = 'Set1')+
   labs(y = "",
        x = "Time")
 print(p2)
@@ -52,13 +55,21 @@ A_pre_plot =  ggplot(data = A_pre_melt,
   geom_tile()+
   theme(panel.background = element_rect(fill = NA),
         plot.background = element_blank(),
-        legend.position = 'none',
+        # legend.title = element_text(size = 15, hjust = 0.2,
+        #                             margin = margin(b = -10)),
+        legend.title = element_blank(),
+        legend.position = c(1.16, 0.55),
+        legend.background = element_rect(fill = NA),
         axis.text = element_blank(),
         axis.ticks = element_blank(),
         axis.title = element_blank(),
         aspect.ratio = 1)+
-    labs(x = '', y = '')
-
+  labs(x = '', y = '',
+       fill = expression(A[ij]))+
+  scale_fill_distiller(palette = 'RdYlBu',
+                       guide = guide_colorbar(barwidth = 0.6,
+                                              barheight = 4.5),
+                       breaks = c(0.5, 0.9, 1.3))
 print(A_pre_plot)
 
 A_post = my_list$A_post
@@ -69,17 +80,29 @@ A_post_plot =  ggplot(data = A_post_melt,
   geom_tile()+
[…]

### 2021-10-26 paper/paper.tex
updated ms with cover letter and title page

@@
 \usepackage{float}
-\usepackage[sort,numbers]{natbib}
+\usepackage{natbib}
+
+\setcitestyle{round,aysep={}}
 
@@
 
+\date{}
+\author{Zachary R. Miller$^{1*}$ (zachmiller@uchicago.edu), Pablo Lech\'{o}n$^{1}$ (plechon@uchicago.edu),\\ and Stefano Allesina$^{1,2}$ (sallesina@uchicago.edu) \\
+	\\
+	\normalsize{$^{1}$Department of Ecology \& Evolution, University of Chicago, Chicago, IL, USA}\\
+	\normalsize{$^{2}$Northwestern Institute on Complex Systems, Evanston, IL, USA}\\
+	\\
+	\normalsize{*corresponding author (mailing: 1101 E. 57th Street, Chicago, IL 60637; phone: 484-331-4593)}\\
+}
+
 \begin{document}
+
 \maketitle
 \setstretch{1.5}
+
+\noindent Running title: No coexistence in a plant-soil feedback model\\
+Keywords: plant-soil feedbacks, legacy effects, coexistence, Bever model, bimatrix games, replicator equation \\
+
+\noindent Author contributions: ZRM and SA performed the mathematical analysis, and PL implemented the numerical simulations.
+ZRM wrote the manuscript, and all authors edited and discussed the text.
+\\
+
+\noindent Data accessibility: Code for reproducing the numerical simulations is available at [ADD REPOSITORY].\\
+
+\noindent Article type: Letter\\
+Number of words in abstract: 137\\
+Number of words in main text: \\
+Number of references: 51 \\
+Number of figures: 3
+
 \linenumbers
@@
 \begin{abstract}
-  [Abstract]
+Plant-soil feedbacks (PSFs) are thought to represent a crucial mechanism generating frequency dependent dynamics in plant communities.
+Negative feedbacks, in particular, are routinely invoked to explain coexistence and the maintenance of diversity in species-rich communities.
+However, the primary modeling framework used to study PSFs considers only two plant species, and we lack clear theoretical expectations for how these complex interactions play out in communities with natural levels of diversity.
+Here, we demonstrate that this canonical model for PSFs is equivalent to a well-studied model from evolutionary game theory, and using this equivalence, we are able to characterize the dynamics with an arbitrary number of plant species.
+Surprisingly, we find that coexistence of more than two species is virtually impossible in this model, suggesting that alternative theoretical frameworks are needed to describe feedbacks observed in diverse natural communities.
 \end{abstract}
@@
 It has become well understood that reciprocal interactions between plants and the soil biota, known as plant-soil feedbacks (PSFs), play an important role in structuring the composition and dynamics of plant communities.
-PSFs operate alongside other factors, including abiotic drivers \cite{bennett2019mechanisms} and above-ground trophic interactions \cite{van2009empirical}, but are thought to be a key mechanism generating negative frequency-dependent feedbacks that promote coexistence and maintain plant diversity \cite{kulmatiski2008plant,van2013plant,bever2015maintenance}.
-The existence of PSFs has long been known \cite{van1993plant,bever1994feedback}, but our understanding of their importance -- particularly in relation to patterns of coexistence -- has developed rapidly in recent years \cite{klironomos2002feedback,petermann2008janzen,mangan2010negative}.
-Broad interest in PSFs was ignited by the development of simple mathematical models, which illustrated the potential of PSFs to mediate plant coexistence \cite{bever1997incorporating,bever2003soil,ke2015incorporating}.
-These models have played a guiding role for a wide range of empirical studies, as well \cite{kulmatiski2008plant,kulmatiski2011testing,pernilla2010plant}.
+PSFs operate alongside other factors, including abiotic drivers \citep{bennett2019mechanisms} and above-ground trophic interactions \citep{van2009empirical}, but are thought to be a key mechanism generating negative frequency-dependent feedbacks that promote coexistence and maintain plant diversity \citep{kulmatiski2008plant,van2013plant,bever2015maintenance}.
+The existence of PSFs has long been known \citep{van1993plant,bever1994feedback}, but our understanding of their importance -- particularly in relation to patterns of coexistence -- has developed rapidly in recent years \citep{klironomos2002feedback,petermann2008janzen,mangan2010negative}.
+Broad interest in PSFs was ignited by the development of simple mathematical models, which illustrated the potential of PSFs to mediate plant coexistence \citep{bever1997incorporating,bever2003soil,ke2015incorporating}.
+These models have played a guiding role for a wide range of empirical studies, as well \citep{kulmatiski2008plant,kulmatiski2011testing,pernilla2010plant}.
 
-The first, and still most widely known and used, model for PSFs was introduced by Bever and colleagues in the 1990s \cite{bever1992ecological,bever1997incorporating,bever1999dynamics,bever2003soil}.
+The first, and still most widely known and used, model for PSFs was introduced by Bever and colleagues in the 1990s \citep{bever1992ecological,bever1997incorporating,bever1999dynamics,bever2003soil}.
 In this framework, often referred to simply as the Bever model, each plant species is assumed to promote the growth of a specific soil component (i.e.
@@
 In turn, the fitness of each plant species is impacted by the relative frequency of different soil components.
-Starting from minimalist assumptions, Bever et al.
-\cite{bever1997incorporating} derived a set of differential equations to capture these dynamics.
+Starting from minimal assumptions, Bever et al.
+\citep{bever1997incorporating} derived a set of differential equations to capture these dynamics.
 PSFs can be either positive (fitness of a plant species is increased by its corresponding soil component) or negative (a plant species experiences lower relative fitness in its own soil).
 Bever et al.
-introduced a single quantity to summarize whether PSFs are positive or negative, and showed that this value characterizes the dynamical behavior of the model.
+introduced a single quantity to summarize whether community-wide PSFs are positive or negative, and showed that this value characterizes the dynamical behavior of the model.
 In the original Bever model of two plant species, positive PSFs lead to exclusion of one species, while negative PSFs result in neutral oscillations.
-It is thus widely suggested that negative PSFs help sustain coexistence in real-world plant communities \cite{kulmatiski2008plant,van2013plant}, perhaps with spatial asynchrony playing a role in stabilizing the cyclic dynamics \cite{revilla2013plant,bever2003soil}.
+It is thus widely suggested that negative PSFs help sustain coexistence in real-world plant communities \citep{kulmatiski2008plant,van2013plant}, perhaps with spatial asynchrony playing a role in stabilizing the cyclic dynamics \citep{revilla2013plant,bever2003soil}.
 
-Subsequent studies have generalized this model to include, for example, more realistic functional forms \cite{umbanhowar2005simple, eppinga2006accumulation}, more explicit representations of the soil community \cite{bever2010rooting}, spatial structure \cite{eppinga2006accumulation,molofsky2002negative,suding2013consequences}, or additional processes such as direct competitive interactions between plants \cite{bever2003soil}.
-However, the original Bever model remains an important touchstone for the theory of PSFs \cite{ke2015incorporating,ke2020effects}, and informs empirical research through the interaction coefficient, $I_s$, derived by Bever et al., which is commonly measured and used to draw conclusions about coexistence in experimental studies.
-Despite the ubiquity of this model, and the fruitful interplay of theory and experiment in the PSF literature, extensions to communities with more than two or three species have appeared only rarely and recently (but see \cite{eppinga2018frequency,mack2019plant}).
-While PSF models motivate hypotheses and conclusions about species-rich natural communities, there is much still unknown about the behavior of these models with natural levels of diversity \cite{van2013plant}.
+Subsequent studies have generalized this model to include, for example, more realistic functional forms \citep{umbanhowar2005simple, eppinga2006accumulation}, more explicit representations of the soil community \citep{bever2010rooting}, spatial structure \citep{eppinga2006accumulation,molofsky2002negative,suding2013consequences}, or additional processes such as direct competitive interactions between plants \citep{bever2003soil}.
+However, the original Bever model remains an important touchstone for the theory of PSFs \citep{ke2015incorporating,ke2020effects}, and informs empirical research through the interaction coefficient, $I_s$, derived by Bever et al., which is commonly measured and used to draw conclusions about coexistence in experimental studies.
+Despite the ubiquity of this model, and the fruitful interplay of theory and experiment in the PSF literature, extensions to communities with more than two or three species have appeared only rarely and recently (but see \citep{eppinga2018frequency,mack2019plant}).
+While PSF models motivate hypotheses and conclusions about species-rich natural communities, there is much still unknown about the behavior of these models with natural levels of diversity \citep{van2013plant}.
 
-In this Letter, we extend the Bever model to include any number of plant species, and show that the model is equivalent to a special form of the replicator equation studied in evolutionary game theory \cite{hofbauer1998evolutionary}.
+Here, we extend the Bever model to include any number of plant species, and show that the model is equivalent to a special form of the replicator equation studied in evolutionary game theory \citep{hofbauer1998evolutionary}.
 In particular, this model corresponds to the class of bimatrix games, where there are two players (here, plants and soil components) which interact with asymmetric strategies and payoffs.
-The replicator dynamics of bimatrix games are well-studied, allowing us to characterize many properties of the Bever model with $n$ species.
+The replicator dynamics of bimatrix games are well-studied, allowing us to characterize many properties of the Bever model with $n$ plant species.
 Surprisingly, using this equivalence, we show that coexistence of more than two species in this model is never robust.
@@
 
-Inspired by emerging empirical evidence for the important role of PSFs in plant community dynamics and coexistence \cite{van1993plant,bever1994feedback}, Bever et al.
-\cite{bever1997incorporating} introduced a simple mathematical model to investigate their behavior.
-In this model, two plant species, $1$ and $2$, grow exponentially with growth rates determined by the state of the soil biota in the system (note we adopt a different notation than Bever et al.).
+Inspired by emerging empirical evidence for the important role of PSFs in plant community dynamics and coexistence \citep{van1993plant,bever1994feedback}, Bever et al.
+\citep{bever1997incorporating} introduced a simple mathematical model to investigate their behavior.
+In this model, two plant species, $1$ and $2$, grow exponentially with growth rates determined by the state of the soil biota in the system.
 These effects of soil on plants are specified by parameters $\alpha_{ij}$, the growth rate of plant species $i$ in soil type $j$.
@@
 \begin{cases}
-p_i^\star &= \frac{1}{1 + \frac{\beta_i}{\beta_j}} \, , \quad  i, j = 1,2 \\
+p_i^\star &= \frac{\beta_j}{\beta_i + \beta_j} \, , \quad  i, j = 1,2 \\
 q_i^\star &= \frac{\alpha_{jj} - \alpha_{ji}}{\alpha_{ii} - \alpha_{ji} - \alpha_{ij} + \alpha_{jj}} \, .
@@
 In fact, these conclusions depend on the existence of a feasible equilibrium (i.e.
-positive equilibrium values), which further requires that $\alpha_{ii} < \alpha_{ji}$, for both $i, j = 1, 2$, in order for the model to exhibit coexistence \cite{bever1997incorporating,ke2015incorporating}.
+positive equilibrium values), which further requires that $\alpha_{ii} < \alpha_{ji}$, for both $i, j = 1, 2$, in order for the model to exhibit coexistence \citep{bever1997incorporating,ke2015incorporating}.
 
@@
 Species frequencies oscillate neutrally, similar to the textbook example of Lotka-Volterra predator-prey dynamics.
-Any stochasticity, external forcing, or time variation in the model parameters can destroy these finely balanced oscillations and cause one species to go extinct \cite{revilla2013plant}.
-However, coupled with mechanisms that buffer the system from extinctions, such as migration between desynchronized patches or the presence of a seed bank, the negative feedbacks in this model might produce sustained coexistence \cite{revilla2013plant, bever2003soil}.
+Any stochasticity, external forcing, or time variation in the model parameters can destroy these finely balanced oscillations and cause one species to go extinct \citep{revilla2013plant}.
+However, coupled with mechanisms that buffer the system from extinctions, such as migration between desynchronized patches or the presence of a seed bank, the negative feedbacks in this model might produce sustained coexistence \citep{revilla2013plant, bever2003soil}.
 
-Of course, most natural plant communities feature more than two coexisting species, and it is precisely in the most diverse communities where mechanisms of coexistence hold the greatest interest \cite{van2013plant}.
+Of course, most natural plant communities feature more than two coexisting species, and it is precisely in the most diverse communities where mechanisms of coexistence hold the greatest interest \citep{van2013plant}.
 While it is not immediately clear how to generalize Eq.~\ref{2spp_relative} to more than two species, Eq.~\ref{2spp_absolute} is naturally extended by maintaining the assumption that the overall growth rate for any plant is a weighted average of its growth rate in each soil type:
@@
 where vectors are indicated in boldface (e.g.
-$\bm{p}$ is the vector of species frequencies $(p_1, p_2, \dots p_n)$ and $\bm{1}$ is a vector of $n$ ones) and $D(\bm{z})$ is the diagonal matrix with vector $\bm{z}$ on the diagonal.
+$\bm{p}$ is the vector of species frequencies $(p_1, p_2, \dots p_n)^T$ and $\bm{1}$ is a vector of $n$ ones) and $D(\bm{z})$ is the diagonal matrix with vector $\bm{z}$ on the diagonal.
 We have introduced the matrices $A = (\alpha_{ij})$ and $B = D(\beta_1, \beta_2, \dots \beta_n)$, specifying soil effects on plants and plant effects on soil, respectively.
-Because $\bm{p}$ and $\bm{q}$ are vectors of frequencies, they must sum to one: $\bm{1}^T \bm{p} = \bm{1}^T \bm{p} = 1$.
+Because $\bm{p}$ and $\bm{q}$ are vectors of frequencies, they must sum to one: $\bm{1}^T \bm{p} = \bm{1}^T \bm{q} = 1$.
 Using these constraints, one can easily show that the Bever model (Eq.~\ref{2spp_relative}) is a special case of Eqs.~\ref{nspp_relative} and \ref{matrix_form} when $n = 2$.
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_1.pdf}
+	\caption[Structure of the $n$-species Bever model]
+	{The model described by Eqs.~\ref{nspp_relative}-\ref{matrix_form} is shown here graphically.
+Plant species (green squares) promote the growth of their respective soil components (brown circles) at a rate $\beta_i$ (gray arrows).
+In turn, growth of each plant is governed by the mix of soil components present in the system, with the effect of soil component $j$ on species $i$ quantified by the parameter $\alpha_{ij}$ (black arrows).
+This model is a straightforward extension of the model proposed by Bever et al.
+\citep{bever1997incorporating} to an arbitrary number of species.
+Note only selected parameter labels are shown for clarity.}
+	\label{fig:concept}
+\end{figure*}
 
@@
 
-Systems taking the form of Eq.~\ref{matrix_form} are well-known and well-studied in evolutionary game theory.
-Our generalization of the Bever model is a special case of the \emph{replicator equation}, corresponding to the class of \emph{bimatrix games} \cite{taylor1979evolutionarily,hofbauer1996evolutionary,hofbauer1998evolutionary,cressman2014replicator}.
-Bimatrix games arise in diverse contexts, such as animal behavior \cite{taylor1979evolutionarily,selten1988note}, evolutionary theory \cite{hofbauer1998evolutionary,cressman2014replicator}, and economics \cite{friedman1991evolutionary}, where they model games with asymmetric players.
+Systems that take the form of Eq.~\ref{matrix_form} are well-known and well-studied in evolutionary game theory.
+Our generalization of the Bever model is a special case of the \emph{replicator equation}, corresponding to the class of \emph{bimatrix games} \citep{taylor1979evolutionarily,hofbauer1996evolutionary,hofbauer1998evolutionary,cressman2014replicator}.
+Bimatrix games arise in diverse contexts, such as animal behavior \citep{taylor1979evolutionarily,selten1988note}, evolutionary theory \citep{hofbauer1998evolutionary,cressman2014replicator}, and economics \citep{friedman1991evolutionary}, where they model games with asymmetric players.
 In a bimatrix game, each player (here, the plant community and the soil) has a distinct set of strategies (plants species and soil components, respectively) and payoffs (realized growth rates).
@@
 Much is known about bimatrix game dynamics, and we can draw on this body of knowledge to characterize the behavior of the Bever model with $n$ species.
-Essential mathematical background and details are presented in the Supplemental Methods; for a detailed introduction to bimatrix games, see Hofbauer and Sigmund (\cite{hofbauer1998evolutionary}).
+Essential mathematical background and details are presented in the Supplemental Methods; for a detailed introduction to bimatrix games, see Hofbauer and Sigmund \citep{hofbauer1998evolutionary}.
 
@@
 As we have seen, when the community consists of two species, the coexistence equilibrium, if feasible, can be either unstable or neutrally stable.
-In fact, the same is true for the $n$-species extension (and, more generally, for any bimatrix game dynamics \cite{eshel1983coevolutionary,selten1988note,hofbauer1998evolutionary}).
+In fact, the same is true for the $n$-species extension (and, more generally, for any bimatrix game dynamics \citep{eshel1983coevolutionary,selten1988note,hofbauer1998evolutionary}).
 This can be established using straightforward local stability analysis, after accounting for the relative abundance constraints, which imply $p_n = 1 - \sum_{i = 1}^{n-1} p_i$ and $q_n = 1 - \sum_{i = 1}^{n-1} q_i$.
@@
 
-Another notable property of bimatrix game dynamics is that the vector field defined by the model equations is divergence-free or incompressible (see \cite{hofbauer1998evolutionary} for a proof).
-The divergence theorem from vector calculus \cite{arfken1985mathematical} then tells us that Eq.~\ref{matrix_form} cannot have any attractors -- that is, regions of the phase space that ``pull in'' trajectories -- with multiple species.
+Another notable property of bimatrix game dynamics is that the vector field defined by the model equations is divergence-free or incompressible (see \citep{hofbauer1998evolutionary} for a proof).
+The divergence theorem from vector calculus \citep{arfken1985mathematical} then tells us that Eq.~\ref{matrix_form} cannot have any attractors -- that is, regions of the phase space that ``pull in'' trajectories -- with multiple species.
 This rules out coexistence in a stable limit cycle or other non-equilibrium attractors (e.g.
@@
 In particular, the model parameters must satisfy the conditions $\alpha_{ij} = \gamma_i + \delta_j$ for some constants $\gamma_i, \delta_i$ in $i = 1, \dots, n$ (when $i \neq j$), and $\alpha_{ii} = \gamma_i + \delta_i- c \beta_i$ (where $c$ is a positive constant independent of $i$).
-In the language of bimatrix games, such systems are called \emph{rescaled zero-sum games} \cite{hofbauer1996evolutionary,hofbauer1998evolutionary}.
-It is a long-standing conjecture in evolutionary game theory that these parameterizations are the \emph{only} cases where $n$-species coexistence can occur \cite{hofbauer1996evolutionary,hofbauer2011deterministic}.
+In the language of bimatrix games, such systems are called \emph{rescaled zero-sum games} \citep{hofbauer1996evolutionary,hofbauer1998evolutionary}.
+It is a long-standing conjecture in evolutionary game theory that these parameterizations are the \emph{only} cases where $n$-species coexistence can occur \citep{hofbauer1996evolutionary,hofbauer2011deterministic}.
 
 Ecologically, these conditions mean there is a fixed effect of each soil type and plant species identity, and the growth rate of plant $i$ in soil type $j$ is the additive combination of these two, with no interaction effects.
-The only exception is for plants growing in their own soil type, which must experience a fitness cost exactly proportional to the rate at which they promote growth of the soil type ($\beta_i$).
+The only exception is for plants growing in their own soil type, which must experience a fitness cost ($\gamma_i + \delta_i - \alpha_{ii}$) exactly proportional to the rate at which they promote growth of the soil type ($\beta_i$).
 These conditions clearly extend the intuitive notion that each plant must have a disadvantage in its corresponding soil type to allow coexistence.
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_2.pdf}
+	\caption[Final community sizes with varying initial richness]
+	{Final community sizes with varying initial richness.
+We show the distribution of final richness (number of species, in red) for 500 (???) communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
+Parameters $\alpha_{ij}$ and $\beta_{i}$  were chosen randomly \emph{iid} from a standard uniform distribution, $U[0,1]$.
+For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.
+\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
+In agreement with the conjecture that coexistence of more than two species is vanishingly unlikely, we found that, regardless of the number of species initially present, every community collapsed to a subset of one or two surviving species.}
+	\label{fig:histograms}
+\end{figure*}
+
 \section{Discussion}
 
-The Bever model has played a central role in motivating PSF research, and continues to guide both theory and experiment in the fast-growing field \cite{bever2015maintenance,kandlikar2019winning,ke2020effects}.
-In this Letter, we extend the Bever model to any number of species, and highlight its equivalence to bimatrix game dynamics.
+The Bever model has played a central role in motivating PSF research, and continues to guide both theory and experiment in the fast-growing field \citep{bever2015maintenance,kandlikar2019winning,ke2020effects}.
+Here, we extend the Bever model to any number of species, and highlight its equivalence to bimatrix game dynamics.
 Taking advantage of the well-developed theory for these dynamics, we are able to characterize the behavior of the generalized Bever model in detail.
@@
 
-This result may be surprising, because a significant body of experimental evidence indicates that PSFs play a key role in mediating the coexistence of more than two species in natural communities \cite{kulmatiski2008plant,petermann2008janzen,mangan2010negative,bever2015maintenance}.
+This result may be surprising, because a significant body of experimental evidence indicates that PSFs can play an important role in mediating the coexistence of more than two species in natural communities \citep{kulmatiski2008plant,petermann2008janzen,mangan2010negative,bever2015maintenance}.
 Apparently, the picture suggested by the two-species Bever model generalizes in nature, but not in the model framework itself.
 We note that this framework was introduced as an intentional simplification to illustrate the potential role of PSFs in mediating coexistence, not to accurately model the biological details of PSFs.
-In this capacity, the model has been wildly successful.
-And indeed, alongside empirical study of PSFs, other modeling approaches have emerged, accounting for more biological realism (e.g.
-\cite{umbanhowar2005simple,eppstein2007invasiveness,bever2010rooting}), or with the demonstrated capacity to produce multispecies coexistence (e.g.
-\cite{bonanomi2005negative,miller2021metapopulations}).
-Some of these are minor modifications of the Bever model framework; others build on distinct foundations \cite{ke2015incorporating,ke2020effects}.
+Indeed, the model has been wildly successfully in spurring research into PSFs.
+Alongside empirical study of these processes, other modeling approaches have emerged, accounting for more biological realism (e.g.
+\citep{umbanhowar2005simple,eppstein2007invasiveness,bever2010rooting}), or with the demonstrated capacity to produce multispecies coexistence (e.g.
+\citep{bonanomi2005negative,miller2021metapopulations}).
+Some of these are minor modifications of the Bever model framework; others build on distinct foundations \citep{ke2015incorporating,ke2020effects}.
 Our results suggest that these various avenues are worth pursuing further.
@@
 Our analysis cautions that applications of this model in multispecies communities might lead to incorrect inference.
-For example, attempts to parameterize the Bever model for three species using empirical data have produced predictions of non-coexistence in plant communities that seem to coexist experimentally \cite{kulmatiski2011testing}.
-In many other studies, the pairwise interaction coefficient, $I_s$, is calculated for species pairs and used to assess whole-community coexistence \cite{kulmatiski2008plant,fitzsimons2010importance,pendergast2013belowground,suding2013consequences,kuebbing2015plant,smith2015plant,bauer2017effects,kandlikar2019winning}.
+For example, attempts to parameterize the Bever model for three species using empirical data have produced predictions of non-coexistence in plant communities that coexist experimentally \citep{kulmatiski2011testing}.
+In many other studies, the pairwise interaction coefficient, $I_s$, is calculated for species pairs and used to assess whole-community coexistence \citep{kulmatiski2008plant,fitzsimons2010importance,pendergast2013belowground,suding2013consequences,kuebbing2015plant,smith2015plant,bauer2017effects,kandlikar2019winning}.
 However, we have seen that whole-community coexistence is virtually impossible within the generalized model, and there is no guarantee that the pairwise coexistence conditions for this model will agree with $n$-species coexistence conditions in other frameworks.
-For example, $I_s < 0$ for all species pairs is neither necessary nor sufficient to produce coexistence in a metapopulation-based model for PSFs \cite{miller2021metapopulations}.
+For example, $I_s < 0$ for all species pairs is neither necessary nor sufficient to produce coexistence in a metapopulation-based model for PSFs \citep{miller2021metapopulations}.
 
-In contrast to other extensions of the Bever model to many-species communities \cite{eppinga2018frequency,mack2019plant}, our approach keeps the dynamics of both plants and soil fully explicit.
+In contrast to other extensions of the Bever model to many-species communities \citep{eppinga2018frequency,mack2019plant}, our approach keeps the dynamics of both plants and soil fully explicit.
 As such, we make no assumptions about the relative timescales of plant and soil dynamics, or whether either of these reach equilibrium.
-This difference likely explains the discrepancy between our conclusions and previously published predictions for $n$-species PSFs \cite{kulmatiski2011testing,eppinga2018frequency,mack2019plant}.
+This difference likely explains the discrepancy between our conclusions and previously published predictions for $n$-species PSFs \citep{kulmatiski2011testing,eppinga2018frequency,mack2019plant}.
 
-PSFs as envisioned in the classic Bever model might facilitate coexistence in conjunction with other mechanisms, such as direct species interactions, or through long transient dynamics, but our analysis shows that they cannot produce robust $n$-species coexistence in isolation.
+PSFs as envisioned in the classic Bever model might facilitate coexistence in conjunction with other mechanisms, such as direct species interactions \citep{bever2003soil}, or through long transient dynamics, but our analysis shows that they cannot produce robust $n$-species coexistence in isolation.
 This finding calls for renewed theoretical investigation of PSFs.
 One important consideration is grounding PSF modeling frameworks in more realistic models for absolute abundances or densities.
-As various researchers have noted \cite{revilla2013plant,eppinga2018frequency,ke2020effects}, the Bever model and its extensions are in fact projections (onto the space of relative abundances or frequencies) of dynamics for plant and soil \emph{abundances}.
+As various researchers have noted \citep{kulmatiski2011testing,revilla2013plant,eppinga2018frequency,ke2020effects}, the Bever model and its extensions are in fact projections (onto the space of relative abundances, or frequencies) of dynamics for plant and soil \emph{abundances}.
 Consequently, the projected dynamics can mask unbiological outcomes in the original model (e.g.
 relative abundances oscillate around equilibrium while absolute abundances shrink to zero or explode to infinity).
-Indeed, the absolute abundance model (Eq.~\ref{nspp_absolute}) used to derive our $n$-species frequency dynamics (Eqs.~\ref{nspp_relative}-\ref{matrix_form}) does not generally possess any fixed points, which is a basic requirement for species coexistence \cite{hutson1990existence,hutson1992permanence}.
-It is usually seen as desirable to study PSFs in the space of species frequencies, both because this facilitates connections to data, and because frequencies are considered a more appropriate metric for analyzing processes that stabilize coexistence (\cite{adler2007niche,eppinga2018frequency} but see \cite{kandlikar2019winning,ke2020effects}).
+Indeed, the absolute abundance model (Eq.~\ref{nspp_absolute}) used to derive our $n$-species frequency dynamics (Eqs.~\ref{nspp_relative}-\ref{matrix_form}) does not generally possess any fixed points, which is a basic requirement for species coexistence \citep{hutson1990existence,hutson1992permanence}.
+It is usually seen as desirable to study PSFs in the space of species frequencies, both because this facilitates connections to data, and because frequencies are considered a more appropriate metric for analyzing processes that stabilize coexistence (\citep{adler2007niche,eppinga2018frequency} but see \citep{kandlikar2019winning,ke2020effects}).
 But models that introduce frequencies through a natural constraint, such as competition for finite space, will likely produce more realistic dynamics.
@@
 From a broader theoretical perspective, the qualitative change in model behavior that we observe as the number of species increases from two to three or more is a striking phenomenon, but not an unprecedented one.
-Ecologists have repeatedly found that intuitions from two-species models can generalize (or fail to generalize) to more diverse communities in surprising ways \cite{strobeck1973n,smale1976differential,barabas2016effect}.
-Our analysis provides another illustration of the fact that ``more is different'' \cite{anderson1972more} in ecology, and highlights the importance of developing theory for species-rich communities.
+Ecologists have repeatedly found that intuitions from two-species models can generalize (or fail to generalize) to more diverse communities in surprising ways \citep{strobeck1973n,smale1976differential,barabas2016effect}.
+Our analysis provides another illustration of the fact that ``more is different'' \citep{anderson1972more} in ecology, and highlights the importance of developing theory for species-rich communities.
 
@@
 
-\bibliographystyle{unsrt}
+\bibliographystyle{ecol_let}
 \bibliography{references}
-
-\section*{Supplemental Methods}
-
-\subsection{Model derivation}
-
-As described in the Main Text, we begin with the system
-
-\begin{align}
-\begin{cases}
-\frac{dx_i}{dt} &= x_i \left(\sum_{j} \alpha_{ij} \, q_j \right) \, , \quad  i = 1, \dots n \\
-\frac{dy_i}{dt} &= y_i \left( \beta_i \, p_i \right)
-\end{cases}
-\end{align}
-governing the time-evolution of plant abundances $x_i$ and soil components $y_i$, where $p_i = x_i / \sum_j x_j$, $q_i = y_i / \sum_j y_j$, and Greek letters denote nonnegative parameters.
-These equations capture the assumptions outlined by Bever et al.
-\cite{bever1997incorporating} for two species and extend them straightforwardly to any $n$ species.
-Following the approach of Bever et al.
-for two species (and consistent with other generalizations of this model \cite{kulmatiski2008plant,eppinga2018frequency}), we derive dynamics for the $p_i$ by applying the chain rule:
-
-\begin{align} \label{derive_relative}
-\begin{split}
-	\frac{dp_i}{dt} &= \frac{d}{dt} \frac{x_i}{\sum x_j}\\  
-	&= \frac{1}{\sum_j x_j} \frac{dx_i}{dt} - \frac{x_i}{(\sum_j x_j)^2} \sum_j \frac{dx_j}{dt} \\
-	&= \frac{x_i}{\sum_j x_j} \left( \sum_{j} \alpha_{ij} \, q_j \right) - \frac{x_i}{\sum_j x_j} \left(\sum_j \frac{x_j}{\sum_k x_k} \sum_{l} \alpha_{jl} \, q_l \right)\\
-	&= p_i \left(\sum_{j} \alpha_{ij} \, q_j -  \sum_{j, k} \alpha_{jk} p_j  q_k \right) \, .
-\end{split}
-\end{align}
-This last expression is identical to the first line of Eq.~\ref{nspp_relative} in the Main Text.
-The dynamics for $q_i$ can be derived in exactly the same way (using the definitions $\beta_{ii} = \beta_i$ and $\beta_{ij} = 0$ to make the parallel clear).
-The two terms of each per capita growth rate in Eq.~\ref{nspp_relative} have natural interpretations in the language and notation of linear algebra: $\sum_{j} \alpha_{ij} \, q_j$ is the $i$th component of the matrix-vector product $A \bm{q}$ and $\sum_{j, k} \alpha_{jk} p_j  q_k$ is the bilinear form $\bm{p}^T A \bm{q}$.
-Here, $A$ (and $B$) is an $n \times n$ matrix and $\bm{p}$ ($\bm{q}$) is a vector of length $n$, as described in the Main Text.
-We can re-write Eq.~\ref{nspp_relative} as
-\begin{align}
-\begin{split}
-	\frac{dp_i}{dt} &= p_i \left((A \bm{q})_i - \bm{p}^T A \bm{q} \right) \\
-	\frac{dq_i}{dt} &= q_i \left((B \bm{p})_i - \bm{q}^T B \bm{p} \right) \\
-\end{split}
-\end{align}
-or even more compactly as
-\begin{align}
-\begin{cases}
-\frac{d\bm{p}}{dt} &= D(\bm{p}) \left(A \bm{q} - (\bm{p}^T A \bm{q}) \bm{1} \right) \\
-\frac{d\bm{q}}{dt} &= D(\bm{q}) \left(B \bm{p} - (\bm{q}^T B \bm{p}) \bm{1}  \right) \,
-\end{cases}
-\end{align}
-which is Eq.~\ref{matrix_form} in the Main Text.
-
-These expressions are identical to standard bimatrix replicator dynamics \cite{hofbauer1996evolutionary,hofbauer1998evolutionary}.
-Bimatrix games have two strategy sets (here, the $p_i$ and $q_i$), and interactions take place only between strategies from opposite sets.
-The growth rate terms we considered above now have interpretations as payoffs or fitnesses: $\sum_{j} \alpha_{ij} \, q_j = (A \bm{q})_i$ is the payoff for strategy $i$ (an average of payoffs playing against each strategy of the other ``player'', weighted by the frequency of each strategy, $q_j$) and $\sum_{j, k} \alpha_{jk} p_j  q_k = \bm{p}^T A \bm{q}$ is the average payoff across the population of strategies.
-A general bimatrix game may have any nonnegative $B$; our model assumptions lead to the special case where $B$ is diagonal.
-We note that one could easily and plausibly consider an extension of the Bever model where each plant species has some affect on (up to) all $n$ of the soil components.
-Then, our PSF model would be map exactly onto the full space of bimatrix game dynamics (rather than just a subset).
-However, all of the results we consider hold for arbitrary bimatrix games, meaning the same conclusions about the dynamics of Eqs.~\ref{nspp_relative}-\ref{matrix_form} would apply to this extended model, as well.
-
-We note two useful properties of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, as they will be important for the analysis that follows.
-First, we have the constraint $\sum_i p_i = \sum_i q_i = 1$ at every point in time.
-Second, the dynamics are completely unchanged by adding a constant to any \emph{column} of the parameter matrices $A$ or $B$.
-The first fact is a direct consequence of our definition for $p_i$ and $q_i$; the second can easily be shown.
-Suppose we have added a constant $c$ to each element in the $l$th column of $A$.
-Then
-
-\begin{align}
-\begin{split}
-\frac{dp_i}{dt} &= p_i \left(\sum_{j} \alpha_{ij} \, q_j + c \, q_l - \sum_{j, k} \alpha_{jk} \, p_j \, q_k - \sum_{j} c \, p_j \, q_l \right)\\
-&= p_i \left(\sum_{j} \alpha_{ij} \, q_j + c \, q_l - \sum_{j, k} \alpha_{jk} \, p_j \, q_k - c \, q_l \right) \\
-&= p_i \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right)
-\end{split}
-\end{align}
-which is precisely the differential equation we obtained prior to adding $c$.
-Clearly the trajectories of both systems (with and without the column shift) must be identical.
-The same considerations apply for the matrix $B$.
-Intuitively, this property reflects the fact that we are always subtracting the average payoff, and so any change to the payoffs that benefits (or harms) each species equally is ``invisible'' to the dynamics.
-
-In the remaining sections, we outline the main behaviors of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, especially with regard to coexistence.
-We closely follow the treatment by Hofbauer and Sigmund \cite{hofbauer1998evolutionary}, and urge interested readers to consult this excellent introduction (see especially chapters 10 and 11).
-Here, we reproduce or sketch the essential details needed to justify the results in the Main Text.
-
-\subsection{Coexistence equilibrium}
-
-Written in matrix form, it is easy to see that the model admits a unique fixed point where all species are present at non-zero frequency.
-This fixed point, $(\bm{p}^\star, \bm{q}^\star)$, must take the form $(c_1 B^{-1} \bm{1}, c_2 A^{-1} \bm{1})$ for some undetermined constants $c_1$ and $c_2$.
-Substituting this ansatz into the growth rates in Eq.~\ref{matrix_form} and equating them to zero, we have
-
-\begin{align}
-	\begin{split}
-	A \bm{q}^\star - ((\bm{p}^\star)^T A \bm{q}^\star) \bm{1} = c_2 A A^{-1} \bm{1} - (c_1 c_2 \bm{1}^T  (B^{-1})^T A A^{-1} \bm{1}) \bm{1} &= c_2 (1 - c_1 \bm{1}^T  (B^{-1})^T \bm{1}) \bm{1} = 0 \\
-	B \bm{p}^\star - ((\bm{q}^\star)^T B \bm{p}^\star) \bm{1} = c_1 B B^{-1} \bm{1} - (c_1 c_2 \bm{1}^T  (A^{-1})^T B B^{-1} \bm{1}) \bm{1} &= c_1 (1 - c_2 \bm{1}^T  (A^{-1})^T \bm{1}) \bm{1} = 0
-	\end{split}
-\end{align}
-From the final two equations, it is clear that $c_1 = \frac{1}{\bm{1}^T  (B^{-1})^T \bm{1}} = \frac{1}{\bm{1}^T  B^{-1} \bm{1}}$ and $c_2 = \frac{1}{\bm{1}^T  (A^{-1})^T \bm{1}} = \frac{1}{\bm{1}^T  A^{-1} \bm{1}}$.
-
-These rescaling factors make intuitive sense, as they ensure that $\sum_i p^\star_i = \sum_i q^\star_i = 1$, consistent with their definition as frequencies.
-
-Describing these equilibrium frequencies in terms of the parameters is a difficult problem that has received significant attention elsewhere \cite{eppinga2018frequency,mack2019plant,saavedra2017structural,servan2018coexistence,pettersson2020stability,saavedra2021feasibility}.
-In particular, one is usually interested in identifying whether all of the frequencies are nonnegative (such a fixed point is said to be feasible).
-The existence of a feasible fixed point is a requirement for the model to exhibit permanence, meaning that no species go extinct or grow to infinity.
-Throughout our analysis, we assume the existence of a feasible fixed point; considering the question of feasibility simultaneously would only make coexistence less likely in each case.
-
-\subsection{Local stability analysis}
-
-Perturbations around the coexistence equilibrium are constrained to respect the conditions $\sum_i p_i = \sum_i q_i = 1$.
-For this reason, it is convenient to remove these constraints before performing a local stability analysis.
-As in the two species case \cite{bever1997incorporating}, this can be done by eliminating the $n$th species and soil component, which leaves us with a $2n - 2$ dimensional system with no special constraints.
-
-We use $p_n = 1 - \sum_{i = 1}^{n-1} p_i \equiv f(\bm{p})$ and $q_n = 1 - \sum_{i=1}^{n-1} q_i \equiv g(\bm{q})$ and write these frequencies as functions of the others.
-The reduced dynamics are given by
-
-\begin{align}
-\begin{cases} \label{reduced}
-\frac{dp_i}{dt} &= p_i \left(\sum_{j}^{n-1} \alpha_{ij} \, q_j + \alpha_{in} g(\bm{q}) - \sum_{j, k}^{n-1} \alpha_{jk} \, p_j \, q_k - f(\bm{p}) \sum_{j}^{n-1} \alpha_{nj} q_j - g(\bm{q}) \sum_{j}^{n-1} \alpha_{jn} p_j - \alpha_{nn} f(\bm{p}) g(\bm{q})  \right) \\
-\frac{dq_i}{dt} &= q_i \left(\beta_{i} \, p_i - \sum_{j}^{n-1} \beta_{j} \, p_j \, q_j  - \beta_n f(\bm{p}) g(\bm{q})  \right) \, , \quad  i = 1, \dots n-1
-\end{cases}
-\end{align}
-Although these equations appear more complex, it is now straightforward to analyze the local stability of the coexistence equilibrium.
-
-The elements of the community matrix (the Jacobian evaluated at the coexistence equilibrium) are easily computed from Eq.~\ref{reduced}.
-First we consider the plant dynamics differentiated with respect to the plant frequencies.
-In these calculations, all frequencies are evaluated at their equilibrium values.
-
-\begin{align} \label{firstblock}
-\begin{split} 
-	\frac{\partial}{\partial p_j} \frac{dp_i}{dt} &= p_i \left( - \sum_{k}^{n-1} \alpha_{jk} \, q_k + \sum_{k}^{n-1} \alpha_{nk} q_k - \alpha_{jn} g(\bm{q}) + \alpha_{nn} g(\bm{q}) \right) \\
-	&= p_i \left( - \sum_{k}^{n} \alpha_{jk} \, q_k + \sum_{k}^{n} \alpha_{nk} q_k \right)\\
-	&= 0
-\end{split}
-\end{align}
-Here, we have used the fact that $A \bm{q^*} \propto \bm{1}$.
-Notice that, because the factors in parentheses in Eq.~\ref{reduced} are zero at equilibrium, these community matrix calculations are valid even for $i = j$.
-
-The other elements are computed similarly:
-
-\begin{align}
-\begin{split} 
-\frac{\partial}{\partial q_j} \frac{dq_i}{dt} &= q_i \left( - \beta_i q_i + \beta_n f(\bm{p}) \right) \\
-&= 0
-\end{split}
-\end{align}
-
-\begin{align}
-\begin{split} 
-\frac{\partial}{\partial q_j} \frac{dp_i}{dt} &= p_i \left(\alpha_{ij} - \alpha_{in} - \sum_{k}^{n-1} \alpha_{kj} \, p_k - \alpha_{nj} \, f(\bm{p}) + \sum_{k}^{n-1} \alpha_{kn} p_k + \alpha_{nn} f(\bm{p}) \right) \\
-&= p_i(\alpha_{ij} - \alpha_{in})
-\end{split}
-\end{align}
-
-\begin{align} \label{lastblock}
-\begin{split} 
-\frac{\partial}{\partial p_j} \frac{dq_i}{dt} &= 
-\begin{cases}
-q_i \left(\beta_{i} - \beta_{n} \right) \, , \quad i = j\\
-0 \, , \quad i \neq j
-\end{cases} \\
-\end{split}
-\end{align}
-
-From these calculations, it is apparent that the trace of the community matrix, given by $\sum_i^{n-1} \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_j^{n-1} \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$ , is zero.
-The trace of a square matrix is equal to the sum of its eigenvalues \cite{horn2012matrix}, so the eigenvalues of the community matrix must include either (i) a mix of positive and negative real parts or (ii) only purely imaginary values.
-In the first case, the coexistence equilibrium is locally unstable, because at least one eigenvalue has positive real part.
-In the second case, the coexistence equilibrium is a neutrally or marginally stable.
-These two possibilities exclude locally stable equilibria.
-In this respect, the behavior of the two-species model is the generic behavior of the generalized $n$-species model.
-
-\subsection{Divergence and attractors}
-
-We can extend this picture beyond a local neighborhood of the coexistence equilibrium by considering the divergence of the vector field associated with Eqs.~\ref{nspp_relative}-\ref{matrix_form}.
-The divergence, defined by $\sum_i \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_i \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$, measures the outgoing flux around a given point.
-It can be shown (see \cite{eshel1983coevolutionary} and \cite{hofbauer1998evolutionary}) that up to a change in velocity (i.e.
-rescaling time by a positive factor), the vector field corresponding to any bimatrix game dynamics has zero divergence everywhere in the interior of the positive orthant (i.e.
-where $p_i, q_i > 0$ for all $i$).
-
-The divergence theorem \cite{arfken1985mathematical} equates the integral of the divergence of a vector field over some $n$-dimensional region to the net flux over the boundary of the region.
-For a vector field with zero divergence, this implies that every closed surface has zero net flux.
-As a consequence, such \emph{divergence-free} vector fields cannot have attractors, or subsets of phase space toward which trajectories of the corresponding dynamical system tend to evolve.
-If an attractor existed, one could define a surface enclosing it sufficiently tightly, and the net flux over this surface would be negative (as trajectories enter, but do not exit, this region).
-But this would present a contradiction, and so we conclude that there can be no attractors, such as limit cycles, for the dynamics.
-
-For our model, these facts mean that attractors can only exist on the boundary of the phase space.
-Because each boundary face for the $n$-dimensional system is another bimatrix replicator system on $n-2$ dimensions, the same logic applies, and the only possible attractors are points where a single species (and corresponding soil component) is present \cite{hofbauer1998evolutionary}.
-States with multiple species present are never attractive.
-This leaves neutrally-stable oscillations as the only potential form of species coexistence.
-
-\subsection{Rescaled zero-sum games are neutrally stable}
-
-In the context of bimatrix games, a zero-sum game is one where $A = -B^T$.
-A rescaled zero-sum game is one where there exist constants $\gamma_i, \delta_j$ and $c > 0$ such that $a_{ij} + \delta_j = -c b_{ji} + \gamma_i$ for all $i$ and $j$ (here, we understand $A = (a_{ij}), B = (b_{ij})$) \cite{hofbauer1998evolutionary}.
-Any rescaled zero-sum game can be turned into a zero-sum game by adding constants (in particular, $-\delta_j$ and $-\gamma_j$) to each column of $A$ and $B$, and then multiplying $B$ by a positive constant $1 / c$.
-As such, the dynamics of a rescaled zero-sum game and its corresponding zero-sum game are the same up to a rescaling of time.
-
-If a rescaled zero-sum game has a feasible coexistence equilibrium, this equilibrium is neutrally stable.
-We can see this by considering the associated community matrix.
-First, we assume without loss of generality that $A = -c B^T$ (otherwise, we shift columns to obtain this form, without altering the dynamics in the process) Now we add the column-constant matrix $\frac{1}{c} \bm{b}_n \bm{1}^T$ to $A$ and $c \bm{a}_n \bm{1}^T$ to $B$, where $\bm{a}_n$ ($\bm{b}_n$) denotes the $n$th column of $A$ ($B$).
-Again, the dynamics, including both equilibrium values and stability properties, are unchanged by this operation.
-From Eqs.~\ref{firstblock}-\ref{lastblock}, we see that the community matrix, $J$, of the resulting system is given by
-
-\begin{equation}
-\begin{pmatrix}
-	0 && D(\bm{p}^\star) (\bar{A} + \frac{1}{c} \bm{b}_n \bm{1}^T - \bm{1} \bm{a}_n^T) \\
-	D(\bm{q}^\star) (\bar{B} + c \bm{a}_n \bm{1}^T - \bm{1} \bm{b}_n^T) && 0
-\end{pmatrix}
-\end{equation}
-where $\bar{A}$ ($\bar{B}$) denotes the $(n-1) \times (n-1)$ submatrix of $A$ ($B$) obtained by dropping the $n$th row and column.
-Finally, we consider the similarity transform $P^{-1} J P$, defined by the change of basis matrix
-
-\begin{equation}
-P = \begin{pmatrix}
-	\sqrt{c} D(\bm{p}^\star)^{1/2} && 0 \\
-	0 && D(\bm{q}^\star)^{1/2}
-\end{pmatrix} \, .
-\end{equation}
-The resulting matrix, $J'$, which shares the same eigenvalues as $J$ \cite{horn2012matrix}, is given by
-
-\begin{equation}
-\begin{pmatrix}
-0 && \sqrt{c} D(\bm{p}^\star)^{1/2} (\bar{A} + \frac{1}{c} \bm{b}_n \bm{1}^T - \bm{1} \bm{a}_n^T) D(\bm{q}^\star)^{1/2} \\
-\sqrt{c} D(\bm{q}^\star)^{1/2} (-\bar{A^T} + \bm{a}_n \bm{1}^T - \frac{1}{c} \bm{1} \bm{b}_n^T) D(\bm{p}^\star)^{1/2} 
-\end{pmatrix}
-\end{equation}
-which is a skew-symmetric matrix.
-Every eigenvalue of a skew-symmetric matrix must have zero real part \cite{horn2012matrix}.
-Thus, the eigenvalues of $J$, the community matrix, have zero real part, and the coexistence equilibrium of our original system is neutrally stable.
-
-Here, we have outlined a proof that applies to all rescaled zero-sum games.
-When $B$ is a diagonal matrix, as in our model of PSFs, the condition for $A$ and $B$ to constitute a rescaled zero-sum game reduces to the condition given in the Main Text.
-
-Rescaled zero-sum games are the only bimatrix games known to produce neutrally stable oscillations.
-It is a long-standing conjecture that no other bimatrix games have this property \cite{hofbauer1996evolutionary,hofbauer1998evolutionary,hofbauer2011deterministic}.
-
-\subsection{Two-species bimatrix games}
-
-For $n > 2$, the rescaled zero-sum game condition is very stringent -- it places exacting equality constraints on the elements of $A$ and $B$.
-However, for $n = 2$, every bimatrix game satisfies $a_{ij} + \delta_j = -c b_{ji} + \gamma_i$ for some $c$ potentially positive (in which case we have a rescaled zero-sum game) or negative (in which case the game is called a \emph{partnership game}, and the coexistence equilibrium is unstable) \cite{hofbauer1998evolutionary}.
-Thus, neutral oscillations arise whenever $c > 0$.
-
-To see that this is true, we first suppose that $A$ and $B$ have the form
-
-\begin{equation}
-	A = \begin{pmatrix}
-	0 && a_1 \\
-	a_2 && 0
-	\end{pmatrix}
-	\quad \quad 
-	B = \begin{pmatrix}
-	0 && b_1 \\
-	b_2 && 0 
-	\end{pmatrix} \, . 
-\end{equation}
-If this is not the case, we can use constant column shifts to arrive at this form (e.g., in general, $a_1 = a_{12} - a_{22}$).
-Now consider the constants $c = -\frac{a_1 + a_2}{b_1 + b_2}$ and $\gamma_1 = \delta_1 = a_1 + c b_2$ and $\gamma_2 = \delta_2 = 0$.
-Examining the equation $a_{ij} + \delta_j - \gamma_i = -c b_{ji}$ for each $i$ and $j$, one verifies
-
-\begin{align}
-	\begin{split}
-	0 + \gamma_1 - \delta_1  &= 0 \\
-	a_1 + \gamma_2 - \delta_1 &= -c b_2 \\
-	a_2 + \gamma_1 - \delta_2 = -c (b_1 + b_2 - b2) &= -c b_1 \\
-	0  &= 0
-	\end{split}
-\end{align}
-and so the parameters $A$ and $B$ always constitute a rescaled zero-sum or partnership game.
-In the particular case of our model, $a_1 + a_2 = -\alpha_{11} + \alpha_{21} + \alpha_{12} - \alpha_{22} = -I_s$ and $b_1 + b_2 = -\beta_1 - \beta_2$.
-$c$ is positive (as needed for cycles) when these signs disagree; since $b_1 + b_2 = -\beta_1 - \beta_2$ is always negative, $a_1 + a_2$ must be positive, meaning $I_s < 0$, as found by Bever et al.
-
-\subsection{Constant of motion}
-
-When $A$ and $B$ satisfy the rescaled zero-sum game condition, the function
-
-\begin{equation}
-	H(\bm{p}, \bm{q}) = \sum_i p_i^\star \log p_i + c \sum_j q_j^\star \log q_j 
-\end{equation}
-is a constant of motion for the dynamics \cite{hofbauer1998evolutionary}.
-As above, we suppose that $A = -c B^T$, and shift the columns of each matrix as needed if this is not the case.
-Then consider the time derivative
-
-\begin{align}
-	\begin{split}
-	\frac{dH}{dt} &= \sum_i p_i^\star \frac{1}{p_i} \frac{dp_i}{dt} + c \sum_j q_j^\star \frac{1}{q_i} \frac{dq_i}{dt} \\
-	&= \sum_i p_i^\star \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right) + c \sum_j q_j^\star \left( \beta_{i} \, p_i - \sum_{j} \beta_{j} \, p_j \, q_j \right) \\
-	&= \sum_{i, j} \alpha_{ij} \, p_i^\star \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k + c \sum_{i} \beta_{i} \, q_i^\star \, p_i - c \sum_{j} \beta_{j} \, p_j \, q_j \\
-	&= \sum_{i, j} \alpha_{ij} \, (p_i^\star - p_i) \, q_j + c \sum_{i} \beta_{i} \, (q_i^\star - q_i) \, p_i 
-	\intertext{Now, because $A = -c B^T$, we have}
-	&= c \sum_i \beta_i \left( -(p_i^\star - p_i) \, q_i + (q_i^\star - q_i) \, p_i \right)\\
-	&= c \sum_i \beta_i \left( -p_i^\star \, q_i + q_i^\star \, p_i \right)
-	\intertext{and because $q_i^* = p_i^* = \frac{Z}{\beta_{i}}$, with $Z$ the normalizing constant,}
-	&= c \, Z \, \sum_i (-q_i + p_i) \\
-	&= 0
-	\end{split}
-\end{align}
-In the last line, we use the fact that both sets of frequencies always sum to one.
-
-Each orbit remains in the level set defined by the initial conditions,$(\bm{p}_0, \bm{q}_0)$:
-
-\begin{equation}
-H(\bm{p}_0, \bm{q}_0) = \sum_i p_i^\star \log p_i + c \sum_j q_j^\star \log q_j 
-\end{equation}
-For the two-species model studied by Bever et al., these level sets precisely define the trajectories in the $(p, q)$ phase plane.
 

### 2021-10-26 paper/supp_methods.tex
updated ms with cover letter and title page

\documentclass[11pt]{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry} % page settings
\usepackage{float}
\usepackage{natbib}

\setcitestyle{round,aysep={}}

\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}

\usepackage{amsmath}
\usepackage{amsthm, thmtools}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}   
\usepackage{enumitem}
\usepackage{listings}
\usepackage[hidelinks]{hyperref}
%\numberwithin{equation}

\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows}
\usepackage[makeroom,samesize]{cancel}

\usepackage{
nameref,
%\nameref
hyperref,
%\autoref
% n.b. \Autoref is defined by thmtools
cleveref,
% \cref
% n.b. cleveref after! hyperref
}

\usepackage{setspace}
\usepackage{lineno}
\title{Supplemental Methods: No robust coexistence in a canonical model of plant-soil feedbacks}

\date{}
\author{Zachary R. Miller$^{1*}$, Pablo Lech\'{o}n$^{1}$, and Stefano Allesina$^{1,2}$ \\
	\\
	\normalsize{$^{1}$Department of Ecology \& Evolution, University of Chicago, Chicago, IL, USA}\\
	\normalsize{$^{2}$Northwestern Institute on Complex Systems, Evanston, IL, USA}\\
	\\
	\normalsize{*corresponding author e-mail: zachmiller@uchicago.edu}\\
}

\begin{document}
\maketitle
\setstretch{1.5}
\linenumbers

\section{Model derivation}

As described in the Main Text, we begin with the system

\begin{align}
\begin{cases}
\frac{dx_i}{dt} &= x_i \left(\sum_{j} \alpha_{ij} \, q_j \right) \, , \quad  i = 1, \dots n \\
\frac{dy_i}{dt} &= y_i \left( \beta_i \, p_i \right)
\end{cases}
\end{align}
governing the time-evolution of plant abundances $x_i$ and soil components $y_i$, where $p_i = x_i / \sum_j x_j$, $q_i = y_i / \sum_j y_j$, and Greek letters denote nonnegative parameters. These equations capture the assumptions outlined by Bever et al. \citep{bever1997incorporating} for two species and extend them straightforwardly to any $n$ species. Following the approach of Bever et al. for two species (and consistent with other generalizations of this model \citep{kulmatiski2008plant,eppinga2018frequency}), we derive dynamics for the $p_i$ by applying the chain rule:

\begin{align} \label{derive_relative}
\begin{split}
	\frac{dp_i}{dt} &= \frac{d}{dt} \frac{x_i}{\sum x_j}\\  
	&= \frac{1}{\sum_j x_j} \frac{dx_i}{dt} - \frac{x_i}{(\sum_j x_j)^2} \sum_j \frac{dx_j}{dt} \\
	&= \frac{x_i}{\sum_j x_j} \left( \sum_{j} \alpha_{ij} \, q_j \right) - \frac{x_i}{\sum_j x_j} \left(\sum_j \frac{x_j}{\sum_k x_k} \sum_{l} \alpha_{jl} \, q_l \right)\\
	&= p_i \left(\sum_{j} \alpha_{ij} \, q_j -  \sum_{j, k} \alpha_{jk} p_j  q_k \right) \, .
\end{split}
\end{align}
This last expression is identical to the first line of Eq.~\ref{nspp_relative} in the Main Text. The dynamics for $q_i$ can be derived in exactly the same way (using the definitions $\beta_{ii} = \beta_i$ and $\beta_{ij} = 0$). The two terms of each per capita growth rate in Eq.~\ref{nspp_relative} have natural interpretations in the language and notation of linear algebra: $\sum_{j} \alpha_{ij} \, q_j$ is the $i$th component of the matrix-vector product $A \bm{q}$ and $\sum_{j, k} \alpha_{jk} p_j  q_k$ is the bilinear form $\bm{p}^T A \bm{q}$. Here, $A$ (and $B$) is an $n \times n$ matrix and $\bm{p}$ ($\bm{q}$) is a vector of length $n$, as described in the Main Text. We can re-write Eq.~\ref{nspp_relative} as
\begin{align}
\begin{split}
	\frac{dp_i}{dt} &= p_i \left((A \bm{q})_i - \bm{p}^T A \bm{q} \right) \\
	\frac{dq_i}{dt} &= q_i \left((B \bm{p})_i - \bm{q}^T B \bm{p} \right) \\
\end{split}
\end{align}
or even more compactly as
\begin{align}
\begin{cases}
\frac{d\bm{p}}{dt} &= D(\bm{p}) \left(A \bm{q} - (\bm{p}^T A \bm{q}) \bm{1} \right) \\
\frac{d\bm{q}}{dt} &= D(\bm{q}) \left(B \bm{p} - (\bm{q}^T B \bm{p}) \bm{1}  \right) \,
\end{cases}
\end{align}
which is Eq.~\ref{matrix_form} in the Main Text.

These expressions are identical to standard bimatrix replicator dynamics \citep{hofbauer1996evolutionary,hofbauer1998evolutionary}. Bimatrix games have two strategy sets (here, the $p_i$ and $q_i$), and interactions take place only between strategies from opposite sets. The growth rate terms we considered above now have interpretations as payoffs or fitnesses: $\sum_{j} \alpha_{ij} \, q_j = (A \bm{q})_i$ is the payoff for strategy $i$ (an average of payoffs playing against each strategy of the other ``player'', weighted by the frequency of each strategy, $q_j$) and $\sum_{j, k} \alpha_{jk} p_j  q_k = \bm{p}^T A \bm{q}$ is the average payoff across the population of strategies. A general bimatrix game may have any nonnegative $B$; our model assumptions lead to the special case where $B$ is diagonal. We note that one could easily and plausibly consider an extension of the Bever model where each plant species has some effect on (up to) all $n$ of the soil components. Then, our PSF model would be map exactly onto the full space of bimatrix game dynamics (rather than just a subset). However, all of the results we consider hold for arbitrary bimatrix games, meaning the same conclusions about the dynamics of Eqs.~\ref{nspp_relative}-\ref{matrix_form} would apply to this extended model, as well.

We note two useful properties of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, as they will be important for the analysis that follows. First, we have the constraint $\sum_i p_i = \sum_i q_i = 1$ at every point in time. Second, the dynamics are completely unchanged by adding a constant to any \emph{column} of the parameter matrices $A$ or $B$. The first fact is a direct consequence of our definition for $p_i$ and $q_i$; the second can easily be shown. Suppose we have added a constant $c$ to each element in the $l$th column of $A$. Then

\begin{align}
\begin{split}
\frac{dp_i}{dt} &= p_i \left(\sum_{j} \alpha_{ij} \, q_j + c \, q_l - \sum_{j, k} \alpha_{jk} \, p_j \, q_k - \sum_{j} c \, p_j \, q_l \right)\\
&= p_i \left(\sum_{j} \alpha_{ij} \, q_j + c \, q_l - \sum_{j, k} \alpha_{jk} \, p_j \, q_k - c \, q_l \right) \\
&= p_i \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right)
\end{split}
\end{align}
which is precisely the differential equation we obtained prior to adding $c$. Clearly the trajectories of both systems (with and without the column shift) must be identical. The same considerations apply for the matrix $B$. Intuitively, this property reflects the fact that we are always subtracting the average payoff, and so any change to the payoffs that benefits (or harms) each species equally is ``invisible'' to the dynamics.

In the remaining sections, we outline the main behaviors of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, especially with regard to coexistence. We closely follow the treatment by Hofbauer and Sigmund \citep{hofbauer1998evolutionary}, and urge interested readers to consult this excellent introduction (see especially chapters 10 and 11). Here, we reproduce or sketch the essential details needed to justify the results in the Main Text.

\section{Coexistence equilibrium}

Written in matrix form, it is easy to see that the model admits a unique fixed point where all species are present at non-zero frequency. This fixed point, $(\bm{p}^\star, \bm{q}^\star)$, must take the form $(c_1 B^{-1} \bm{1}, c_2 A^{-1} \bm{1})$ for some undetermined constants $c_1$ and $c_2$. Substituting this ansatz into the growth rates in Eq.~\ref{matrix_form} and equating them to zero, we have

\begin{align}
	\begin{split}
	A \bm{q}^\star - ((\bm{p}^\star)^T A \bm{q}^\star) \bm{1} = c_2 A A^{-1} \bm{1} - (c_1 c_2 \bm{1}^T  (B^{-1})^T A A^{-1} \bm{1}) \bm{1} &= c_2 (1 - c_1 \bm{1}^T  (B^{-1})^T \bm{1}) \bm{1} = 0 \\
	B \bm{p}^\star - ((\bm{q}^\star)^T B \bm{p}^\star) \bm{1} = c_1 B B^{-1} \bm{1} - (c_1 c_2 \bm{1}^T  (A^{-1})^T B B^{-1} \bm{1}) \bm{1} &= c_1 (1 - c_2 \bm{1}^T  (A^{-1})^T \bm{1}) \bm{1} = 0
	\end{split}
\end{align}
From the final two equations, it is clear that $c_1 = \frac{1}{\bm{1}^T  (B^{-1})^T \bm{1}} = \frac{1}{\bm{1}^T  B^{-1} \bm{1}}$ and $c_2 = \frac{1}{\bm{1}^T  (A^{-1})^T \bm{1}} = \frac{1}{\bm{1}^T  A^{-1} \bm{1}}$.

These rescaling factors make intuitive sense, as they ensure that $\sum_i p^\star_i = \sum_i q^\star_i = 1$, consistent with their definition as frequencies.

Describing these equilibrium frequencies in terms of the parameters is a difficult problem that has received significant attention elsewhere \citep{eppinga2018frequency,mack2019plant,saavedra2017structural,servan2018coexistence,pettersson2020stability,saavedra2021feasibility}. In particular, one is usually interested in identifying whether all of the frequencies are nonnegative (such a fixed point is said to be feasible). The existence of a feasible fixed point is a requirement for the model to exhibit permanence, meaning that no species go extinct or grow to infinity. Throughout our analysis, we assume the existence of a feasible fixed point; considering the question of feasibility simultaneously would only make coexistence less likely in each case.

\section{Local stability analysis}

Perturbations around the coexistence equilibrium are constrained to respect the conditions $\sum_i p_i = \sum_i q_i = 1$. For this reason, it is convenient to remove these constraints before performing a local stability analysis. As in the two species case \citep{bever1997incorporating}, this can be done by eliminating the $n$th species and soil component, which leaves us with a $2n - 2$ dimensional system with no special constraints.

We use $p_n = 1 - \sum_{i = 1}^{n-1} p_i \equiv f(\bm{p})$ and $q_n = 1 - \sum_{i=1}^{n-1} q_i \equiv g(\bm{q})$ and write these frequencies as functions of the others. The reduced dynamics are given by

\begin{align}
\begin{cases} \label{reduced}
\frac{dp_i}{dt} &= p_i \left(\sum_{j}^{n-1} \alpha_{ij} \, q_j + \alpha_{in} g(\bm{q}) - \sum_{j, k}^{n-1} \alpha_{jk} \, p_j \, q_k - f(\bm{p}) \sum_{j}^{n-1} \alpha_{nj} q_j - g(\bm{q}) \sum_{j}^{n-1} \alpha_{jn} p_j - \alpha_{nn} f(\bm{p}) g(\bm{q})  \right) \\
\frac{dq_i}{dt} &= q_i \left(\beta_{i} \, p_i - \sum_{j}^{n-1} \beta_{j} \, p_j \, q_j  - \beta_n f(\bm{p}) g(\bm{q})  \right) \, , \quad  i = 1, \dots n-1
\end{cases}
\end{align}
Although these equations appear more complex, it is now straightforward to analyze the local stability of the coexistence equilibrium.

The elements of the community matrix (the Jacobian evaluated at the coexistence equilibrium) are easily computed from Eq.~\ref{reduced}. First we consider the plant dynamics differentiated with respect to the plant frequencies. In these calculations, all frequencies are evaluated at their equilibrium values.

\begin{align} \label{firstblock}
\begin{split} 
	\frac{\partial}{\partial p_j} \frac{dp_i}{dt} &= p_i \left( - \sum_{k}^{n-1} \alpha_{jk} \, q_k + \sum_{k}^{n-1} \alpha_{nk} q_k - \alpha_{jn} g(\bm{q}) + \alpha_{nn} g(\bm{q}) \right) \\
	&= p_i \left( - \sum_{k}^{n} \alpha_{jk} \, q_k + \sum_{k}^{n} \alpha_{nk} q_k \right)\\
	&= 0
\end{split}
\end{align}
Here, we have used the fact that $A \bm{q^*} \propto \bm{1}$. Notice that, because the factors in parentheses in Eq.~\ref{reduced} are zero at equilibrium, these community matrix calculations are valid even for $i = j$.

The other elements are computed similarly:

\begin{align}
\begin{split} 
\frac{\partial}{\partial q_j} \frac{dq_i}{dt} &= q_i \left( - \beta_i q_i + \beta_n f(\bm{p}) \right) \\
&= 0
\end{split}
\end{align}

\begin{align}
\begin{split} 
\frac{\partial}{\partial q_j} \frac{dp_i}{dt} &= p_i \left(\alpha_{ij} - \alpha_{in} - \sum_{k}^{n-1} \alpha_{kj} \, p_k - \alpha_{nj} \, f(\bm{p}) + \sum_{k}^{n-1} \alpha_{kn} p_k + \alpha_{nn} f(\bm{p}) \right) \\
&= p_i(\alpha_{ij} - \alpha_{in})
\end{split}
\end{align}

\begin{align} \label{lastblock}
\begin{split} 
\frac{\partial}{\partial p_j} \frac{dq_i}{dt} &= 
\begin{cases}
q_i \, \beta_{i} \, , \quad i = j\\
0 \, , \quad i \neq j
\end{cases} \\
\end{split}
\end{align}

From these calculations, it is apparent that the trace of the community matrix, given by $\sum_i^{n-1} \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_j^{n-1} \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$ , is zero. The trace of a square matrix is equal to the sum of its eigenvalues \citep{horn2012matrix}, so the eigenvalues of the community matrix must include either (i) a mix of positive and negative real parts or (ii) only purely imaginary values. In the first case, the coexistence equilibrium is locally unstable, because at least one eigenvalue has positive real part. In the second case, the coexistence equilibrium is a neutrally or marginally stable. These two possibilities exclude locally stable equilibria. In this respect, the behavior of the two-species model is the generic behavior of the generalized $n$-species model.

\section{Zero divergence implies no attractors}

We can extend this picture beyond a local neighborhood of the coexistence equilibrium by considering the divergence of the vector field associated with Eqs.~\ref{nspp_relative}-\ref{matrix_form}. The divergence, defined by $\sum_i \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_i \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$, measures the outgoing flux around a given point. It can be shown (see \citep{eshel1983coevolutionary} and \citep{hofbauer1998evolutionary}) that up to a change in velocity (i.e. rescaling time by a positive factor), the vector field corresponding to any bimatrix game dynamics has zero divergence everywhere in the interior of the positive orthant (i.e. where $p_i, q_i > 0$ for all $i$).

The divergence theorem \citep{arfken1985mathematical} equates the integral of the divergence of a vector field over some $n$-dimensional region to the net flux over the boundary of the region. For a vector field with zero divergence, this implies that every closed surface has zero net flux. As a consequence, such \emph{divergence-free} vector fields cannot have attractors, or subsets of phase space toward which trajectories of the corresponding dynamical system tend to evolve. If an attractor existed, one could define a surface enclosing it sufficiently tightly, and the net flux over this surface would be negative (as trajectories enter, but do not exit, this region). But this would present a contradiction, and so we conclude that there can be no attractors, such as limit cycles, for the dynamics.

For our model, these facts mean that attractors can only exist on the boundary of the phase space. Because each boundary face for the $n$-dimensional system is another bimatrix replicator system on $n-2$ dimensions, the same logic applies, and the only possible attractors are points where a single species (and corresponding soil component) is present \citep{hofbauer1998evolutionary}. States with multiple species present are never attractive. This leaves neutrally-stable oscillations as the only potential form of species coexistence.

\section{Rescaled zero-sum games are neutrally stable}

In the context of bimatrix games, a zero-sum game is one where $A = -B^T$. A rescaled zero-sum game is one where there exist constants $\gamma_i, \delta_j$ and $c > 0$ such that $a_{ij} + \delta_j = -c b_{ji} + \gamma_i$ for all $i$ and $j$ (here, we understand $A = (a_{ij}), B = (b_{ij})$) \citep{hofbauer1998evolutionary}. Any rescaled zero-sum game can be turned into a zero-sum game by adding constants (in particular, $-\delta_j$ and $-\gamma_j$) to each column of $A$ and $B$, and then multiplying $B$ by a positive constant $1 / c$. As such, the dynamics of a rescaled zero-sum game and its corresponding zero-sum game are the same up to a rescaling of time.

If a rescaled zero-sum game has a feasible coexistence equilibrium, this equilibrium is neutrally stable. We can see this by considering the associated community matrix. First, we assume without loss of generality that $A = -c B^T$ (otherwise, we shift columns to obtain this form, without altering the dynamics in the process) Now we add the column-constant matrix $\frac{1}{c} \bm{b}_n \bm{1}^T$ to $A$ and $c \bm{a}_n \bm{1}^T$ to $B$, where $\bm{a}_n$ ($\bm{b}_n$) denotes the $n$th column of $A$ ($B$). Again, the dynamics, including both equilibrium values and stability properties, are unchanged by this operation. From Eqs.~\ref{firstblock}-\ref{lastblock}, we see that the community matrix, $J$, of the resulting system is given by

\begin{equation}
\begin{pmatrix}
	0 && D(\bm{p}^\star) (\bar{A} + \frac{1}{c} \bm{b}_n \bm{1}^T - \bm{1} \bm{a}_n^T) \\
	D(\bm{q}^\star) (\bar{B} + c \bm{a}_n \bm{1}^T - \bm{1} \bm{b}_n^T) && 0
\end{pmatrix}
\end{equation}
where $\bar{A}$ ($\bar{B}$) denotes the $(n-1) \times (n-1)$ submatrix of $A$ ($B$) obtained by dropping the $n$th row and column. Finally, we consider the similarity transform $P^{-1} J P$, defined by the change of basis matrix

\begin{equation}
P = \begin{pmatrix}
	\sqrt{c} D(\bm{p}^\star)^{1/2} && 0 \\
	0 && D(\bm{q}^\star)^{1/2}
\end{pmatrix} \, .
\end{equation}
The resulting matrix, $J'$, which shares the same eigenvalues as $J$ \citep{horn2012matrix}, is given by

\begin{equation}
\begin{pmatrix}
0 && \sqrt{c} D(\bm{p}^\star)^{1/2} (\bar{A} + \frac{1}{c} \bm{b}_n \bm{1}^T - \bm{1} \bm{a}_n^T) D(\bm{q}^\star)^{1/2} \\
\sqrt{c} D(\bm{q}^\star)^{1/2} (-\bar{A^T} + \bm{a}_n \bm{1}^T - \frac{1}{c} \bm{1} \bm{b}_n^T) D(\bm{p}^\star)^{1/2} && 0
\end{pmatrix}
\end{equation}
which is a skew-symmetric matrix. Every eigenvalue of a skew-symmetric matrix must have zero real part \citep{horn2012matrix}. Thus, the eigenvalues of $J$, the community matrix, have zero real part, and the coexistence equilibrium of our original system is neutrally stable.

Here, we have outlined a proof that applies to all rescaled zero-sum games. When $B$ is a diagonal matrix, as in our model of PSFs, the condition for $A$ and $B$ to constitute a rescaled zero-sum game reduces to the condition given in the Main Text.

Rescaled zero-sum games are the only bimatrix games known to produce neutrally stable oscillations. It is a long-standing conjecture that no other bimatrix games have this property \citep{hofbauer1996evolutionary,hofbauer1998evolutionary,hofbauer2011deterministic}.

\section{Two-species bimatrix games}

For $n > 2$, the rescaled zero-sum game condition is very stringent -- it places exacting equality constraints on the elements of $A$ and $B$. However, for $n = 2$, every bimatrix game satisfies $a_{ij} + \delta_j = -c b_{ji} + \gamma_i$ for some $c$ potentially positive (in which case we have a rescaled zero-sum game) or negative (in which case the game is called a \emph{partnership game}, and the coexistence equilibrium is unstable) \citep{hofbauer1998evolutionary}. Thus, neutral oscillations arise whenever $c > 0$.

To see that this is true, we first suppose that $A$ and $B$ have the form

\begin{equation}
	A = \begin{pmatrix}
	0 && a_1 \\
	a_2 && 0
	\end{pmatrix}
	\quad \quad 
	B = \begin{pmatrix}
	0 && b_1 \\
	b_2 && 0 
	\end{pmatrix} \, . 
\end{equation}
If this is not the case, we can use constant column shifts to arrive at this form (e.g., in general, $a_1 = a_{12} - a_{22}$). Now consider the constants $c = -\frac{a_1 + a_2}{b_1 + b_2}$ and $\gamma_1 = \delta_1 = a_1 + c b_2$ and $\gamma_2 = \delta_2 = 0$. Examining the equation $a_{ij} + \delta_j - \gamma_i = -c b_{ji}$ for each $i$ and $j$, one verifies

\begin{align}
	\begin{split}
	0 + \gamma_1 - \delta_1  &= 0 \\
	a_1 + \gamma_2 - \delta_1 &= -c b_2 \\
	a_2 + \gamma_1 - \delta_2 = -c (b_1 + b_2 - b2) &= -c b_1 \\
	0  &= 0
	\end{split}
\end{align}
and so the parameters $A$ and $B$ always constitute a rescaled zero-sum or partnership game. In the particular case of our model, $a_1 + a_2 = -\alpha_{11} + \alpha_{21} + \alpha_{12} - \alpha_{22} = -I_s$ and $b_1 + b_2 = -\beta_1 - \beta_2$. $c$ is positive (as needed for cycles) when these signs disagree; since $b_1 + b_2 = -\beta_1 - \beta_2$ is always negative, $a_1 + a_2$ must be positive, meaning $I_s < 0$, as found by Bever et al.

\section{Constants of motion}

When $A$ and $B$ satisfy the rescaled zero-sum game condition, the function

\begin{equation}
	H(\bm{p}, \bm{q}) = \sum_i p_i^\star \log p_i + c \sum_j q_j^\star \log q_j 
\end{equation}
is a constant of motion for the dynamics \citep{hofbauer1998evolutionary}. As above, we suppose that $A = -c B^T$, and shift the columns of each matrix as needed if this is not the case. Then consider the time derivative

\begin{align}
	\begin{split}
	\frac{dH}{dt} &= \sum_i p_i^\star \frac{1}{p_i} \frac{dp_i}{dt} + c \sum_j q_j^\star \frac{1}{q_i} \frac{dq_i}{dt} \\
	&= \sum_i p_i^\star \left(\sum_{j} \alpha_{ij} \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k \right) + c \sum_j q_j^\star \left( \beta_{i} \, p_i - \sum_{j} \beta_{j} \, p_j \, q_j \right) \\
	&= \sum_{i, j} \alpha_{ij} \, p_i^\star \, q_j - \sum_{j, k} \alpha_{jk} \, p_j \, q_k + c \sum_{i} \beta_{i} \, q_i^\star \, p_i - c \sum_{j} \beta_{j} \, p_j \, q_j \\
	&= \sum_{i, j} \alpha_{ij} \, (p_i^\star - p_i) \, q_j + c \sum_{i} \beta_{i} \, (q_i^\star - q_i) \, p_i 
	\intertext{Now, because $A = -c B^T$, we have}
	&= c \sum_i \beta_i \left( -(p_i^\star - p_i) \, q_i + (q_i^\star - q_i) \, p_i \right)\\
	&= c \sum_i \beta_i \left( -p_i^\star \, q_i + q_i^\star \, p_i \right)
	\intertext{and because $q_i^* = p_i^* = \frac{Z}{\beta_{i}}$, with $Z$ the normalizing constant,}
	&= c \, Z \, \sum_i (-q_i + p_i) \\
	&= 0
	\end{split}
\end{align}
In the last line, we use the fact that both sets of frequencies always sum to one.

Each orbit remains in the level set defined by the initial conditions,$(\bm{p}_0, \bm{q}_0)$:

\begin{equation}
H(\bm{p}_0, \bm{q}_0) = \sum_i p_i^\star \log p_i + c \sum_j q_j^\star \log q_j 
\end{equation}
For the two-species model studied by Bever et al., these level sets precisely define the trajectories in the $(p, q)$ phase plane.

\bibliographystyle{ecol_let}
\bibliography{references_supp}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-10-26 paper/paper.tex
- Changed many \citep to \citet to avoid double parentheses - added a placeholder for figure 3, and updated references to figs in the text - minimal changes to main text - hard-coded references to equations in the main text across the Supplement - really great work!

@@
 In turn, the fitness of each plant species is impacted by the relative frequency of different soil components.
-Starting from minimal assumptions, Bever et al.
-\citep{bever1997incorporating} derived a set of differential equations to capture these dynamics.
+Starting from minimal assumptions, \citet{bever1997incorporating} derived a set of differential equations to capture these dynamics.
 PSFs can be either positive (fitness of a plant species is increased by its corresponding soil component) or negative (a plant species experiences lower relative fitness in its own soil).
-Bever et al.
-introduced a single quantity to summarize whether community-wide PSFs are positive or negative, and showed that this value characterizes the dynamical behavior of the model.
+Bever \textit{et al.} introduced a single quantity to summarize whether community-wide PSFs are positive or negative, and showed that this value characterizes the dynamical behavior of the model.
 In the original Bever model of two plant species, positive PSFs lead to exclusion of one species, while negative PSFs result in neutral oscillations.
@@
 Subsequent studies have generalized this model to include, for example, more realistic functional forms \citep{umbanhowar2005simple, eppinga2006accumulation}, more explicit representations of the soil community \citep{bever2010rooting}, spatial structure \citep{eppinga2006accumulation,molofsky2002negative,suding2013consequences}, or additional processes such as direct competitive interactions between plants \citep{bever2003soil}.
-However, the original Bever model remains an important touchstone for the theory of PSFs \citep{ke2015incorporating,ke2020effects}, and informs empirical research through the interaction coefficient, $I_s$, derived by Bever et al., which is commonly measured and used to draw conclusions about coexistence in experimental studies.
-Despite the ubiquity of this model, and the fruitful interplay of theory and experiment in the PSF literature, extensions to communities with more than two or three species have appeared only rarely and recently (but see \citep{eppinga2018frequency,mack2019plant}).
+However, the original Bever model remains an important touchstone for the theory of PSFs \citep{ke2015incorporating,ke2020effects}, and informs empirical research through the interaction coefficient, $I_s$, derived by Bever \textit{et al.}, which is commonly measured and used to draw conclusions about coexistence in experimental studies.
+Despite the ubiquity of this model, and the fruitful interplay of theory and experiment in the PSF literature, extensions to communities with more than two or three species have appeared only rarely and recently \citep[but see][]{eppinga2018frequency,mack2019plant}.
 While PSF models motivate hypotheses and conclusions about species-rich natural communities, there is much still unknown about the behavior of these models with natural levels of diversity \citep{van2013plant}.
@@
 
-Inspired by emerging empirical evidence for the important role of PSFs in plant community dynamics and coexistence \citep{van1993plant,bever1994feedback}, Bever et al.
-\citep{bever1997incorporating} introduced a simple mathematical model to investigate their behavior.
+Inspired by emerging empirical evidence for the important role of PSFs in plant community dynamics and coexistence \citep{van1993plant,bever1994feedback}, \citet{bever1997incorporating} introduced a simple mathematical model to investigate their behavior.
 In this model, two plant species, $1$ and $2$, grow exponentially with growth rates determined by the state of the soil biota in the system.
@@
 There is a soil component corresponding to each plant species, which grows exponentially in the presence of its associated plant at a rate $\beta_i$.
-Bever et al.
-set an important precedent by considering dynamics of \emph{relative} abundances in such a system; starting from dynamics of the form
+Bever \textit{et al.} set an important precedent by considering dynamics of \emph{relative} abundances in such a system; starting from dynamics of the form
 
@@
 p_i^\star &= \frac{\beta_j}{\beta_i + \beta_j} \, , \quad  i, j = 1,2 \\
-q_i^\star &= \frac{\alpha_{jj} - \alpha_{ji}}{\alpha_{ii} - \alpha_{ji} - \alpha_{ij} + \alpha_{jj}} \, .
+q_i^\star &= \frac{\alpha_{jj} - \alpha_{ji}}{\alpha_{ii} - \alpha_{ij} + \alpha_{jj} - \alpha_{ji}} \, .
 \end{cases}
@@
 
-A central finding of the analysis by Bever et al.
-was that the denominator of $q_i^*$, which they termed the ``interaction coefficient'', $I_s$ ($= \alpha_{ii} - \alpha_{ji} - \alpha_{ij} + \alpha_{jj}$), controls the model dynamics: When $I_s > 0$, which represents a community with positive feedbacks, the equilibrium in Eq.~\ref{2spp_eq} is unstable, and the two species cannot coexist.
+A central finding of the analysis by Bever \textit{et al.} was that the denominator of $q_i^*$, which they termed the ``interaction coefficient'', $I_s$ ($= \alpha_{ii} - \alpha_{ij} + \alpha_{jj} - \alpha_{ji}$), controls the model dynamics: When $I_s > 0$, which represents a community with positive feedbacks, the equilibrium in Eq.~\ref{2spp_eq} is unstable, and the two species cannot coexist.
 On the other hand, when $I_s < 0$, the equilibrium is neutrally stable, and the dynamics cycle around it, providing a form of non-equilibrium coexistence.
@@
 \end{align}
-From Eq.~\ref{nspp_absolute}, one can derive the $n$-species analogue of Eq.~\ref{2spp_relative},
+From Eq.~\ref{nspp_absolute}, one can derive the $n$-species analogue of Eq.~\ref{2spp_relative} (Fig.~\ref{fig:concept}),
 
@@
 In turn, growth of each plant is governed by the mix of soil components present in the system, with the effect of soil component $j$ on species $i$ quantified by the parameter $\alpha_{ij}$ (black arrows).
-This model is a straightforward extension of the model proposed by Bever et al.
-\citep{bever1997incorporating} to an arbitrary number of species.
+This model is a straightforward extension of the model proposed by \citet{bever1997incorporating} to an arbitrary number of species.
 Note only selected parameter labels are shown for clarity.}
@@
 Much is known about bimatrix game dynamics, and we can draw on this body of knowledge to characterize the behavior of the Bever model with $n$ species.
-Essential mathematical background and details are presented in the Supplemental Methods; for a detailed introduction to bimatrix games, see Hofbauer and Sigmund \citep{hofbauer1998evolutionary}.
+Essential mathematical background and details are presented in the Supplemental Methods; for a detailed introduction to bimatrix games, see \citet{hofbauer1998evolutionary}.
 
@@
 As we have seen, when the community consists of two species, the coexistence equilibrium, if feasible, can be either unstable or neutrally stable.
-In fact, the same is true for the $n$-species extension (and, more generally, for any bimatrix game dynamics \citep{eshel1983coevolutionary,selten1988note,hofbauer1998evolutionary}).
+In fact, the same is true for the $n$-species extension  \citep[and, more generally, for any bimatrix game dynamics,][]{eshel1983coevolutionary,selten1988note,hofbauer1998evolutionary}.
 This can be established using straightforward local stability analysis, after accounting for the relative abundance constraints, which imply $p_n = 1 - \sum_{i = 1}^{n-1} p_i$ and $q_n = 1 - \sum_{i = 1}^{n-1} q_i$.
@@
 
-Another notable property of bimatrix game dynamics is that the vector field defined by the model equations is divergence-free or incompressible (see \citep{hofbauer1998evolutionary} for a proof).
+Another notable property of bimatrix game dynamics is that the vector field defined by the model equations is divergence-free or incompressible \citep[see][for a proof]{hofbauer1998evolutionary}.
 The divergence theorem from vector calculus \citep{arfken1985mathematical} then tells us that Eq.~\ref{matrix_form} cannot have any attractors -- that is, regions of the phase space that ``pull in'' trajectories -- with multiple species.
@@
 
-Based on the local stability properties of the coexistence equilibrium, Bever et al.
-concluded that such neutral cycles arise for two species when $\alpha_{11} < \alpha_{21}$ and $\alpha_{22} < \alpha_{12}$.
+Based on the local stability properties of the coexistence equilibrium, Bever \textit{et al.} concluded that such neutral cycles arise for two species when $\alpha_{11} < \alpha_{21}$ and $\alpha_{22} < \alpha_{12}$.
 The equivalence between their model and a bimatrix game with two strategies allows us to give a fuller picture of these cycles.
@@
 Because neutral cycles provide the only possible form of coexistence in this model, a key question becomes whether and when neutral cycles with $n$ species can arise.
-Do the ``negative feedback'' conditions identified by Bever et al.
-generalize in richer communities?
+Do the ``negative feedback'' conditions identified by Bever \textit{et al.} generalize in richer communities?
 Indeed, they do; however, for more than two species, these conditions are very severe.
-The model in Eq.~\ref{matrix_form} supports oscillations with $n$ species -- for any $n$ -- if matrices $A$ and $B$ satisfy a precise relationship (Fig ???; see Supplemental Methods for details).
+The model in Eq.~\ref{matrix_form} supports oscillations with $n$ species -- for any $n$ -- if matrices $A$ and $B$ satisfy a precise relationship (Fig.~\ref{fig:fragile}; see Supplemental Methods for details).
 In particular, the model parameters must satisfy the conditions $\alpha_{ij} = \gamma_i + \delta_j$ for some constants $\gamma_i, \delta_i$ in $i = 1, \dots, n$ (when $i \neq j$), and $\alpha_{ii} = \gamma_i + \delta_i- c \beta_i$ (where $c$ is a positive constant independent of $i$).
@@
 But the parameters of the model are constrained so strongly that we never expect to observe cycles with more than two species in practice.
-When $n > 2$, a great deal of fine-tuning is necessary to satisfy the zero-sum game condition; the probability that random parameters will be suitable is infinitesimally small.
-We confirm this numerically with simulations shown in Fig ???.
-Although $n$-species cycles are clearly possible (as in Fig ???), for parameters drawn independently at random, communities always collapse to zero, one, or two species, regardless of the initial richness.
+When $n > 2$, a great deal of fine-tuning is necessary to satisfy the conditions yielding a rescaled zero-sum game; the probability that random parameters will be suitable is infinitesimally small.
+We confirm this numerically with simulations shown in Fig.~\ref{fig:histograms}.
+Although $n$-species cycles are clearly possible (as in Fig.~\ref{fig:fragile}), for parameters drawn independently at random, communities always collapse to zero, one, or two species, regardless of the initial richness.
 
@@
 $\alpha_{ij} - \alpha_{ik} = \alpha_{lj} - \alpha_{lk}$ for all $i, j, k,$ and $l$).
-Even if mechanisms exist to generate the requisite qualitative patterns, inevitable quantitative variation in real-world communities will disrupt coexistence (Fig ???).
+Even if mechanisms exist to generate the requisite qualitative patterns, inevitable quantitative variation in real-world communities will disrupt coexistence (Fig.~\ref{fig:fragile}).
 Coexistence of $n > 2$ species -- even in the weak sense of neutral cycles -- is not robust to small changes in the model parameters.
@@
 It can be shown (see Supplemental Methods) that all $2 \times 2$ bimatrix games take the same general form as a rescaled zero-sum game, although the constant $c$ may be positive or negative, depending on the parameters.
-When $I_s$, the interaction coefficient identified by Bever et al., is negative, $c$ is positive, ensuring (neutral) stability.
-This condition amounts to an inequality constraint, rather than an equality constraint, and so it \emph{is} generally robust to small variations in model parameters (Fig ???).
+When $I_s$, the interaction coefficient identified by Bever \textit{et al.}, is negative, $c$ is positive, ensuring (neutral) stability.
+This condition amounts to an inequality constraint, rather than an equality constraint, and so it \emph{is} generally robust to small variations in model parameters (Fig.~\ref{fig:fragile}).
 As we can now see, the case $n = 2$ is unique in this regard.
@@
 	{Final community sizes with varying initial richness.
-We show the distribution of final richness (number of species, in red) for 500 (???) communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
-Parameters $\alpha_{ij}$ and $\beta_{i}$  were chosen randomly \emph{iid} from a standard uniform distribution, $U[0,1]$.
-For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.
-\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
+We show the distribution of final richness (number of species, in red) for 500 \textbf{(???)} communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
+Parameters $\alpha_{ij}$ and $\beta_{i}$  were sampled independently from a standard uniform distribution, $U[0,1]$.
+For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.~\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
 In agreement with the conjecture that coexistence of more than two species is vanishingly unlikely, we found that, regardless of the number of species initially present, every community collapsed to a subset of one or two surviving species.}
@@
 
+\begin{figure*}
+	\centering
+	%\includegraphics[width=\textwidth]{figure_3.pdf}
+	\caption[HERE GOES FIGURE 3]
+	{LABEL FOR FIG 3}
+	\label{fig:fragile}
+\end{figure*}
+
 \section{Discussion}
 
-The Bever model has played a central role in motivating PSF research, and continues to guide both theory and experiment in the fast-growing field \citep{bever2015maintenance,kandlikar2019winning,ke2020effects}.
+The Bever model has played a central role in motivating PSF research, and continues to guide both theory and experiment in this fast-growing field \citep{bever2015maintenance, kandlikar2019winning, ke2020effects}.
 Here, we extend the Bever model to any number of species, and highlight its equivalence to bimatrix game dynamics.
@@
 Indeed, the model has been wildly successfully in spurring research into PSFs.
-Alongside empirical study of these processes, other modeling approaches have emerged, accounting for more biological realism (e.g.
-\citep{umbanhowar2005simple,eppstein2007invasiveness,bever2010rooting}), or with the demonstrated capacity to produce multispecies coexistence (e.g.
-\citep{bonanomi2005negative,miller2021metapopulations}).
+Alongside empirical study of these processes, other modeling approaches have emerged, accounting for more biological realism \citep[e.g.,][]{umbanhowar2005simple,eppstein2007invasiveness,bever2010rooting}, or with the demonstrated capacity to produce multispecies coexistence  \citep[e.g.,][]{bonanomi2005negative,miller2021metapopulations}.
 Some of these are minor modifications of the Bever model framework; others build on distinct foundations \citep{ke2015incorporating,ke2020effects}.
@@
 Indeed, the absolute abundance model (Eq.~\ref{nspp_absolute}) used to derive our $n$-species frequency dynamics (Eqs.~\ref{nspp_relative}-\ref{matrix_form}) does not generally possess any fixed points, which is a basic requirement for species coexistence \citep{hutson1990existence,hutson1992permanence}.
-It is usually seen as desirable to study PSFs in the space of species frequencies, both because this facilitates connections to data, and because frequencies are considered a more appropriate metric for analyzing processes that stabilize coexistence (\citep{adler2007niche,eppinga2018frequency} but see \citep{kandlikar2019winning,ke2020effects}).
+It is usually seen as desirable to study PSFs in the space of species frequencies, both because this facilitates connections to data, and because frequencies are considered a more appropriate metric for analyzing processes that stabilize coexistence (\citet{adler2007niche,eppinga2018frequency}, but see \citet{kandlikar2019winning,ke2020effects}).
 But models that introduce frequencies through a natural constraint, such as competition for finite space, will likely produce more realistic dynamics.

### 2021-10-26 paper/supp_methods.tex
- Changed many \citep to \citet to avoid double parentheses - added a placeholder for figure 3, and updated references to figs in the text - minimal changes to main text - hard-coded references to equations in the main text across the Supplement - really great work!

@@
 
+\renewcommand{\theequation}{S\arabic{equation}} % to distinguish them from main text
+
 \usepackage{setspace}
@@
 governing the time-evolution of plant abundances $x_i$ and soil components $y_i$, where $p_i = x_i / \sum_j x_j$, $q_i = y_i / \sum_j y_j$, and Greek letters denote nonnegative parameters.
-These equations capture the assumptions outlined by Bever et al.
-\citep{bever1997incorporating} for two species and extend them straightforwardly to any $n$ species.
-Following the approach of Bever et al.
-for two species (and consistent with other generalizations of this model \citep{kulmatiski2008plant,eppinga2018frequency}), we derive dynamics for the $p_i$ by applying the chain rule:
+These equations capture the assumptions outlined by \citet{bever1997incorporating} for two species and extend them straightforwardly to any $n$ species.
+Following the approach of Bever \textit{et al.} for two species \citep[and consistent with other generalizations of this model,][]{kulmatiski2008plant,eppinga2018frequency}, we derive dynamics for the $p_i$ by applying the chain rule:
 
@@
 \end{align}
-This last expression is identical to the first line of Eq.~\ref{nspp_relative} in the Main Text.
+This last expression is identical to the first line of Eq.~5 in the Main Text.
 The dynamics for $q_i$ can be derived in exactly the same way (using the definitions $\beta_{ii} = \beta_i$ and $\beta_{ij} = 0$).
-The two terms of each per capita growth rate in Eq.~\ref{nspp_relative} have natural interpretations in the language and notation of linear algebra: $\sum_{j} \alpha_{ij} \, q_j$ is the $i$th component of the matrix-vector product $A \bm{q}$ and $\sum_{j, k} \alpha_{jk} p_j  q_k$ is the bilinear form $\bm{p}^T A \bm{q}$.
+The two terms of each per capita growth rate in Eq.~5 have natural interpretations in the language and notation of linear algebra: $\sum_{j} \alpha_{ij} \, q_j$ is the $i$th component of the matrix-vector product $A \bm{q}$ and $\sum_{j, k} \alpha_{jk} p_j  q_k$ is the bilinear form $\bm{p}^T A \bm{q}$.
 Here, $A$ (and $B$) is an $n \times n$ matrix and $\bm{p}$ ($\bm{q}$) is a vector of length $n$, as described in the Main Text.
-We can re-write Eq.~\ref{nspp_relative} as
+We can re-write Eq.~5 as
 \begin{align}
@@
 \end{align}
-which is Eq.~\ref{matrix_form} in the Main Text.
+which is Eq.~6 in the Main Text.
 
@@
 Then, our PSF model would be map exactly onto the full space of bimatrix game dynamics (rather than just a subset).
-However, all of the results we consider hold for arbitrary bimatrix games, meaning the same conclusions about the dynamics of Eqs.~\ref{nspp_relative}-\ref{matrix_form} would apply to this extended model, as well.
+However, all of the results we consider hold for arbitrary bimatrix games, meaning the same conclusions about the dynamics of Eqs.~5-6 would apply to this extended model, as well.
 
-We note two useful properties of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, as they will be important for the analysis that follows.
+We note two useful properties of Eqs.~5-6, as they will be important for the analysis that follows.
 First, we have the constraint $\sum_i p_i = \sum_i q_i = 1$ at every point in time.
@@
 
-In the remaining sections, we outline the main behaviors of Eqs.~\ref{nspp_relative}-\ref{matrix_form}, especially with regard to coexistence.
-We closely follow the treatment by Hofbauer and Sigmund \citep{hofbauer1998evolutionary}, and urge interested readers to consult this excellent introduction (see especially chapters 10 and 11).
+In the remaining sections, we outline the main behaviors of Eqs.~5-6, especially with regard to coexistence.
+We closely follow the treatment by \citet{hofbauer1998evolutionary}, and urge interested readers to consult this excellent introduction (see especially chapters 10 and 11).
 Here, we reproduce or sketch the essential details needed to justify the results in the Main Text.
@@
 This fixed point, $(\bm{p}^\star, \bm{q}^\star)$, must take the form $(c_1 B^{-1} \bm{1}, c_2 A^{-1} \bm{1})$ for some undetermined constants $c_1$ and $c_2$.
-Substituting this ansatz into the growth rates in Eq.~\ref{matrix_form} and equating them to zero, we have
+Substituting this ansatz into the growth rates in Eq.~6 and equating them to zero, we have
 
@@
 
-We can extend this picture beyond a local neighborhood of the coexistence equilibrium by considering the divergence of the vector field associated with Eqs.~\ref{nspp_relative}-\ref{matrix_form}.
+We can extend this picture beyond a local neighborhood of the coexistence equilibrium by considering the divergence of the vector field associated with Eqs.~5-6.
 The divergence, defined by $\sum_i \frac{\partial}{\partial p_i} \frac{dp_i}{dt} + \sum_i \frac{\partial}{\partial q_i} \frac{dq_i}{dt}$, measures the outgoing flux around a given point.
-It can be shown (see \citep{eshel1983coevolutionary} and \citep{hofbauer1998evolutionary}) that up to a change in velocity (i.e.
-rescaling time by a positive factor), the vector field corresponding to any bimatrix game dynamics has zero divergence everywhere in the interior of the positive orthant (i.e.
-where $p_i, q_i > 0$ for all $i$).
+It can be shown \citep[see][]{eshel1983coevolutionary, hofbauer1998evolutionary} that up to a change in velocity (i.e., rescaling time by a positive factor), the vector field corresponding to any bimatrix game dynamics has zero divergence everywhere in the interior of the positive orthant (i.e., where $p_i, q_i > 0$ for all $i$).
 
@@
 In the particular case of our model, $a_1 + a_2 = -\alpha_{11} + \alpha_{21} + \alpha_{12} - \alpha_{22} = -I_s$ and $b_1 + b_2 = -\beta_1 - \beta_2$.
-$c$ is positive (as needed for cycles) when these signs disagree; since $b_1 + b_2 = -\beta_1 - \beta_2$ is always negative, $a_1 + a_2$ must be positive, meaning $I_s < 0$, as found by Bever et al.
+$c$ is positive (as needed for cycles) when these signs disagree; since $b_1 + b_2 = -\beta_1 - \beta_2$ is always negative, $a_1 + a_2$ must be positive, meaning $I_s < 0$, as found by Bever \textit{et al}.
 
@@
 \end{equation}
-For the two-species model studied by Bever et al., these level sets precisely define the trajectories in the $(p, q)$ phase plane.
+For the two-species model studied by Bever \textit{et al.}, these level sets precisely define the trajectories in the $(p, q)$ phase plane.
 

### 2021-10-26 paper/compile.sh

@@ -1,0 +1,15 @@
+pdflatex paper.tex
+pdflatex paper.tex
+bibtex paper
+pdflatex paper.tex
+pdflatex paper.tex
+evince paper.pdf &
+
+pdflatex supp_methods.tex
+pdflatex supp_methods.tex
+bibtex supp_methods
+pdflatex supp_methods.tex
+pdflatex supp_methods.tex
+evince supp_methods.pdf &
+
+rm *.aux *.bbl *.blg *.log *.out *.synctex.gz

### 2021-10-26 code/dynamics_collapse.R
updated figure 3

@@ -4,42 +4,45 @@ library(reshape2)
 library(cowplot)
 library(gridExtra)
 library(RColorBrewer)
+library(ggplot2)
 
 
-data = read_csv('../data/zach_results.csv')
+data <- read_csv('../data/zach_results.csv')
+data <- data %>% mutate(flag = ifelse(p_l == 0 | p_l == 1, 'one', 'two'))
 dev.off()
 p1 = ggplot(data = data,
        aes(x = t, y = p_ab,
            color = as.factor(p_l)))+
-  geom_line(alpha = 0.2)+
-  geom_ma(ma_fun = EMA, n = 150, linetype = 'solid')+
+  geom_line(alpha = 1)+
   theme(aspect.ratio = 3/4,
             panel.background = element_blank(),
             panel.border = element_rect(fill = NA),
             legend.position = 'none',
             axis.title.x = element_text(size = 15),
             axis.title.y = element_text(size = 15),
-            axis.text = element_blank(),
+            #axis.text = element_blank(),
             axis.ticks = element_blank())+
-  scale_color_brewer(palette = 'Set1')+
+  scale_y_log10(limits = c(1e-39, 1))+
+  scale_color_brewer(palette = 'Set1', direction = -1)+
   labs(y = "Abundance",
        x = "Time")
 print(p1)
 
 p2 = ggplot(data = data,
-            aes(x = t, y = p_ab_per,
+            aes(x = t, y = p_ab_per))+
+  geom_line(aes(linetype = flag,
                 color = as.factor(p_l)))+
-  geom_line(alpha = 0.2)+
-  geom_ma(ma_fun = EMA, n = 150, linetype = 'solid')+
   theme(aspect.ratio = 3/4,
         panel.background = element_blank(),
         panel.border = element_rect(fill = NA),
         legend.position = 'none',
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15),
-        axis.text = element_blank(),
+        #axis.text = element_blank(),
         axis.ticks = element_blank())+
-  scale_color_brewer(palette = 'Set1')+
+  scale_linetype_manual(values = c('dashed', 'solid'))+
+  scale_y_log10(limits = c(1e-39, 1))+
+  scale_color_brewer(palette = 'Set1', direction = -1)+
   labs(y = "",
[…]

### 2021-10-26 code/functions.py

@@ -22,7 +22,7 @@ def model(t, z, A, B, n):
         import ipdb; ipdb.set_trace(context = 20)
     return(list(dpdt.reshape(n)) + list(dqdt.reshape(n)))
 
-def integrate_PSF(fun, t_span, z0, args, method = 'BDF'):
+def integrate_PSF(fun, t_span, z0, args, method = 'BDF', max_step = np.inf):
     '''
     Wrapper for integrator
     '''
@@ -143,7 +143,7 @@ def check_equilibrium_bis(plant_vector, soil_vector):
     '''
     n_plants = sum(plant_vector > 0)
     n_soils = sum(soil_vector > 0)
-    if n_plants > 2 or n_soils > 2:
+    if n_plants > 2:
         #If either plant/soil have more than 2, we have not reached
         #equilibnrium
         equilibrium = False

### 2021-10-26 code/quick_simulation.py

@@ -39,13 +39,14 @@ def main(argv):
     t_vec = []
     p_l = []
     present = np.ones(n)
-    t_max = 2000
+    t_max = 500
     while not equilibrium:
         #Integrate unperurbed system
         plant_ab_per, soil_ab_per, t_per = integrate_PSF(model, [0, t_max], 
                                                          z0_per, 
-                                                         (A_per, B, n), 
-                                                         method = 'RK45')
+                                                         (A_per, B, n),
+                                                         method = 'RK45', 
+                                                         max_step = 1)
         #Delete extinctions 
         rem_plant, rem_soil = remove_extinctions(plant_ab_per[:, -1], 
                                                  soil_ab_per[:, -1], 
@@ -61,8 +62,8 @@ def main(argv):
                                                    soil_ab_per[:, -1],
                                                    rem_soil)
         #Set extinct species to 0
-        plant_ab_per[ext_plant, :] = 0
-        soil_ab_per[ext_soil, :] = 0
+        plant_ab_per[plant_ab_per[:, -1] < tol, :] = 0
+        #soil_ab_per[soil_ab_per[:, -1] < tol, :] = 0
         #Store plant abundance
         p_ab_per += list(plant_ab_per.flatten())
         #Store time

### 2021-10-26 paper/cover_letter/PSF_cover.tex

\documentclass[11pt]{letter} 
\usepackage{graphicx} % Required for including pictures
\usepackage[export]{adjustbox} % to left align image
\usepackage[outdir=./]{epstopdf}

\pagestyle{empty} % Suppress headers and footers

\setlength\parindent{0cm} % Paragraph indentation

% Create a new command for the horizontal rule in the document which allows thickness specification
\makeatletter
\newcommand{\vhrulefill}[1]{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%----------------------------------------------------------------------------------------
%	DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\usepackage{geometry} % Required for adjusting page dimensions

\geometry{
	top=1cm, % Top margin
	bottom=1.5cm, % Bottom margin
	left=2.5cm, % Left margin
	right=2.5cm, % Right margin
	%showframe, % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	DEFINE CUSTOM COMMANDS
%----------------------------------------------------------------------------------------

\newcommand{\logo}[1]{\renewcommand{\logo}{#1}}

\newcommand{\Who}[1]{\renewcommand{\Who}{#1}}
\newcommand{\Title}[1]{\renewcommand{\Title}{#1}}

\newcommand{\headerlineone}[1]{\renewcommand{\headerlineone}{#1}}
\newcommand{\headerlinetwo}[1]{\renewcommand{\headerlinetwo}{#1}}

\newcommand{\authordetails}[1]{\renewcommand{\authordetails}{#1}}

%----------------------------------------------------------------------------------------
%	AUTHOR DETAILS STRUCTURE
%----------------------------------------------------------------------------------------

\newcommand{\authordetailsblock}{
	\hspace{\fill} % Move the author details to the far right
	\parbox[t]{0.58\textwidth}{ % Box holding the author details; width value specifies where it starts and ends, increase to move details left
		\footnotesize % Use a smaller font size for the details
		%\Who\\ % Author name
		\authordetails % The author details text, all italicised
	}
}

%----------------------------------------------------------------------------------------
%	HEADER STRUCTURE
%----------------------------------------------------------------------------------------

\address{
	\includegraphics[width=.38in,left]{./uchicago_shield.pdf} % Include the logo of author institution
	\hspace{0\textwidth} % this does nothing anymore due to left above, otherwise move the logo left and right
	\vskip -0.116\textheight~\\ % move the logo up or down
	\Large\hspace{0.2\textwidth} \hfill ~\\[-0.006\textheight] % First line of institution name, if desires, add before hfill. move up and down
	\hspace{0.08\textwidth}\includegraphics[width=1.6in]{./uchicago_text.pdf}\hfill \normalsize % Second line of institution name, adjust to move the logo left and right
	\makebox[0ex][r]{\textbf{
	%\Who\Title
	}}\hspace{0.01\textwidth} % Print author name and title with a little whitespace to the right
	~\\[-0.01\textheight] % Reduce the whitespace above the horizontal rule
	\hspace{0.08\textwidth}\vhrulefill{1pt} \\ % adjust length and with of header line. usually want this value to match the value for the second line of the inst name
	\noindent\authordetailsblock % Include the letter author's details on the right side of the page under the horizontal rule
	\hspace{-0.23\textwidth} % Horizontal position of the author details block, increase to move left, decrease to move right
	\vspace{-0.11\textheight} % Move the date and letter content up for a more compact look
}

%----------------------------------------------------------------------------------------
%	COMPOSE THE ENTIRE HEADER
%----------------------------------------------------------------------------------------

\renewcommand{\opening}[1]{
	{\centering\noindent\fromaddress\vspace{0.05\textheight} \\ % Print the header and from address here, add whitespace to move date down
	\hspace*{\longindentation}
	%\today
	\hspace*{\fill}\par} % Print today's date, remove \today to not display it
	{\raggedright \toname \\ \toaddress \par} % Print the to name and address
	\vspace{0.5cm} % White space after the to address
	\noindent #1 % Print the opening line
}

%----------------------------------------------------------------------------------------
%	SIGNATURE STRUCTURE
%----------------------------------------------------------------------------------------

\signature{\Who\Title} % The signature is a combination of the author's name and title

\renewcommand{\closing}[1]{
	\vspace{2.5mm} % Some whitespace after the letter content and before the signature
	\noindent % Stop paragraph indentation
	% \hspace*{9.5cm} % Move the signature right to the value of \longindentation
	\parbox{10cm}{
		\raggedright
		#1 % Print the signature text
		\vskip .2cm % Whitespace between the closing text and author's name for a physical signature
		Zachary R. Miller % Prints the value of \signature{}, i.e. author name and title
	}
}

%----------------------------------------------------------------------------------------
%	Your details
%----------------------------------------------------------------------------------------

\Who{Zachary R. Miller} % Your name

\Title{} % Your title, leave blank for no title

\authordetails{
	%\noindent \\
	%\Who \\
	%Department of Ecology \& Evolution\\ % Your department/institution
	%University of Chicago \\
	%1101 E.~57th St.\\ % Your address
	%Chicago, IL 60637, USA\\ % Your city, zip code, country, etc
	%zachmiller@uchicago.edu\\ % Your email address	
}

\begin{document}

%----------------------------------------------------------------------------------------
%	TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}{

	\today
}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\opening{Dear Editor,}

\frenchspacing

Please find attached the manuscript, ``No robust coexistence in a canonical model of plant-soil feedbacks,'' by Zachary R.
Miller, Pablo Lech\'{o}n, and Stefano Allesina, which we would like to submit to \textit{Ecology Letters} as a Letter.

In recent decades, plant-soil feedbacks (PSFs) -- indirect interactions between plants mediated by changes to local soil microbiota and chemistry -- have emerged as a leading mechanism to explain widespread coexistence in natural plant communities. The study of PSFs, both in theory and in the field, has been guided by a modeling framework introduced by Bever, Westover, and Antonovics in a landmark 1997 paper. However, the now-canonical ``Bever model'' is formulated for only two species, while natural communities may contain tens or hundreds of co-occurring plants. Bridging this gap has been a crucial ongoing challenge for the field.

Here, we extend the Bever model to include an arbitrary number of species and use results from evolutionary game theory to thoroughly characterize coexistence in this extended model. Our central finding is that coexistence of more than two species is virtually impossible within this framework. We show that multi-species coexistence can only arise under highly artificial parameter combinations, and collapses when these conditions are violated even slightly. Such fine-tuned conditions never arise by chance, are not robust to inevitable biological variation, and cannot explain coexistence in real-world systems.

These surprising results have significant implications for plant community ecology. The Bever model undergirds the analysis and interpretation of PSF experiments across a wide array of systems, and our findings suggest that applying two-species predictions to richer communities will yield incorrect inference about plant coexistence. Abundant empirical evidence indicates that PSFs do play an important role in maintaining diversity in plant communities; the apparent contradiction between these observations and the model predictions calls for critical reevaluation of model assumptions and urgent exploration of alternative theoretical frameworks.

Our analysis illuminates a long-standing disconnect between the low diversity of a simple, but widely utilized, PSF model, and the high diversity of real-world ecosystems. By demonstrating that this well-known model is incompatible with multi-species coexistence mediated by PSFs, we believe our results will kickstart new interest and progress in the theory behind this rapidly advancing area of ecology. Publishing these findings in Ecology Letters will allow us to reach the broad range of theoreticians and experimentalists across community, plant, and soil ecology who work with and rely on PSF models.

We thank you for your time, and we look forward to hearing from you.

\closing{Sincerely,}
\end{letter}
\end{document}

### 2021-10-27 paper/supp_methods.tex
added text explaining simulations

@@
 or even more compactly as
-\begin{align}
+\begin{align}\label{compact_form}
 \begin{cases}
@@
 
+\section{Numerical Simulations}
+In order to offer numerical evidence that complements our analytical findings, we  integrate  Eq.~\ref{compact_form} with 2, 3, 5, and 6 initial plant species and soil communities.
+For each case, we perform 5000 numerical integrations sampling the payoff matrices at random each time.
+At each integration we (1) sample non-singular payoff matrices $ A $ and $ B $ from the uniform distribution U(0, 1), and (2) integrate the system until an equilibrium with $\leq 2$ species is reached.
+
 \bibliographystyle{ecol_let}

### 2021-10-27 code/dynamics_collapse.R

@@ -9,99 +9,38 @@ library(ggplot2)
 
 data <- read_csv('../data/zach_results.csv')
 data <- data %>% mutate(flag = ifelse(p_l == 0 | p_l == 1, 'one', 'two'))
+colnames(data) <- c('X1', 'Perturbed', 't', 'p_l', 'Unperturbed', 'flag')
+
+data = data %>% 
+  pivot_longer(cols = c(Unperturbed, Perturbed),
+               names_to = "type",
+               values_to = "abundance")
+
 dev.off()
 p1 = ggplot(data = data,
-       aes(x = t, y = p_ab,
+       aes(x = t, y = abundance,
            color = as.factor(p_l)))+
   geom_line(alpha = 1)+
+  facet_wrap(.~factor(type, levels = rev(levels(factor(type)))))+
+  theme_classic() + 
   theme(aspect.ratio = 3/4,
             panel.background = element_blank(),
             panel.border = element_rect(fill = NA),
             legend.position = 'none',
             axis.title.x = element_text(size = 15),
-            axis.title.y = element_text(size = 15),
-            #axis.text = element_blank(),
-            axis.ticks = element_blank())+
-  scale_y_log10(limits = c(1e-39, 1))+
+            axis.title.y = element_text(size = 15))+
+  scale_y_log10(limits = c(1e-6, 1))+
   scale_color_brewer(palette = 'Set1', direction = -1)+
   labs(y = "Abundance",
        x = "Time")
 print(p1)
 
-p2 = ggplot(data = data,
-            aes(x = t, y = p_ab_per))+
-  geom_line(aes(linetype = flag,
-                color = as.factor(p_l)))+
-  theme(aspect.ratio = 3/4,
-        panel.background = element_blank(),
-        panel.border = element_rect(fill = NA),
-        legend.position = 'none',
-        axis.title.x = element_text(size = 15),
-        axis.title.y = element_text(size = 15),
-        #axis.text = element_blank(),
-        axis.ticks = element_blank())+
-  scale_linetype_manual(values = c('dashed', 'solid'))+
-  scale_y_log10(limits = c(1e-39, 1))+
-  scale_color_brewer(palette = 'Set1', direction = -1)+
-  labs(y = "",
-       x = "Time")
-print(p2)
-
 #Load matrices
 load('../data/example_perturbation_parameters.RData')
 
 A_pre = my_list$A_pre
-A_pre_melt = melt(A_pre)
-
[…]

### 2021-10-27 code/functions.py

@@ -179,19 +179,7 @@ def find_extinct_indices(plants, plants_rem, soils, soils_rem):
     ext_soil = list(set(soils) - set(soils_rem))
     o, ext_soil_ind, o = np.intersect1d(soils, ext_soil, 
                                         return_indices = True)
-    return(ext_plant_ind, ext_soil_ind)
+    if len(ext_plant_ind) != len(ext_soil_ind):
+        import ipdb; ipdb.set_trace(context = 20)
 
-def homogenize(vector):
-    '''
-    Homogenize a vector such that all the elements that are one integer
-    appart become equal to the first integer in the sequence
-    '''
-    n = len(vector)
-    new_vector = np.zeros(n)
-    for i in range(n-1):
-        if vector[i+1] - vector[i] == 1:
-            new_vector[i+1] = vector[i]
-        else:
-            new_vector[i+1] = vector[i+1]
-    return new_vector
-
+    return np.intersect1d(ext_plant_ind, ext_soil_ind)

### 2021-10-27 code/histograms_collapse.R

@@ -1,7 +1,7 @@
 library(tidyverse)
 
 data = read_csv('../data/results.csv')
-n_sim = round(max(data$n_sim), digits = -2) # number of simulations per inital richness
+dn_sim = round(max(data$n_sim), digits = -2) # number of simulations per inital richness
 
 # Remove non-convergent data (need to filter on n_p_f, rather than n_p)
 # Hopefully we can deal with these cases in the simulation code and eventually

### 2021-10-27 code/plant_feedback_model.py

@@ -81,12 +81,10 @@ def main(argv):
             #(3) the number of integration cycles surpasses our limit. 
             while (not equilibrium) & (n_int <= n_int_max):
                 #Find extinct indices of plants and soils 
-                ext_plant, ext_soil = find_extinct_indices(plant_ab[:, -1], 
-                                                           rem_plant, 
-                                                           soil_ab[:, -1],
-                                                           rem_soil)
-                A = remove_extinctions_matrix(A, ext_plant)
-                B = remove_extinctions_matrix(B, ext_soil)
+                ext_ind = find_extinct_indices(plant_ab[:, -1], rem_plant, 
+                                                soil_ab[:, -1], rem_soil)
+                A = remove_extinctions_matrix(A, ext_ind)
+                B = remove_extinctions_matrix(B, ext_ind)
                 #Set initial conditions to the final state of previous 
                 #integration
                 z0 = list(np.hstack([rem_plant, rem_soil]))

### 2021-10-27 paper/paper.tex
edited my last name

@@
 \date{}
-\author{Zachary R. Miller$^{1*}$ (zachmiller@uchicago.edu), Pablo Lech\'{o}n$^{1}$ (plechon@uchicago.edu),\\ and Stefano Allesina$^{1,2}$ (sallesina@uchicago.edu) \\
+\author{Zachary R. Miller$^{1*}$ (zachmiller@uchicago.edu), Pablo Lech\'{o}n-Alonso$^{1}$ (plechon@uchicago.edu),\\ and Stefano Allesina$^{1,2}$ (sallesina@uchicago.edu) \\
 	\\
@@
 
-\noindent Author contributions: ZRM and SA performed the mathematical analysis, and PL implemented the numerical simulations.
+\noindent Author contributions: ZRM and SA performed the mathematical analysis, and PLA implemented the numerical simulations.
 ZRM wrote the manuscript, and all authors edited and discussed the text.

### 2021-10-27 code/functions.py
added figure with 5000 simulations

@@ -179,7 +179,5 @@ def find_extinct_indices(plants, plants_rem, soils, soils_rem):
     ext_soil = list(set(soils) - set(soils_rem))
     o, ext_soil_ind, o = np.intersect1d(soils, ext_soil, 
                                         return_indices = True)
-    if len(ext_plant_ind) != len(ext_soil_ind):
-        import ipdb; ipdb.set_trace(context = 20)
 
     return np.intersect1d(ext_plant_ind, ext_soil_ind)

### 2021-10-27 code/histograms_collapse.R

@@ -1,12 +1,12 @@
 library(tidyverse)
 
 data = read_csv('../data/results.csv')
-dn_sim = round(max(data$n_sim), digits = -2) # number of simulations per inital richness
+n_sim = round(max(data$n_sim), digits = -2) # number of simulations per inital richness
 
 # Remove non-convergent data (need to filter on n_p_f, rather than n_p)
 # Hopefully we can deal with these cases in the simulation code and eventually
 # remove this line
-data = data %>% filter(n_p_f != -1)
+#data = data %>% filter(n_p_f != -1)
 
 # Reshape data for easier ggplotting
 data = data %>% 
@@ -14,6 +14,8 @@ data = data %>%
   pivot_longer(cols = c(n_p_s, n_p_f),
                names_to = "type",
                values_to = "num_species")
+
+data$type = factor(data$type, levels = c('n_p_s', 'n_p_f'))
 
 data %>% ggplot(aes(x = num_species,
                     after_stat(density),
@@ -23,23 +25,23 @@ data %>% ggplot(aes(x = num_species,
   geom_histogram(position = "identity",
                  alpha = 0.5,
                  binwidth = 1) +
-  facet_wrap(.~n_p, scales = "free") + 
+  facet_wrap(.~n_p, nrow = 4) + 
   theme_classic() + 
   theme(axis.title = element_text(size = 15),
          axis.text = element_text(size = 15),
          strip.text = element_text(size = 15),
          legend.title = element_blank(),
-         legend.position = c(0.35, 0.85),
+         legend.position = c(0.78, 0.94),
          legend.text = element_text(size = 12, hjust = 0.1),
          panel.background = element_blank(), 
          panel.border = element_rect(color = 'black', fill = NA), 
          aspect.ratio = 0.5) + 
-  scale_fill_manual(values = c('red', 'blue'),
-                    labels = c('Final', 'Initial')) + 
-  scale_color_manual(values = c('red', 'blue'),
-                     labels = c('Final', 'Initial'))+
-  scale_linetype_manual(values = c('solid', 'dashed'),
-                        labels = c('Final', 'Initial'))+
+  scale_fill_manual(values = c('blue', 'red'),
+                    labels = c('Initial', 'Final')) + 
+  scale_color_manual(values = c('blue', 'red'),
+                     labels = c('Initial', 'Final'))+
+  scale_linetype_manual(values = c('dashed', 'solid'),
[…]

### 2021-10-27 code/plant_feedback_model.py

@@ -62,13 +62,15 @@ def main(argv):
             #Delete extinctions (only those that happen simultaneously in both
             #vectors). Failing to do so would be a conceptual mistake, since
             #If a plant goes extinct, so does its soil.
-            rem_plant, rem_soil = remove_extinctions(plant_ab[:, -1], 
-                                                     soil_ab[:, -1], 
-                                                     tol) 
+            #Set extinctions to 0
+            plant_ab[plant_ab[:, -1] < tol, :] = 0
+            #rem_plant, rem_soil = remove_extinctions(plant_ab[:, -1], 
+            #                                         soil_ab[:, -1], 
+            #                                         tol) 
             #Check if we have reached equilibrium
-            equilibrium = check_equilibrium(rem_plant, rem_soil)
+            equilibrium = check_equilibrium_bis(plant_ab[:, -1], soil_ab[:, -1])
             #Update number of plants
-            n_plant = len(rem_plant)
+            #n_plant = len(rem_plant)
             ###################################################################
             #NOTE THAT THIS IS NOT ENTIRELY CORRECT, NEED TO ADD A SNIPPET
             #CHECKING FOR PARTNERSHIP/ZERO-SUM GAME
@@ -81,17 +83,18 @@ def main(argv):
             #(3) the number of integration cycles surpasses our limit. 
             while (not equilibrium) & (n_int <= n_int_max):
                 #Find extinct indices of plants and soils 
-                ext_ind = find_extinct_indices(plant_ab[:, -1], rem_plant, 
-                                                soil_ab[:, -1], rem_soil)
-                A = remove_extinctions_matrix(A, ext_ind)
-                B = remove_extinctions_matrix(B, ext_ind)
+                #ext_ind = find_extinct_indices(plant_ab[:, -1], rem_plant, 
+                #                                soil_ab[:, -1], rem_soil)
+                #A = remove_extinctions_matrix(A, ext_ind)
+                #B = remove_extinctions_matrix(B, ext_ind)
                 #Set initial conditions to the final state of previous 
                 #integration
-                z0 = list(np.hstack([rem_plant, rem_soil]))
+                #z0 = list(np.hstack([rem_plant, rem_soil]))
+                z0 = list(np.hstack([plant_ab[:, -1], soil_ab[:, -1]]))
                 #Solve diferential equations again
                 #Re-integrate system
                 plant_ab, soil_ab, t = integrate_PSF(model, [1, 2000], z0, 
-                                                     (A, B, n_plant))
+                                                     (A, B, n))
[…]

### 2021-10-27 paper/cover_letter/PSF_cover.tex
final edits and updated cover letter

@@
 
-\opening{Dear Editor,}
+\opening{Dear Dr. Coulson,}
 
@@
 Please find attached the manuscript, ``No robust coexistence in a canonical model of plant-soil feedbacks,'' by Zachary R.
-Miller, Pablo Lech\'{o}n, and Stefano Allesina, which we would like to submit to \textit{Ecology Letters} as a Letter.
+Miller, Pablo Lech\'{o}n-Alonso, and Stefano Allesina, which we would like to submit to \textit{Ecology Letters} as a Letter.
 
-In recent decades, plant-soil feedbacks (PSFs) -- indirect interactions between plants mediated by changes to local soil microbiota and chemistry -- have emerged as a leading mechanism to explain widespread coexistence in natural plant communities.
+In recent decades, plant-soil feedbacks (PSFs) -- indirect interactions between plants mediated by changes to local soil microbiota -- have emerged as a leading mechanism to explain widespread coexistence in plant communities.
 The study of PSFs, both in theory and in the field, has been guided by a modeling framework introduced by Bever, Westover, and Antonovics in a landmark 1997 paper.
@@
 We show that multi-species coexistence can only arise under highly artificial parameter combinations, and collapses when these conditions are violated even slightly.
-Such fine-tuned conditions never arise by chance, are not robust to inevitable biological variation, and cannot explain coexistence in real-world systems.
+Such fine-tuned conditions never arise by chance, are not robust to inevitable biological variation, and cannot explain coexistence in real-world communities.
 
@@
 The Bever model undergirds the analysis and interpretation of PSF experiments across a wide array of systems, and our findings suggest that applying two-species predictions to richer communities will yield incorrect inference about plant coexistence.
-Abundant empirical evidence indicates that PSFs do play an important role in maintaining diversity in plant communities; the apparent contradiction between these observations and the model predictions calls for critical reevaluation of model assumptions and urgent exploration of alternative theoretical frameworks.
+Abundant empirical evidence indicates that PSFs do play an important role in maintaining diversity in plant communities; the apparent contradiction between these observations and the lack of coexistence predicted by this model calls for critical reevaluation of model assumptions and urgent exploration of alternative theory.
 
-Our analysis illuminates a long-standing disconnect between the low diversity of a simple, but widely utilized, PSF model, and the high diversity of real-world ecosystems.
-By demonstrating that this well-known model is incompatible with multi-species coexistence mediated by PSFs, we believe our results will kickstart new interest and progress in the theory behind this rapidly advancing area of ecology.
-Publishing these findings in Ecology Letters will allow us to reach the broad range of theoreticians and experimentalists across community, plant, and soil ecology who work with and rely on PSF models.
+\textbf{Statement of novelty.} Our analysis illuminates a long-standing disconnect between the low diversity of a simple, but widely utilized, PSF model, and the high diversity of real-world ecosystems.
+To our knowledge, this work is the first to characterize the dynamics of the Bever model in full generality, representing an advancement on recent efforts (e.g.
+Eppinga \textit{et al.}, \textit{Nature Ecology \& Evolution 2018}; Mack \textit{et al.}, \textit{PLoS ONE} 2019).
+By demonstrating that this well-known PSF model is incompatible with multi-species coexistence, we believe our results will kickstart new interest and progress in the theory behind this rapidly advancing area of ecology.
+Publishing our findings in \textit{Ecology Letters} will allow us to reach the broad range of theoreticians and experimentalists across community, plant, and soil ecology who work with and rely on PSF models.
+
+We have not previously published research directly related to the Bever model for PSFs.
+In a recent manuscript (Miller \& Allesina 2021, \textit{in submission}), we introduce a mathematical model for a broad class of environmentally-mediated interactions, including PSFs as a typical example.
+This approach, which we cite in our present study, demonstrates how and when PSFs can facilitate the coexistence of many species, and represents an alternative framework for modeling these complex interactions.
 

### 2021-10-27 paper/paper.tex
included figure 3 in manuscript
final edits and updated cover letter

@@
 \noindent Author contributions: ZRM and SA performed the mathematical analysis, and PLA implemented the numerical simulations.
-ZRM wrote the manuscript, and all authors edited and discussed the text.
-\\
+ZRM wrote the manuscript, and all authors edited and discussed the text.\\
 
-\noindent Data accessibility: Code for reproducing the numerical simulations is available at [ADD REPOSITORY].\\
+\noindent Data accessibility: Code for reproducing the numerical simulations is available at\\ \url{https://github.com/pablolich/plant_soil_feedback}.\\
 
 \noindent Article type: Letter\\
-Number of words in abstract: 137\\
-Number of words in main text: \\
-Number of references: 51 \\
+Number of words in abstract: 135\\
+Number of words in main text: 3297\\
+Number of references: 52 \\
 Number of figures: 3
+
+\pagebreak
 
@@
 However, the primary modeling framework used to study PSFs considers only two plant species, and we lack clear theoretical expectations for how these complex interactions play out in communities with natural levels of diversity.
-Here, we demonstrate that this canonical model for PSFs is equivalent to a well-studied model from evolutionary game theory, and using this equivalence, we are able to characterize the dynamics with an arbitrary number of plant species.
+Here, we demonstrate that this canonical model for PSFs is equivalent to a well-studied model from evolutionary game theory, and we use this equivalence to characterize the dynamics with an arbitrary number of plant species.
 Surprisingly, we find that coexistence of more than two species is virtually impossible in this model, suggesting that alternative theoretical frameworks are needed to describe feedbacks observed in diverse natural communities.
@@
 
-A central finding of the analysis by Bever \textit{et al.} was that the denominator of $q_i^*$, which they termed the ``interaction coefficient'', $I_s$ ($= \alpha_{ii} - \alpha_{ij} + \alpha_{jj} - \alpha_{ji}$), controls the model dynamics: When $I_s > 0$, which represents a community with positive feedbacks, the equilibrium in Eq.~\ref{2spp_eq} is unstable, and the two species cannot coexist.
+A central finding of the analysis by Bever \textit{et al.} was that the denominator of $q_i^\star$, which they termed the ``interaction coefficient'', $I_s$ ($= \alpha_{ii} - \alpha_{ij} + \alpha_{jj} - \alpha_{ji}$), controls the model dynamics: When $I_s > 0$, which represents a community with positive feedbacks, the equilibrium in Eq.~\ref{2spp_eq} is unstable, and the two species cannot coexist.
 On the other hand, when $I_s < 0$, the equilibrium is neutrally stable, and the dynamics cycle around it, providing a form of non-equilibrium coexistence.
 In fact, these conclusions depend on the existence of a feasible equilibrium (i.e.
-positive equilibrium values), which further requires that $\alpha_{ii} < \alpha_{ji}$, for both $i, j = 1, 2$, in order for the model to exhibit coexistence \citep{bever1997incorporating,ke2015incorporating}.
+positive equilibrium values), which further requires that $\alpha_{ii} < \alpha_{ji}$ for both $i, j = 1, 2$, in order for the model to exhibit coexistence \citep{bever1997incorporating,ke2015incorporating}.
 
@@
 
-Of course, most natural plant communities feature more than two coexisting species, and it is precisely in the most diverse communities where mechanisms of coexistence hold the greatest interest \citep{van2013plant}.
+Of course, most natural plant communities feature more than two coexisting species, and it is precisely in the most diverse communities that mechanisms of coexistence hold the greatest interest \citep{van2013plant}.
 While it is not immediately clear how to generalize Eq.~\ref{2spp_relative} to more than two species, Eq.~\ref{2spp_absolute} is naturally extended by maintaining the assumption that the overall growth rate for any plant is a weighted average of its growth rate in each soil type:
@@
 \end{align}
-From Eq.~\ref{nspp_absolute}, one can derive the $n$-species analogue of Eq.~\ref{2spp_relative} (Fig.~\ref{fig:concept}),
+From Eq.~\ref{nspp_absolute}, one can derive the $n$-species analogue of Eq.~\ref{2spp_relative},
 
@@
 \end{align}
-giving the dynamics for species and soil component frequencies. Eq.~\ref{nspp_relative} is conveniently expressed in matrix form as
+giving the dynamics for species and soil component frequencies (Fig.~\ref{fig:concept}). Eq.~\ref{nspp_relative} is conveniently expressed in matrix form as
 
@@
 Using these constraints, one can easily show that the Bever model (Eq.~\ref{2spp_relative}) is a special case of Eqs.~\ref{nspp_relative} and \ref{matrix_form} when $n = 2$.
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{figure_1.pdf}
-	\caption[Structure of the $n$-species Bever model]
-	{The model described by Eqs.~\ref{nspp_relative}-\ref{matrix_form} is shown here graphically.
-Plant species (green squares) promote the growth of their respective soil components (brown circles) at a rate $\beta_i$ (gray arrows).
-In turn, growth of each plant is governed by the mix of soil components present in the system, with the effect of soil component $j$ on species $i$ quantified by the parameter $\alpha_{ij}$ (black arrows).
-This model is a straightforward extension of the model proposed by \citet{bever1997incorporating} to an arbitrary number of species.
-Note only selected parameter labels are shown for clarity.}
-	\label{fig:concept}
-\end{figure*}
 
@@
 Another notable property of bimatrix game dynamics is that the vector field defined by the model equations is divergence-free or incompressible \citep[see][for a proof]{hofbauer1998evolutionary}.
-The divergence theorem from vector calculus \citep{arfken1985mathematical} then tells us that Eq.~\ref{matrix_form} cannot have any attractors -- that is, regions of the phase space that ``pull in'' trajectories -- with multiple species.
+The divergence theorem from vector calculus \citep{arfken1985mathematical} then dictates that Eq.~\ref{matrix_form} cannot have any attractors -- that is, regions of the phase space that ``pull in'' trajectories -- with multiple species.
 This rules out coexistence in a stable limit cycle or other non-equilibrium attractors (e.g.
@@
 The level curves of $H$ form closed orbits around the equilibrium when the equilibrium is neutrally stable.
-Thus, $H$ implicitly defines the trajectories of the model, and can be used to determine characteristics such as the amplitude of oscillations arising from particular initial frequencies.
+Thus, $H$ implicitly defines the trajectories of the model, and can be used to determine characteristics such as the amplitude of oscillations arising from particular initial frequencies \citep{volterra1926variazioni}.
 
@@
 Indeed, they do; however, for more than two species, these conditions are very severe.
-The model in Eq.~\ref{matrix_form} supports oscillations with $n$ species -- for any $n$ -- if matrices $A$ and $B$ satisfy a precise relationship (Fig.~\ref{fig:fragile}; see Supplemental Methods for details).
+The model in Eq.~\ref{matrix_form} supports oscillations with $n$ species -- for any $n$ -- if matrices $A$ and $B$ satisfy a precise relationship (see Supplemental Methods for details).
 In particular, the model parameters must satisfy the conditions $\alpha_{ij} = \gamma_i + \delta_j$ for some constants $\gamma_i, \delta_i$ in $i = 1, \dots, n$ (when $i \neq j$), and $\alpha_{ii} = \gamma_i + \delta_i- c \beta_i$ (where $c$ is a positive constant independent of $i$).
@@
 Ecologically, these conditions mean there is a fixed effect of each soil type and plant species identity, and the growth rate of plant $i$ in soil type $j$ is the additive combination of these two, with no interaction effects.
-The only exception is for plants growing in their own soil type, which must experience a fitness cost ($\gamma_i + \delta_i - \alpha_{ii}$) exactly proportional to the rate at which they promote growth of the soil type ($\beta_i$).
+The only exception is for plants growing in their own soil type, which must experience a fitness cost ($\gamma_i + \delta_i - \alpha_{ii}$) exactly proportional to the rate at which they promote growth of their soil type ($\beta_i$).
 These conditions clearly extend the intuitive notion that each plant must have a disadvantage in its corresponding soil type to allow coexistence.
@@
 We confirm this numerically with simulations shown in Fig.~\ref{fig:histograms}.
-Although $n$-species cycles are clearly possible (as in Fig.~\ref{fig:fragile}), for parameters drawn independently at random, communities always collapse to zero, one, or two species, regardless of the initial richness.
+Although $n$-species cycles are clearly possible (as in Fig.~\ref{fig:fragile}), for parameters drawn independently at random, communities always collapse to one or two species, regardless of the initial richness.
 
@@
 As we can now see, the case $n = 2$ is unique in this regard.
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{figure_2.pdf}
-	\caption[Final community sizes with varying initial richness]
-	{Final community sizes with varying initial richness.
-We show the distribution of final richness (number of species, in red) for 500 \textbf{(???)} communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
-Parameters $\alpha_{ij}$ and $\beta_{i}$  were sampled independently from a standard uniform distribution, $U[0,1]$.
-For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.~\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
-In agreement with the conjecture that coexistence of more than two species is vanishingly unlikely, we found that, regardless of the number of species initially present, every community collapsed to a subset of one or two surviving species.}
-	\label{fig:histograms}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	%\includegraphics[width=\textwidth]{figure_3.pdf}
-	\caption[HERE GOES FIGURE 3]
-	{LABEL FOR FIG 3}
-	\label{fig:fragile}
-\end{figure*}
 
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_1.pdf}
+	\caption[Structure of the $n$-species Bever model]
+	{The model described by Eqs.~\ref{nspp_relative}-\ref{matrix_form} is shown here graphically.
+Plant species (green squares) promote the growth of their respective soil components (brown circles) at a rate $\beta_i$ (gray arrows).
+In turn, growth of each plant is governed by the mix of soil components present in the system, with the effect of soil component $j$ on species $i$ quantified by the parameter $\alpha_{ij}$ (black arrows).
+This model is a straightforward extension of the model proposed by \citet{bever1997incorporating} to an arbitrary number of species.
+Note only selected parameter labels are shown for clarity.}
+	\label{fig:concept}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_2.pdf}
+	\caption[Final community sizes with varying initial richness]
+	{Final community sizes with varying initial richness.
+We show the distribution of final richness (number of species, in red) for 5000 communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
+Parameters $\alpha_{ij}$ and $\beta_{i}$ were sampled independently from a standard uniform distribution, $U(0,1)$.
+For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.~\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
+In agreement with the conjecture that coexistence of more than two species is vanishingly unlikely, we found that regardless of the number of species initially present, every community collapsed to a subset of one or two surviving species.}
+	\label{fig:histograms}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_3.pdf}
+	\caption[Coexistence of three or more species is not robust]{Coexistence of three or more species is not robust.
+It is possible to obtain neutrally stable oscillations with any number of species if the model parameters constitute a rescaled zero-sum game (see text for conditions).
+Here, for example, we show sustained oscillations with 4 species using fine-tuned parameters, $A_u$ (left).
+However, if we randomly perturb $A_u$ by a small amount to obtain new parameters, $A_p$, the dynamics quickly collapse to a two-species subset (right).
+Any slight perturbation is enough to disrupt coexistence; for this example, the parameters $A_u$ and $A_p$ are highly correlated (inset) and differ in value by less than 3\% on average.}
+	\label{fig:fragile}
+\end{figure*}
+
 \end{document}

### 2021-10-27 paper/supp_methods.tex
edited my last name
final edits and updated cover letter

--- 2021-10-26 paper/supp_methods.tex
@@
 \date{}
-\author{Zachary R. Miller$^{1*}$, Pablo Lech\'{o}n$^{1}$, and Stefano Allesina$^{1,2}$ \\
+\author{Zachary R. Miller$^{1*}$, Pablo Lech\'{o}n-Alonso$^{1}$, and Stefano Allesina$^{1,2}$ \\
 	\\
@@
 These equations capture the assumptions outlined by \citet{bever1997incorporating} for two species and extend them straightforwardly to any $n$ species.
-Following the approach of Bever \textit{et al.} for two species \citep[and consistent with other generalizations of this model,][]{kulmatiski2008plant,eppinga2018frequency}, we derive dynamics for the $p_i$ by applying the chain rule:
+Following the approach of Bever \textit{et al.} for two species \citep[and consistent with other generalizations of this model, e.g.,][]{kulmatiski2008plant,eppinga2018frequency}, we derive dynamics for the $p_i$ by applying the chain rule:
 
@@
 \end{align}
-Here, we have used the fact that $A \bm{q^*} \propto \bm{1}$.
+Here, we have used the fact that $A \bm{q^\star} \propto \bm{1}$.
 Notice that, because the factors in parentheses in Eq.~\ref{reduced} are zero at equilibrium, these community matrix calculations are valid even for $i = j$.
@@
 	&= c \sum_i \beta_i \left( -p_i^\star \, q_i + q_i^\star \, p_i \right)
-	\intertext{and because $q_i^* = p_i^* = \frac{Z}{\beta_{i}}$, with $Z$ the normalizing constant,}
+	\intertext{and because $q_i^\star = p_i^\star = \frac{Z}{\beta_{i}}$, with $Z$ the normalizing constant,}
 	&= c \, Z \, \sum_i (-q_i + p_i) \\
@@
 
-Each orbit remains in the level set defined by the initial conditions,$(\bm{p}_0, \bm{q}_0)$:
+Each orbit remains in the level set defined by the initial conditions, $(\bm{p}_0, \bm{q}_0)$:
 
@@
 For the two-species model studied by Bever \textit{et al.}, these level sets precisely define the trajectories in the $(p, q)$ phase plane.
+
+\section{Numerical Simulations}
+To complement our analytical findings, we investigated the dynamics of many randomly parameterized communities using numerical simulations.
+In particular, we integrated  Eq.~\ref{compact_form} with 2, 3, 5, or 6 initial plant species and corresponding soil components.
+For each case, we sampled 5000 parameter sets at random and integrated the dynamics in Python using SciPy's (version 1.7.1) solve\_ivp function with the ``BDF'' method.
+We sampled non-singular payoff matrices $ A $ and $ B $ with each non-zero element drawn independently from the uniform distribution U(0, 1).
+For every choice of parameters, we integrated the system until a subset with $\leq 2$ species was reached (which occurred in all cases).
 

### 2021-10-27 code/dynamics_collapse.R

@@ -5,76 +5,78 @@ library(cowplot)
 library(gridExtra)
 library(RColorBrewer)
 library(ggplot2)
-
-
-data <- read_csv('../data/zach_results.csv')
-data <- data %>% mutate(flag = ifelse(p_l == 0 | p_l == 1, 'one', 'two'))
-colnames(data) <- c('X1', 'Perturbed', 't', 'p_l', 'Unperturbed', 'flag')
-
-data = data %>% 
-  pivot_longer(cols = c(Unperturbed, Perturbed),
-               names_to = "type",
-               values_to = "abundance")
-
-dev.off()
-p1 = ggplot(data = data,
-       aes(x = t, y = abundance,
-           color = as.factor(p_l)))+
-  geom_line(alpha = 1)+
-  facet_wrap(.~factor(type, levels = rev(levels(factor(type)))))+
-  theme_classic() + 
-  theme(aspect.ratio = 3/4,
-            panel.background = element_blank(),
-            panel.border = element_rect(fill = NA),
-            legend.position = 'none',
-            axis.title.x = element_text(size = 15),
-            axis.title.y = element_text(size = 15))+
-  scale_y_log10(limits = c(1e-6, 1))+
-  scale_color_brewer(palette = 'Set1', direction = -1)+
-  labs(y = "Abundance",
-       x = "Time")
-print(p1)
-
-#Load matrices
-load('../data/example_perturbation_parameters.RData')
-
-A_pre = my_list$A_pre
-
-A_post = my_list$A_post
-
-data = data.frame(x = as.vector(A_pre), y = as.vector(A_post))
-
-r_2 = paste("R^2 == ", round(summary(lm(data$y~data$x))$r.squared, digits = 3))
-
-correlation_plot = ggplot(data = data,
-                          aes(x = x, y = y))+
-  geom_point()+
-  geom_abline(slope = 1)+
-  theme(aspect.ratio = 1,
-        panel.background = element_rect(fill = 'transparent',
-                                        colour = NA),
-        plot.background = element_rect(fill = 'transparent', 
-                                       colour = NA),
-        panel.border = element_rect(fill = 'transparent'),
-        legend.position = 'none',
-        axis.title = element_text(size = 13),
-        axis.text = element_text(size = 8))+
-  scale_x_continuous(breaks = c(0.6, 0.9, 1.2, 1.5),
[…]

### 2021-10-28 paper/biorxiv_v1.tex
removed cover letter directory

--- 2021-10-27 paper/paper.tex
@@
 \date{}
-\author{Zachary R. Miller$^{1*}$ (zachmiller@uchicago.edu), Pablo Lech\'{o}n-Alonso$^{1}$ (plechon@uchicago.edu),\\ and Stefano Allesina$^{1,2}$ (sallesina@uchicago.edu) \\
+\author{Zachary R. Miller$^{1*}$, Pablo Lech\'{o}n-Alonso$^{1}$, and Stefano Allesina$^{1,2}$\\
 	\\
@@
 	\\
-	\normalsize{*corresponding author (mailing: 1101 E. 57th Street, Chicago, IL 60637; phone: 484-331-4593)}\\
+	\normalsize{*corresponding author (zachmiller@uchicago.edu)}\\
 }
@@
 
-\noindent Running title: No coexistence in a plant-soil feedback model\\
-Keywords: plant-soil feedbacks, legacy effects, coexistence, Bever model, bimatrix games, replicator equation \\
-
-\noindent Author contributions: ZRM and SA performed the mathematical analysis, and PLA implemented the numerical simulations.
-ZRM wrote the manuscript, and all authors edited and discussed the text.\\
-
-\noindent Data accessibility: Code for reproducing the numerical simulations is available at\\ \url{https://github.com/pablolich/plant_soil_feedback}.\\
-
-\noindent Article type: Letter\\
-Number of words in abstract: 135\\
-Number of words in main text: 3297\\
-Number of references: 52 \\
-Number of figures: 3
-
-\pagebreak
-
 \linenumbers
@@
 \begin{abstract}
-Plant-soil feedbacks (PSFs) are thought to represent a crucial mechanism generating frequency dependent dynamics in plant communities.
+Plant-soil feedbacks (PSFs) are thought to represent a crucial mechanism generating frequency-dependent dynamics in plant communities.
 Negative feedbacks, in particular, are routinely invoked to explain coexistence and the maintenance of diversity in species-rich communities.
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.9\textwidth]{figure_1.pdf}
+	\caption[Structure of the $n$-species Bever model]
+	{The model described by Eqs.~\ref{nspp_relative}-\ref{matrix_form} is shown here graphically.
+Plant species (green squares) promote the growth of their respective soil components (brown circles) at a rate $\beta_i$ (gray arrows).
+In turn, growth of each plant is governed by the mix of soil components present in the system, with the effect of soil component $j$ on species $i$ quantified by the parameter $\alpha_{ij}$ (black arrows).
+This model is a straightforward extension of the model proposed by \citet{bever1997incorporating} to an arbitrary number of species.
+Note only selected parameter labels are shown for clarity.}
+	\label{fig:concept}
+\end{figure*}
+
 \subsection{Equivalence to bimatrix game dynamics}
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_2.pdf}
+	\caption[Final community sizes with varying initial richness]
+	{Final community sizes with varying initial richness.
+We show the distribution of final richness (number of species, in red) for 5000 communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
+Parameters $\alpha_{ij}$ and $\beta_{i}$ were sampled independently from a standard uniform distribution, $U(0,1)$.
+For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.~\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
+In agreement with the conjecture that coexistence of more than two species is vanishingly unlikely, we found that regardless of the number of species initially present, every community collapsed to a subset with one or two surviving species.}
+	\label{fig:histograms}
+\end{figure*}
+
 Ecologically, these conditions mean there is a fixed effect of each soil type and plant species identity, and the growth rate of plant $i$ in soil type $j$ is the additive combination of these two, with no interaction effects.
@@
 As we can now see, the case $n = 2$ is unique in this regard.
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figure_3.pdf}
+	\caption[Coexistence of three or more species is not robust]{Coexistence of three or more species is not robust.
+It is possible to obtain neutrally stable oscillations with any number of species if the model parameters constitute a rescaled zero-sum game (see text for conditions).
+Here, for example, we show sustained oscillations with 4 species using fine-tuned parameters, $A_u$ (left).
+However, if we randomly perturb $A_u$ by a small amount to obtain new parameters, $A_p$, the dynamics quickly collapse to a two-species subset (right).
+Any slight perturbation is enough to disrupt coexistence; for this example, the parameters $A_u$ and $A_p$ are highly correlated (inset) and differ in value by less than 3\% on average.}
+	\label{fig:fragile}
+\end{figure*}
 
@@
 Ecologists have repeatedly found that intuitions from two-species models can generalize (or fail to generalize) to more diverse communities in surprising ways \citep{strobeck1973n,smale1976differential,barabas2016effect}.
-Our analysis provides another illustration of the fact that ``more is different'' \citep{anderson1972more} in ecology, and highlights the importance of developing theory for species-rich communities.
+Our analysis provides another illustration of the fact that ``more is different'' \citep{anderson1972more} in ecology, and highlights the importance of developing theory for species-rich communities.\\
 
-\section*{Acknowledgments}
-We thank ...
+\noindent \textbf{Data accessibility:} Code for reproducing the numerical simulations is available at\\ \url{https://github.com/pablolich/plant_soil_feedback}.\\
 
@@
 \bibliography{references}
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{figure_1.pdf}
-	\caption[Structure of the $n$-species Bever model]
-	{The model described by Eqs.~\ref{nspp_relative}-\ref{matrix_form} is shown here graphically.
-Plant species (green squares) promote the growth of their respective soil components (brown circles) at a rate $\beta_i$ (gray arrows).
-In turn, growth of each plant is governed by the mix of soil components present in the system, with the effect of soil component $j$ on species $i$ quantified by the parameter $\alpha_{ij}$ (black arrows).
-This model is a straightforward extension of the model proposed by \citet{bever1997incorporating} to an arbitrary number of species.
-Note only selected parameter labels are shown for clarity.}
-	\label{fig:concept}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{figure_2.pdf}
-	\caption[Final community sizes with varying initial richness]
-	{Final community sizes with varying initial richness.
-We show the distribution of final richness (number of species, in red) for 5000 communities governed by the $n$-species Bever model, initialized with 2, 3, 5, or 6 plant species.
-Parameters $\alpha_{ij}$ and $\beta_{i}$ were sampled independently from a standard uniform distribution, $U(0,1)$.
-For each random parameterization at each level of initial richness, we integrated the dynamics of Eq.~\ref{matrix_form} until the system reached a periodic orbit or until only one species remained.
-In agreement with the conjecture that coexistence of more than two species is vanishingly unlikely, we found that regardless of the number of species initially present, every community collapsed to a subset of one or two surviving species.}
-	\label{fig:histograms}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{figure_3.pdf}
-	\caption[Coexistence of three or more species is not robust]{Coexistence of three or more species is not robust.
-It is possible to obtain neutrally stable oscillations with any number of species if the model parameters constitute a rescaled zero-sum game (see text for conditions).
-Here, for example, we show sustained oscillations with 4 species using fine-tuned parameters, $A_u$ (left).
-However, if we randomly perturb $A_u$ by a small amount to obtain new parameters, $A_p$, the dynamics quickly collapse to a two-species subset (right).
-Any slight perturbation is enough to disrupt coexistence; for this example, the parameters $A_u$ and $A_p$ are highly correlated (inset) and differ in value by less than 3\% on average.}
-	\label{fig:fragile}
-\end{figure*}
 

### 2021-10-28 code/histograms_collapse.R
added nice sketch figure1

@@ -24,7 +24,8 @@ data %>% ggplot(aes(x = num_species,
                     linetype = type)) + 
   geom_histogram(position = "identity",
                  alpha = 0.5,
-                 binwidth = 1) +
+                 binwidth = 1,
+                 size = 0.3) +
   facet_wrap(.~n_p, nrow = 4) + 
   theme_classic() + 
   theme(axis.title = element_text(size = 15),

### 2021-10-28 code/plant_feedback_model.py
.

@@ -69,12 +69,6 @@ def main(argv):
             #                                         tol) 
             #Check if we have reached equilibrium
             equilibrium = check_equilibrium_bis(plant_ab[:, -1], soil_ab[:, -1])
-            #Update number of plants
-            #n_plant = len(rem_plant)
-            ###################################################################
-            #NOTE THAT THIS IS NOT ENTIRELY CORRECT, NEED TO ADD A SNIPPET
-            #CHECKING FOR PARTNERSHIP/ZERO-SUM GAME
-            ###################################################################
             #Initialize counter for number of integrations (integration cycles 
             #will stop if n_int is more than n_int_max)
             n_int = 0
